---
title: "Filtering of differential loop Anchors in Cohesin mut AML by TSS presence"
author: "Alexander Fischer"
date: "02 2023"
output: html
---

# Loading libraries and set directories
```{r}
# libraries
library(reshape2)
library(ggplot2)
library(ggrepel)
library(corrplot)
library(cowplot)
# paths
DIR_DATA="/misc/data"
WORKDIR=file.path(DIR_DATA,"analysis/project_cohesin/Cohesin_AML")

CHIPDIR<-file.path(WORKDIR,"ChIP_analysis")
peakdir<-file.path(CHIPDIR,"peaks")
diffpeakdir<-file.path(CHIPDIR,"diffPeaks")
diffpeakdirH3K<-file.path(WORKDIR,"ChIP_analysis/H3K27ac/diffPeaks")
diffpeakdirSA1vsSA2<-file.path(diffpeakdir,"SA1vsSA2")
GEXdir<-file.path(WORKDIR,"RNAseq/Analysis/Resulttables/RNAseq_Cohesin_AML_AF3")
ATACdir<-file.path(WORKDIR,"ATAC")

LOOPDIR=file.path(WORKDIR,"HiC/loops")
diffanchorsn2t=file.path(LOOPDIR,"differentialanchorsn2t")
AllEorEP=file.path(diffanchorsn2t,"H3K27overlap")
AnchorsOverlapPairs<-file.path(diffanchorsn2t,"AnchorsOverlapPairs")
TSSOverlap<-file.path(AnchorsOverlapPairs,"TSSoverlap")
DTmatrixdir<-file.path(WORKDIR,"HiC/figures/Heatmaps/DeeptoolsMatrices/n2tresults/PairedAnch")

# variables
MUTs<-c("SA2mut","RAD21mut")
Filts<-c("up.FC1", "down.FC1")
strands<-c("plus","minus")
```

# get the top RAD21 peak for each Anchor to be used for centering in plots
## read files
```{r}
#get RAD21 peak coverage
diffpeaks<-read.table(file.path(diffpeakdir,"SA2mutvsCTRL.RAD21.Peaks_DESEQ.model.all.txt"),header=TRUE,row.names=1)
#get TSS RAD21 association
TSSRAD21<-read.table(file.path(TSSOverlap,"RAD21centred","TSS.RAD21.2wayoverlap.bed"),header=F, sep="\t")
TSSRAD21<-TSSRAD21[,c(1:4,7:10)]
colnames(TSSRAD21)<-c("TSS_chr","TSS_start","TSS_end","TSS_name","RAD21_chr","RAD21_start","RAD21_end","RAD21_peakID")
RAD21TSS<-read.table(file.path(TSSOverlap,"RAD21centred","TSS.RAD21.2wayoverlap.RAD.bed"),header=F, sep="\t")
RAD21TSS<-RAD21TSS[,c(1:4,7:10)]
colnames(RAD21TSS)<-c("RAD21_chr","RAD21_start","RAD21_end","RAD21_peakID","TSS_chr","TSS_start","TSS_end","TSS_name")


#read in EP/EE/PP anchor H3K27ac-RAD21 overlap files
types<-c("EP_anchors_enhancers","EP_anchors_promoters","PP_anchors_promoters.plus","PP_anchors_promoters.minus","EE_anchors_enhancers.plus","EE_anchors_enhancers.minus")
RAD21anchoverllist<-list()
for (MUT in MUTs) {
  for (FILT in Filts){
      for (type in types){
path<-file.path(TSSOverlap,"RAD21centred",paste0(MUT,"_",FILT,"_paired_",type,".RAD21.2wayoverlap.bed"))
RAD21anchoverllist[[paste0(FILT,"_",MUT,"_",type)]]<-read.table(path, header=F, sep="\t") #read in  the bedfiles as tab sep. table
RAD21anchoverllist[[paste0(FILT,"_",MUT,"_",type)]]<-RAD21anchoverllist[[paste0(FILT,"_",MUT,"_",type)]][,c(1:4,5:8,11)]
colnames(RAD21anchoverllist[[paste0(FILT,"_",MUT,"_",type)]])<-c("Anchorchr","Anchorstart","Anchorend","LoopID","RAD21_chr","RAD21_start","RAD21_end","RAD21_peakID","RAD21anchoverlap") #rename columns
}}}

#how many have no RAD21?
RAD21check<-lapply(RAD21anchoverllist,function(x) nrow(subset(x,RAD21anchoverlap=="0")))
RAD21checkdf<-data.frame()
for (MUT in MUTs) {
  for (FILT in Filts){
      for (type in types){
RAD21checkdf[paste0(FILT,"_",MUT),paste0(type,"_all")]<-length(unique(RAD21anchoverllist[[paste0(FILT,"_",MUT,"_",type)]]$LoopID))
RAD21checkdf[paste0(FILT,"_",MUT),paste0(type,"_noRAD21")]<-nrow(subset(RAD21anchoverllist[[paste0(FILT,"_",MUT,"_",type)]],RAD21anchoverlap=="0"))
}}}
###ca 10% of enhancersL; less for promoters


#read in reg-struc (RS) anchors RAD21 overlaps
typesRS<-c("ES_anchors_enhancers","ES_anchors_structural","PS_anchors_promoters","PS_anchors_structural","PsP_anchors_promoters","PsP_anchors_structuralPromoter","EsP_anchors_enhancers","EsP_anchors_structuralPromoter")
RAD21anchoverllist_RS<-list()
for (MUT in MUTs) {
  for (FILT in Filts){
      for (type in typesRS){
path<-file.path(TSSOverlap,"RAD21centred","reg_struc_pairs",paste0(MUT,"_",FILT,"_paired_",type,".RAD21.2wayoverlap.bed"))
RAD21anchoverllist_RS[[paste0(FILT,"_",MUT,"_",type)]]<-read.table(path, header=F, sep="\t") #read in  the bedfiles as tab sep. table
RAD21anchoverllist_RS[[paste0(FILT,"_",MUT,"_",type)]]<-RAD21anchoverllist_RS[[paste0(FILT,"_",MUT,"_",type)]][,c(1:4,5:8,11)]
colnames(RAD21anchoverllist_RS[[paste0(FILT,"_",MUT,"_",type)]])<-c("Anchorchr","Anchorstart","Anchorend","LoopID","RAD21_chr","RAD21_start","RAD21_end","RAD21_peakID","RAD21anchoverlap") #rename columns
}}}

#how many have no RAD21?
RAD21check2<-lapply(RAD21anchoverllist_RS,function(x) nrow(subset(x,RAD21anchoverlap=="0")))
RAD21checkdf2<-data.frame()
for (MUT in MUTs) {
  for (FILT in Filts){
      for (type in types){
RAD21checkdf2[paste0(FILT,"_",MUT),paste0(type,"_all")]<-length(unique(RAD21anchoverllist_RS[[paste0(FILT,"_",MUT,"_",type)]]$LoopID))
RAD21checkdf2[paste0(FILT,"_",MUT),paste0(type,"_noRAD21")]<-nrow(subset(RAD21anchoverllist_RS[[paste0(FILT,"_",MUT,"_",type)]],RAD21anchoverlap=="0"))
}}}
```


## filter for top RAD21 peak in each anchor and genereate RAD21 centred output bedfiles ## use the Top RAD21 peak by by basemean (DESEQ2 analysis)
```{r}
LoopTSSRADlist<-list()
for (MUT in MUTs) {
  for (FILT in Filts){
      for (type in types){
df<-RAD21anchoverllist[[paste0(FILT,"_",MUT,"_",type)]]
Looplist<-list()
Loopsinset<-unique(df$LoopID)
for (Loop in Loopsinset) {
Loop_AllRAD<-subset(df,df$LoopID == Loop) #all entries for the anchor of this loopID
Loop_AllRADcpm<-subset(diffpeaks,row.names(diffpeaks) %in% Loop_AllRAD$RAD21_peakID) #look up RAD21 peaks in table
TopRAD<-row.names(Loop_AllRADcpm[which.max(Loop_AllRADcpm$baseMean),]) #select the max baseMean peak 
Loop_1RAD<-subset(Loop_AllRAD,Loop_AllRAD$RAD21_peakID == TopRAD) #select entry with max baseMean peak
Looplist[[Loop]]<-Loop_1RAD #save in list
}
LoopRADdf<-do.call(rbind.data.frame, Looplist)
LoopTSSRADlist[[paste0(FILT,"_",MUT,"_",type)]]<-LoopRADdf
}}}

#now filter out  pairs where one anchor has no RAD21

summary(LoopTSSRADlist[[paste0("down.FC1_SA2mut_EP_anchors_enhancers")]]$LoopID %in% LoopTSSRADlist[[paste0("down.FC1_SA2mut_EP_anchors_promoters")]]$LoopID)
summary(LoopTSSRADlist[[paste0("down.FC1_SA2mut_EP_anchors_promoters")]]$LoopID %in% LoopTSSRADlist[[paste0("down.FC1_SA2mut_EP_anchors_enhancers")]]$LoopID)
#   Mode   FALSE    TRUE 
#logical      15     115 
LoopRADlistfilt<-list()
for (MUT in MUTs) {
  for (FILT in Filts){
LoopRADlistfilt[[paste0(FILT,"_",MUT,"_EP_anchors_enhancers")]]<-subset(LoopTSSRADlist[[paste0(FILT,"_",MUT,"_EP_anchors_enhancers")]],LoopTSSRADlist[[paste0(FILT,"_",MUT,"_EP_anchors_enhancers")]]$LoopID %in% LoopTSSRADlist[[paste0(FILT,"_",MUT,"_EP_anchors_promoters")]]$LoopID)
LoopRADlistfilt[[paste0(FILT,"_",MUT,"_EP_anchors_promoters")]]<-subset(LoopTSSRADlist[[paste0(FILT,"_",MUT,"_EP_anchors_promoters")]],LoopTSSRADlist[[paste0(FILT,"_",MUT,"_EP_anchors_promoters")]]$LoopID %in% LoopTSSRADlist[[paste0(FILT,"_",MUT,"_EP_anchors_enhancers")]]$LoopID)

LoopRADlistfilt[[paste0(FILT,"_",MUT,"_PP_anchors_promoters.plus")]]<-subset(LoopTSSRADlist[[paste0(FILT,"_",MUT,"_PP_anchors_promoters.plus")]],LoopTSSRADlist[[paste0(FILT,"_",MUT,"_PP_anchors_promoters.plus")]]$LoopID %in% LoopTSSRADlist[[paste0(FILT,"_",MUT,"_PP_anchors_promoters.minus")]]$LoopID)
LoopRADlistfilt[[paste0(FILT,"_",MUT,"_PP_anchors_promoters.minus")]]<-subset(LoopTSSRADlist[[paste0(FILT,"_",MUT,"_PP_anchors_promoters.minus")]],LoopTSSRADlist[[paste0(FILT,"_",MUT,"_PP_anchors_promoters.minus")]]$LoopID %in% LoopTSSRADlist[[paste0(FILT,"_",MUT,"_PP_anchors_promoters.plus")]]$LoopID)

LoopRADlistfilt[[paste0(FILT,"_",MUT,"_EE_anchors_enhancers.plus")]]<-subset(LoopTSSRADlist[[paste0(FILT,"_",MUT,"_EE_anchors_enhancers.plus")]],LoopTSSRADlist[[paste0(FILT,"_",MUT,"_EE_anchors_enhancers.plus")]]$LoopID %in% LoopTSSRADlist[[paste0(FILT,"_",MUT,"_EE_anchors_enhancers.minus")]]$LoopID)
LoopRADlistfilt[[paste0(FILT,"_",MUT,"_EE_anchors_enhancers.minus")]]<-subset(LoopTSSRADlist[[paste0(FILT,"_",MUT,"_EE_anchors_enhancers.minus")]],LoopTSSRADlist[[paste0(FILT,"_",MUT,"_EE_anchors_enhancers.minus")]]$LoopID %in% LoopTSSRADlist[[paste0(FILT,"_",MUT,"_EE_anchors_enhancers.plus")]]$LoopID)
}}


#write new bedfiles of matched pairs
dir.create(file.path(TSSOverlap,"TopRAD21centred"))

for (MUT in MUTs) {
  for (FILT in Filts){
      for (type in types){
#get coords
bed<-LoopRADlistfilt[[paste0(FILT,"_",MUT,"_",type)]][,c(5:7,4,8)]
bed$V6<-1 #important add the column with "1" otherwise deeptools will change the name according to coordinates when clustering!
#colnames(bed)<-c("chr","start","stop","Loopname")
#write as bedfiles
write.table(bed, file.path(TSSOverlap,"TopRAD21centred",paste0(MUT,"_",FILT,"_",type,"_RAD21top.bed")),row.names=F,quote=F,sep="\t",col.names=F)
}}}
```


## stats and barplot
```{r}
RAD21EPdf<-data.frame()
for (MUT in MUTs) {
  for (FILT in Filts){
RAD21EPdf[paste0(FILT,"_",MUT),"EP"]<-nrow(LoopRADlistfilt[[paste0(FILT,"_",MUT,"_EP_anchors_enhancers")]])
RAD21EPdf[paste0(FILT,"_",MUT),"EE"]<-nrow(LoopRADlistfilt[[paste0(FILT,"_",MUT,"_EE_anchors_enhancers.plus")]])
RAD21EPdf[paste0(FILT,"_",MUT),"PP"]<-nrow(LoopRADlistfilt[[paste0(FILT,"_",MUT,"_PP_anchors_promoters.plus")]])
}}
RAD21EPdfm<-melt(t(RAD21EPdf))
colnames(RAD21EPdfm)<-c("LoopType","con","Freq")
RAD21EPdfm$group<-factor(c(rep("SA2mut",6),rep("RAD21mut",6)),levels=c("SA2mut","RAD21mut"))
RAD21EPdfm$change<-c(rep("strengthened",3),rep("weakened",3),rep("strengthened",3),rep("weakened",3))
RAD21EPdfm$LoopType<-factor(RAD21EPdfm$LoopType,levels=c("EE", "EP", "PP"))

p<-ggplot(RAD21EPdfm, aes(change, Freq,fill=LoopType)) +
  geom_bar(position = "stack", stat="identity") +
    xlab("") + ylab("number of loops") +
    ggtitle("Paired Cohesin-H3K27ac-anchor loop type summary") +
 theme(
  axis.text.x = element_text(angle=90,hjust=0.95,vjust=0.2),
  axis.title=element_text(size=20,face="bold"), plot.title = element_text(size = 20, face = "bold"), 
  panel.border = element_blank(),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.background = element_blank(),
  axis.line = element_line(colour = "black"),) + ylim(0,310)+
 guides(fill=guide_legend(title=""))+
 facet_wrap(~ group)
ggsave(p,file=file.path(WORKDIR,"HiC","figures","Loopstats","Diff.Paired_EP_Cohesin_looptypes.pdf"))
```

## same for unpaired H3K27ac loops: filter for top RAD21 peak in each anchor and genereate RAD21 centred output bedfiles
```{r}
LoopTSSRADlist2<-list()
for (MUT in MUTs) {
  for (FILT in Filts){
      for (type in typesRS){
df<-RAD21anchoverllist_RS[[paste0(FILT,"_",MUT,"_",type)]]
Looplist<-list()
Loopsinset<-unique(df$LoopID)
for (Loop in Loopsinset) {
Loop_AllRAD<-subset(df,df$LoopID == Loop) #all entries for the anchor of this loopID
Loop_AllRADcpm<-subset(diffpeaks,row.names(diffpeaks) %in% Loop_AllRAD$RAD21_peakID) #look up RAD21 peaks in table
TopRAD<-row.names(Loop_AllRADcpm[which.max(Loop_AllRADcpm$baseMean),]) #select the max baseMean peak 
Loop_1RAD<-subset(Loop_AllRAD,Loop_AllRAD$RAD21_peakID == TopRAD) #select entry with max baseMean peak
Looplist[[Loop]]<-Loop_1RAD #save in list
}
LoopRADdf<-do.call(rbind.data.frame, Looplist)
LoopTSSRADlist2[[paste0(FILT,"_",MUT,"_",type)]]<-LoopRADdf
}}}

#now filter out  pairs where one anchor has no RAD21

summary(LoopTSSRADlist2[[paste0("down.FC1_SA2mut_ES_anchors_structural")]]$LoopID %in% LoopTSSRADlist2[[paste0("down.FC1_SA2mut_ES_anchors_enhancers")]]$LoopID)
summary(LoopTSSRADlist2[[paste0("down.FC1_SA2mut_ES_anchors_enhancers")]]$LoopID %in% LoopTSSRADlist2[[paste0("down.FC1_SA2mut_ES_anchors_enhancers")]]$LoopID)


LoopTSSRADlistfilt2<-list()
for (MUT in MUTs) {
  for (FILT in Filts){
LoopTSSRADlistfilt2[[paste0(FILT,"_",MUT,"_ES_anchors_enhancers")]]<-subset(LoopTSSRADlist2[[paste0(FILT,"_",MUT,"_ES_anchors_enhancers")]],LoopTSSRADlist2[[paste0(FILT,"_",MUT,"_ES_anchors_enhancers")]]$LoopID %in% LoopTSSRADlist2[[paste0(FILT,"_",MUT,"_ES_anchors_structural")]]$LoopID)
LoopTSSRADlistfilt2[[paste0(FILT,"_",MUT,"_ES_anchors_structural")]]<-subset(LoopTSSRADlist2[[paste0(FILT,"_",MUT,"_ES_anchors_structural")]],LoopTSSRADlist2[[paste0(FILT,"_",MUT,"_ES_anchors_structural")]]$LoopID %in% LoopTSSRADlist2[[paste0(FILT,"_",MUT,"_ES_anchors_enhancers")]]$LoopID)
LoopTSSRADlistfilt2[[paste0(FILT,"_",MUT,"_PS_anchors_promoters")]]<-subset(LoopTSSRADlist2[[paste0(FILT,"_",MUT,"_PS_anchors_promoters")]],LoopTSSRADlist2[[paste0(FILT,"_",MUT,"_PS_anchors_promoters")]]$LoopID %in% LoopTSSRADlist2[[paste0(FILT,"_",MUT,"_PS_anchors_structural")]]$LoopID)
LoopTSSRADlistfilt2[[paste0(FILT,"_",MUT,"_PS_anchors_structural")]]<-subset(LoopTSSRADlist2[[paste0(FILT,"_",MUT,"_PS_anchors_structural")]],LoopTSSRADlist2[[paste0(FILT,"_",MUT,"_PS_anchors_structural")]]$LoopID %in% LoopTSSRADlist2[[paste0(FILT,"_",MUT,"_PS_anchors_promoters")]]$LoopID)
LoopTSSRADlistfilt2[[paste0(FILT,"_",MUT,"_PsP_anchors_promoters")]]<-subset(LoopTSSRADlist2[[paste0(FILT,"_",MUT,"_PsP_anchors_promoters")]],LoopTSSRADlist2[[paste0(FILT,"_",MUT,"_PsP_anchors_promoters")]]$LoopID %in% LoopTSSRADlist2[[paste0(FILT,"_",MUT,"_PsP_anchors_structuralPromoter")]]$LoopID)
LoopTSSRADlistfilt2[[paste0(FILT,"_",MUT,"_PsP_anchors_structuralPromoter")]]<-subset(LoopTSSRADlist2[[paste0(FILT,"_",MUT,"_PsP_anchors_structuralPromoter")]],LoopTSSRADlist2[[paste0(FILT,"_",MUT,"_PsP_anchors_structuralPromoter")]]$LoopID %in% LoopTSSRADlist2[[paste0(FILT,"_",MUT,"_PsP_anchors_promoters")]]$LoopID)
LoopTSSRADlistfilt2[[paste0(FILT,"_",MUT,"_EsP_anchors_enhancers")]]<-subset(LoopTSSRADlist2[[paste0(FILT,"_",MUT,"_EsP_anchors_enhancers")]],LoopTSSRADlist2[[paste0(FILT,"_",MUT,"_EsP_anchors_enhancers")]]$LoopID %in% LoopTSSRADlist2[[paste0(FILT,"_",MUT,"_EsP_anchors_structuralPromoter")]]$LoopID)
LoopTSSRADlistfilt2[[paste0(FILT,"_",MUT,"_EsP_anchors_structuralPromoter")]]<-subset(LoopTSSRADlist2[[paste0(FILT,"_",MUT,"_EsP_anchors_structuralPromoter")]],LoopTSSRADlist2[[paste0(FILT,"_",MUT,"_EsP_anchors_structuralPromoter")]]$LoopID %in% LoopTSSRADlist2[[paste0(FILT,"_",MUT,"_EsP_anchors_enhancers")]]$LoopID)
}}

#write new bedfiles of matched pairs
dir.create(file.path(TSSOverlap,"TopRAD21centred","reg_struc_pairs"))
for (MUT in MUTs) {
  for (FILT in Filts){
      for (type in typesRS){
#get coords
bed<-LoopTSSRADlistfilt2[[paste0(FILT,"_",MUT,"_",type)]][,c(5:7,4,8)]
bed$V6<-1 #important add the column with "1" otherwise deeptools will change the name according to coordinates when clustering!
#colnames(bed)<-c("chr","start","stop","Loopname")
#write as bedfiles
write.table(bed, file.path(TSSOverlap,"TopRAD21centred","reg_struc_pairs",paste0(MUT,"_",FILT,"_",type,"_RAD21top.bed")),row.names=F,quote=F,sep="\t",col.names=F)
}}}
```



## stats and barplot: reg-struct
```{r}
RAD21EPdf2<-data.frame()
for (MUT in MUTs) {
  for (FILT in Filts){
RAD21EPdf2[paste0(FILT,"_",MUT),"ES"]<-nrow(LoopTSSRADlistfilt2[[paste0(FILT,"_",MUT,"_ES_anchors_enhancers")]])
RAD21EPdf2[paste0(FILT,"_",MUT),"PS"]<-nrow(LoopTSSRADlistfilt2[[paste0(FILT,"_",MUT,"_PS_anchors_promoters")]])
RAD21EPdf2[paste0(FILT,"_",MUT),"PsP"]<-nrow(LoopTSSRADlistfilt2[[paste0(FILT,"_",MUT,"_PsP_anchors_structuralPromoter")]])
RAD21EPdf2[paste0(FILT,"_",MUT),"EsP"]<-nrow(LoopTSSRADlistfilt2[[paste0(FILT,"_",MUT,"_EsP_anchors_structuralPromoter")]])
}}
RAD21EPdfm2<-melt(t(RAD21EPdf2))
colnames(RAD21EPdfm2)<-c("LoopType","con","Freq")
RAD21EPdfm2$group<-factor(c(rep("SA2mut",8),rep("RAD21mut",8)),levels=c("SA2mut","RAD21mut"))
RAD21EPdfm2$change<-c(rep("strengthened",4),rep("weakened",4),rep("strengthened",4),rep("weakened",4))
RAD21EPdfm2$LoopType<-factor(RAD21EPdfm2$LoopType,levels=c("ES", "PS", "PsP","EsP"))

p<-ggplot(RAD21EPdfm2, aes(change, Freq,fill=LoopType)) +
  geom_bar(position = "stack", stat="identity") +
    xlab("") + ylab("number of loops") +
    ggtitle("Regulatory-structural \n Cohesin-H3K27ac-anchor differential loop type summary") +
 theme(
  axis.text.x = element_text(angle=90,hjust=0.95,vjust=0.2),
  axis.title=element_text(size=20,face="bold"), plot.title = element_text(size = 20, face = "bold"), 
  panel.border = element_blank(),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.background = element_blank(),
  axis.line = element_line(colour = "black"),) + ylim(0,650)+
 guides(fill=guide_legend(title=""))+
 facet_wrap(~ group)
ggsave(p,file=file.path(WORKDIR,"HiC","figures","Loopstats","Diff.Paired_reg-struc_Cohesin_looptypes.pdf"))

#stats and barplot: reg-struct + reg-reg
RAD21EPdf_RR_RS<-rbind(RAD21EPdfm,RAD21EPdfm2)
RAD21EPdf_RR_RS$LoopType<-factor(RAD21EPdf_RR_RS$LoopType,levels=c("EP","PP","EE","EsP","PsP","ES","PS"))
p<-ggplot(RAD21EPdf_RR_RS, aes(change, Freq,fill=LoopType)) +
  geom_bar(position = "stack", stat="identity") +
    xlab("") + ylab("number of loops") +
    ggtitle("Differential loops \n with H3K27ac-marked cohesin-anchor(s) ") +
  scale_fill_manual(values = c("EP"="darkgoldenrod3","PP"="indianred1","EE"="gold2","EsP"="khaki","PsP"="lightpink","PS"="ivory3","ES"="ivory4"),
  labels = c("Enhancer-Promoter","Promoter-Promoter","Enhancer-Enhancer","Enhancer-structural Promoter","Promoter-structural Promoter","Enhancer-Structural","Promoter-Structural"))+
 theme(
  axis.text.x = element_text(angle=90,hjust=0.95,vjust=0.2),
  axis.title=element_text(size=20,face="bold"), plot.title = element_text(size = 20, face = "bold"), 
  panel.border = element_blank(),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.background = element_blank(),
  axis.line = element_line(colour = "black"),) + ylim(0,1000)+
 guides(fill=guide_legend(title=""))+
 facet_wrap(~ group)
ggsave(p,file=file.path(WORKDIR,"HiC","figures","Loopstats","Diff.Paired_Cohesin_reg-reg_reg-struct_looptypes.CPM1.pdf"))
```