---
title: "Diff Loop Anchor stats in Cohesin-mutant AML"
author: "Alexander Fischer"
output: html_document
---

# set paths and load libraries
```{r setup, include=FALSE}
DIR_DATA="/misc/data"
WORKDIR=file.path(DIR_DATA,"analysis/project_cohesin/Cohesin_AML/HiC")
FIGURESDIR=file.path(WORKDIR,"figures/Loopstats")
LOOPDIR=file.path(WORKDIR,"loops")
diffanchorsn2t=file.path(LOOPDIR,"differentialanchorsn2t")
library(ggplot2)
library(reshape2)
```

# Define variables
```{r}
MUTs<-c("SA2mut","RAD21mut")
Filts<-c("up.FC1", "down.FC1")
anchorcategories<-c("CohesinAssEnhancers.CTCF",
"CohesinAssEnhancers.wo.CTCF",
"wo.CohesinAssEnhancers.RAD21.CTCF",
"wo.CohesinAssEnhancers.RAD21.woCTCF",
"wo.CohesinAssEnhancers.wo.RAD21.CTCF",
"wo.CohesinAssEnhancers.wo.RAD21.woCTCF.H3K27ac",
"wo.CohesinAssEnhancers.wo.RAD21.woCTCF.woH3K27ac")
```

# General summary of loop anchors by ChIP feature overlap
## read in the Anchor-ChIP-peak overlap bedfiles and count the entries by category
```{r}
loopstatslist<-vector(mode = "list", length = 0)
for (MUT in MUTs) {
loopstatslist[[MUT]]<-data.frame()
for (cat in anchorcategories) {
  for (FILT in Filts){
path<-file.path(diffanchorsn2t,"H3K27overlap",paste0(MUT,"vsCTRL.loopAnchors.red.", FILT, ".",cat,".Overlap.bed"))
info <- file.info(path)
if (info$size == 0L) { #check if file if empty 
loopstatslist[[MUT]][cat,FILT]<-0 #add a 0 count for empty files
} else {
loopstatslist[[MUT]][cat,FILT]<-nrow(read.table(path, header=F, sep="\t")) #add count for proper files
}
}}}
```

## STACKED BARPLOTS
```{r}
plotlist<-vector(mode = "list", length = 0)
for (MUT in MUTs) {
df <- loopstatslist[[MUT]]
df$categories <- row.names(df)
data.m <- melt(df, id.vars="categories")
write.table(data.m,file=file.path(FIGURESDIR,paste0(MUT,"diffloopAnchorFeatures.txt")),sep="\t",row.names=FALSE,quote=FALSE)
# Create a stacked barplot with absolute numbers
p1<-ggplot(data.m , aes(x = variable, y = value, fill = categories)) +
    geom_bar(stat = "identity",position="stack") +
    scale_fill_manual(values = c("CohesinAssEnhancers.CTCF" = "#a69f1a", "CohesinAssEnhancers.wo.CTCF" = "#ffee03",
    "wo.CohesinAssEnhancers.RAD21.CTCF" = "#9c2a85", "wo.CohesinAssEnhancers.RAD21.woCTCF"= "#dd23b8", "wo.CohesinAssEnhancers.wo.RAD21.CTCF"= "#1411dc",
    "wo.CohesinAssEnhancers.wo.RAD21.woCTCF.H3K27ac"= "#a4d833", "wo.CohesinAssEnhancers.wo.RAD21.woCTCF.woH3K27ac"= "#aaaaa1"),
    labels=c("Cohesin E/P + CTCF", "Cohesin E/P", "Cohesin + CTCF", "Cohesin", "CTCF", "E/P", "other")) + 
    xlab(" ") + ylab("number of loop anchors") +
    ggtitle(paste0("Differential loop anchors in ", MUT)) +
    theme_classic() +
    theme(axis.title=element_text(size=20,face="bold"),axis.text.x = element_text(size=15),axis.text.y = element_text(size=15),
    legend.text=element_text(size=15))
ggsave(file.path(FIGURESDIR,paste0(MUT,"diffloopAnchorFeatures.pdf")), plot=p1) #directly save to pdf
plotlist[[paste0(MUT,".abs")]]<-p1 #also store in list so it can be plotted in grid later
# Create a stacked barplot with propotions 
p2<-ggplot(data.m , aes(x = variable, y = value, fill = categories)) +
    geom_col(position="fill") +
    scale_fill_manual(values = c("CohesinAssEnhancers.CTCF" = "#a69f1a", "CohesinAssEnhancers.wo.CTCF" = "#ffee03",
    "wo.CohesinAssEnhancers.RAD21.CTCF" = "#9c2a85", "wo.CohesinAssEnhancers.RAD21.woCTCF"= "#dd23b8", "wo.CohesinAssEnhancers.wo.RAD21.CTCF"= "#1411dc",
    "wo.CohesinAssEnhancers.wo.RAD21.woCTCF.H3K27ac"= "#a4d833", "wo.CohesinAssEnhancers.wo.RAD21.woCTCF.woH3K27ac"= "#aaaaa1"),
    labels=c("Cohesin E/P + CTCF", "Cohesin E/P", "Cohesin + CTCF", "Cohesin", "CTCF", "E/P", "other")) + 
    xlab(" ") + ylab("proportion of loop anchors") +
    ggtitle(paste0("Differential loop anchors in ", MUT)) +
    theme_classic() +
    theme(axis.title=element_text(size=20,face="bold"),axis.text.x = element_text(size=15),axis.text.y = element_text(size=15),
    legend.text=element_text(size=15))
ggsave(file.path(FIGURESDIR,paste0(MUT,"diffloopAnchorFeatures.proportions.pdf")), plot=p2)
plotlist[[paste0(MUT,".perc")]]<-p2
}
```

## Combine STACKED BARPLOTS for STAG2-mut and RAD21-mut in grid
```{r}
library(cowplot)
pdf(file = file.path(file.path(FIGURESDIR,"SA2RAD21.diffloopAnchorFeatures.absolute.pdf")),height = 6,width = 14)
plot_grid(plotlist =list(plotlist[["SA2mut.abs"]],plotlist[["RAD21mut.abs"]]),ncol=2)
dev.off()
pdf(file = file.path(file.path(FIGURESDIR,"SA2RAD21.diffloopAnchorFeatures.proportions.pdf")),height = 6,width = 14)
plot_grid(plotlist =list(plotlist[["SA2mut.perc"]],plotlist[["RAD21mut.perc"]]),ncol=2)
dev.off()
```


#  Define loop Anchor pairs by H3K27ac overlap on both (="paired") or one (="unpaired") of the anchors
## read in  bed files into list
```{r}

loopstatsbed<-vector(mode = "list", length = 0)
for (MUT in MUTs) {
  for (FILT in Filts){
#loop anchors
path<-file.path(diffanchorsn2t,paste0(MUT,"vsCTRL.loopAnchors.red.", FILT, "sig.n2t.bed")) #set path
loopstatsbed[[paste0("totalDiffAnchors_",FILT,"_",MUT)]]<-read.table(path, header=F, sep="\t") #read in  the bedfiles as tab sep. table
loopstatsbed[[paste0("totalDiffAnchors_",FILT,"_",MUT)]]<-tidyr::separate_rows(loopstatsbed[[paste0("totalDiffAnchors_",FILT,"_",MUT)]],V4,sep=",",convert=FALSE) #make separate rows in case of same anchors used for different lopps
i<-c(1:nrow(loopstatsbed[[paste0("totalDiffAnchors_",FILT,"_",MUT)]]))
loopstatsbed[[paste0("totalDiffAnchors_",FILT,"_",MUT)]]$Loopname[i]<-sapply(loopstatsbed[[paste0("totalDiffAnchors_",FILT,"_",MUT)]]$V4[i],function(x) strsplit(as.character(x), "_")[[1]][1]) #create separate loopname column wihtout the _plus or _minus suffix
loopstatsbed[[paste0("totalDiffAnchors_",FILT,"_",MUT,"_plus")]]<-as.data.frame(loopstatsbed[[paste0("totalDiffAnchors_",FILT,"_",MUT)]])[as.data.frame(loopstatsbed[[paste0("totalDiffAnchors_",FILT,"_",MUT)]])$V6=="+",]#separate list of plus-anchors
loopstatsbed[[paste0("totalDiffAnchors_",FILT,"_",MUT,"_minus")]]<-as.data.frame(loopstatsbed[[paste0("totalDiffAnchors_",FILT,"_",MUT)]])[as.data.frame(loopstatsbed[[paste0("totalDiffAnchors_",FILT,"_",MUT)]])$V6=="-",]#separate list of minus-anchors
  }
}
#add AnchorChIPoverlaps
cats1=c("CTCFover","RAD21over","MED12over")
#loop anchors with ChIPseq peaks
for (MUT in MUTs) {
  for (FILT in Filts){
   for (cat in cats1){
path<-file.path(diffanchorsn2t,"AnchorChIPoverlap",paste0(MUT,"vsCTRL.loopAnchors.red.", FILT, ".n2t.",cat,".bed"))
loopstatsbed[[paste0(cat,"_",FILT,"_",MUT)]]<-read.table(path, header=F, sep="\t")
loopstatsbed[[paste0(cat,"_",FILT,"_",MUT)]]<-tidyr::separate_rows(loopstatsbed[[paste0(cat,"_",FILT,"_",MUT)]],V4,sep=",",convert=FALSE) #make separate rows in case of same anchors used for different lopps
i<-c(1:nrow(loopstatsbed[[paste0(cat,"_",FILT,"_",MUT)]]))
loopstatsbed[[paste0(cat,"_",FILT,"_",MUT)]]$Loopname[i]<-sapply(loopstatsbed[[paste0(cat,"_",FILT,"_",MUT)]]$V4[i],function(x) strsplit(as.character(x), "_")[[1]][1]) #create separate loopname column wihtout the _plus or _minus suffix
loopstatsbed[[paste0(cat,"_",FILT,"_",MUT,"_plus")]]<-as.data.frame(loopstatsbed[[paste0(cat,"_",FILT,"_",MUT)]])[as.data.frame(loopstatsbed[[paste0(cat,"_",FILT,"_",MUT)]])$V6=="+",]#separate list of plus-anchors
loopstatsbed[[paste0(cat,"_",FILT,"_",MUT,"_minus")]]<-as.data.frame(loopstatsbed[[paste0(cat,"_",FILT,"_",MUT)]])[as.data.frame(loopstatsbed[[paste0(cat,"_",FILT,"_",MUT)]])$V6=="-",]#separate list of minus-anchors
  }
}
}
#double overlaps
cats2=c("RAD21andCTCF","CTCFwoRAD21")
for (MUT in MUTs) {
  for (FILT in Filts){
   for (cat in cats2){
path<-file.path(diffanchorsn2t,"AnchorChIPoverlap",paste0(MUT,"vsCTRL.loopAnchors.red.", FILT,".",cat,".overlap.bed"))
loopstatsbed[[paste0(cat,"_",FILT,"_",MUT)]]<-read.table(path, header=F, sep="\t")
loopstatsbed[[paste0(cat,"_",FILT,"_",MUT)]]<-tidyr::separate_rows(loopstatsbed[[paste0(cat,"_",FILT,"_",MUT)]],V4,sep=",",convert=FALSE) #make separate rows in case of same anchors used for different lopps
i<-c(1:nrow(loopstatsbed[[paste0(cat,"_",FILT,"_",MUT)]]))
loopstatsbed[[paste0(cat,"_",FILT,"_",MUT)]]$Loopname[i]<-sapply(loopstatsbed[[paste0(cat,"_",FILT,"_",MUT)]]$V4[i],function(x) strsplit(as.character(x), "_")[[1]][1]) #create separate loopname column wihtout the _plus or _minus suffix
loopstatsbed[[paste0(cat,"_",FILT,"_",MUT,"_plus")]]<-as.data.frame(loopstatsbed[[paste0(cat,"_",FILT,"_",MUT)]])[as.data.frame(loopstatsbed[[paste0(cat,"_",FILT,"_",MUT)]])$V6=="+",]#separate list of plus-anchors
loopstatsbed[[paste0(cat,"_",FILT,"_",MUT,"_minus")]]<-as.data.frame(loopstatsbed[[paste0(cat,"_",FILT,"_",MUT)]])[as.data.frame(loopstatsbed[[paste0(cat,"_",FILT,"_",MUT)]])$V6=="-",]#separate list of minus-anchors
  }}}
#Enhancers at loop anchors
cats3=c("H3K27acOverlap","CohesinAssEnhancers.Overlap","CohesinAssEnhancersCTCF.overlap")
for (MUT in MUTs) {
  for (FILT in Filts){
   for (cat in cats3){
path<-file.path(diffanchorsn2t,"H3K27overlap",paste0(MUT,"vsCTRL.loopAnchors.red.", FILT,".",cat,".bed"))
loopstatsbed[[paste0(cat,"_",FILT,"_",MUT)]]<-read.table(path, header=F, sep="\t")
loopstatsbed[[paste0(cat,"_",FILT,"_",MUT)]]<-tidyr::separate_rows(loopstatsbed[[paste0(cat,"_",FILT,"_",MUT)]],V4,sep=",",convert=FALSE) #make separate rows in case of same anchors used for different lopps
i<-c(1:nrow(loopstatsbed[[paste0(cat,"_",FILT,"_",MUT)]]))
loopstatsbed[[paste0(cat,"_",FILT,"_",MUT)]]$Loopname[i]<-sapply(loopstatsbed[[paste0(cat,"_",FILT,"_",MUT)]]$V4[i],function(x) strsplit(as.character(x), "_")[[1]][1]) #create separate loopname column wihtout the _plus or _minus suffix
loopstatsbed[[paste0(cat,"_",FILT,"_",MUT,"_plus")]]<-as.data.frame(loopstatsbed[[paste0(cat,"_",FILT,"_",MUT)]])[as.data.frame(loopstatsbed[[paste0(cat,"_",FILT,"_",MUT)]])$V6=="+",]#separate list of plus-anchors
loopstatsbed[[paste0(cat,"_",FILT,"_",MUT,"_minus")]]<-as.data.frame(loopstatsbed[[paste0(cat,"_",FILT,"_",MUT)]])[as.data.frame(loopstatsbed[[paste0(cat,"_",FILT,"_",MUT)]])$V6=="-",]#separate list of minus-anchors
  }}}
#updown overlap ambigous anchors
for (MUT in MUTs) {
path<-file.path(diffanchorsn2t,paste0(MUT,"vsCTRL.loopAnchors.overlap.FC1.n2t.bed"))
loopstatsbed[[paste0("UpdownOverlap_",MUT)]]<-read.table(path, header=F, sep="\t")
loopstatsbed[[paste0("UpdownOverlap_",MUT)]]<-tidyr::separate_rows(loopstatsbed[[paste0("UpdownOverlap_",MUT)]],V4,sep=",",convert=FALSE)
i<-c(1:nrow(loopstatsbed[[paste0("UpdownOverlap_",MUT)]]))
loopstatsbed[[paste0("UpdownOverlap_",MUT)]]$Loopname[i]<-sapply(loopstatsbed[[paste0("UpdownOverlap_",MUT)]]$V4[i],function(x) strsplit(as.character(x), "_")[[1]][1]) #create separate loopname column wihtout the _plus or _minus suffix
loopstatsbed[[paste0("UpdownOverlap_",MUT,"_plus")]]<-as.data.frame(loopstatsbed[[paste0("UpdownOverlap_",MUT)]])[as.data.frame(loopstatsbed[[paste0("UpdownOverlap_",MUT)]])$V6=="+",]#separate list of plus-anchors
loopstatsbed[[paste0("UpdownOverlap_",MUT,"_minus")]]<-as.data.frame(loopstatsbed[[paste0("UpdownOverlap_",MUT)]])[as.data.frame(loopstatsbed[[paste0("UpdownOverlap_",MUT)]])$V6=="-",]#separate list of minus-anchors
}
```


# create new pairs of files (+/-) for paired loop anchors (for the different feature overlap categories)
```{r}
##dir.create(file.path(diffanchorsn2t,"AnchorsOverlapPairs"))
allcat=c("totalDiffAnchors",cats1,cats2,cats3)
for (cat in allcat) {
for (MUT in MUTs) {
  for (FILT in Filts){
plusAnchorspaired<-subset(loopstatsbed[[paste0(cat,"_",FILT,"_",MUT,"_plus")]],loopstatsbed[[paste0(cat,"_",FILT,"_",MUT,"_plus")]]$Loopname %in% loopstatsbed[[paste0(cat,"_",FILT,"_",MUT,"_minus")]]$Loopname)
minusAnchorspaired<-subset(loopstatsbed[[paste0(cat,"_",FILT,"_",MUT,"_minus")]],loopstatsbed[[paste0(cat,"_",FILT,"_",MUT,"_minus")]]$Loopname %in% loopstatsbed[[paste0(cat,"_",FILT,"_",MUT,"_plus")]]$Loopname)
bedplus <- plusAnchorspaired[,c('V1', 'V2', 'V3', 'Loopname','V5','V6')]
colnames(bedplus) <- c('chrom', 'chromStart', 'chromEnd', 'Loopname','score','strand')
write.table(bedplus, file.path(diffanchorsn2t,"AnchorsOverlapPairs",paste0(cat,"_",FILT,"_anchors_",MUT,"_paired_plus.bed")),row.names=F,quote=F,sep="\t",col.names=F)
bedminus <- minusAnchorspaired[,c('V1', 'V2', 'V3', 'Loopname','V5','V6')]
colnames(bedminus) <- c('chrom', 'chromStart', 'chromEnd', 'Loopname','score','strand')
write.table(bedminus, file.path(diffanchorsn2t,"AnchorsOverlapPairs",paste0(cat,"_",FILT,"_anchors_",MUT,"_paired_minus.bed")),row.names=F,quote=F,sep="\t",col.names=F)
}}}
```

# for H3K27ac anchors: also create pairs of files (+/-) for pairs where only one of the anchors has the enhancer/promoter mark
```{r}
cat<-"H3K27acOverlap"
for (MUT in MUTs) {
  for (FILT in Filts){
#for "unpaired" plus anchors with H3K27ac
unpaired_plus<-subset(loopstatsbed[[paste0(cat,"_",FILT,"_",MUT,"_plus")]],!(loopstatsbed[[paste0(cat,"_",FILT,"_",MUT,"_plus")]]$Loopname %in% loopstatsbed[[paste0(cat,"_",FILT,"_",MUT,"_minus")]]$Loopname)) #252
unpaired_plus_minus<-subset(loopstatsbed[[paste0("totalDiffAnchors_",FILT,"_",MUT,"_minus")]],loopstatsbed[[paste0("totalDiffAnchors_",FILT,"_",MUT,"_minus")]]$Loopname %in% unpaired_plus$Loopname) #238 --> some are missing?
unpaired_plus_minus_missing<-subset(unpaired_plus,!(unpaired_plus$Loopname %in% unpaired_plus_minus$Loopname)) #14 ##they are in the ambigous up-down fraction?
unpaired_plus_minus_missing_check<-subset(loopstatsbed[[paste0("UpdownOverlap_",MUT,"_minus")]], loopstatsbed[[paste0("UpdownOverlap_",MUT,"_minus")]]$Loopname %in%  unpaired_plus_minus_missing$Loopname) #14
unpaired_plus_minus_complete<-rbind(unpaired_plus_minus,unpaired_plus_minus_missing_check)###add the missing ones to have a complete list of partners
sorted_unpaired_plus_minus<-unpaired_plus_minus_complete[match(unpaired_plus$Loopname, unpaired_plus_minus_complete$Loopname),] #sort the complete minus anchors to order of unpaired plus
bedplus <- unpaired_plus[,c('V1', 'V2', 'V3', 'Loopname','V5','V6')]
write.table(bedplus, file.path(diffanchorsn2t,"AnchorsOverlapPairs",paste0(cat,"_",FILT,"_anchors_",MUT,"_unpairedPLUS_plus.bed")),row.names=F,quote=F,sep="\t",col.names=F)
bedminus <- sorted_unpaired_plus_minus[,c('V1', 'V2', 'V3', 'Loopname','V5','V6')]
write.table(bedminus, file.path(diffanchorsn2t,"AnchorsOverlapPairs",paste0(cat,"_",FILT,"_anchors_",MUT,"_unpairedPLUS_minus.bed")),row.names=F,quote=F,sep="\t",col.names=F)

#for "unpaired" minus anchors with H3K27ac
unpaired_minus<-subset(loopstatsbed[[paste0(cat,"_",FILT,"_",MUT,"_minus")]],!(loopstatsbed[[paste0(cat,"_",FILT,"_",MUT,"_minus")]]$Loopname %in% loopstatsbed[[paste0(cat,"_",FILT,"_",MUT,"_plus")]]$Loopname)) #248
unpaired_minus_plus<-subset(loopstatsbed[[paste0("totalDiffAnchors_",FILT,"_",MUT,"_plus")]],loopstatsbed[[paste0("totalDiffAnchors_",FILT,"_",MUT,"_plus")]]$Loopname %in% unpaired_minus$Loopname) #226 --> some are missing?
unpaired_minus_plus_missing<-subset(unpaired_minus,!(unpaired_minus$Loopname %in% unpaired_minus_plus$Loopname)) #22
unpaired_minus_plus_missing_check<-subset(loopstatsbed[[paste0("UpdownOverlap_",MUT,"_plus")]], loopstatsbed[[paste0("UpdownOverlap_",MUT,"_plus")]]$Loopname %in%  unpaired_minus_plus_missing$Loopname) #22
unpaired_minus_plus_complete<-rbind(unpaired_minus_plus,unpaired_minus_plus_missing_check)###add the missing ones to have a complete list of partners
sorted_unpaired_minus_plus<-unpaired_minus_plus_complete[match(unpaired_minus$Loopname, unpaired_minus_plus_complete$Loopname),] #sort the complete minus anchors to order of unpaired plus
bedplus <- sorted_unpaired_minus_plus[,c('V1', 'V2', 'V3', 'Loopname','V5','V6')]
write.table(bedplus, file.path(diffanchorsn2t,"AnchorsOverlapPairs",paste0(cat,"_",FILT,"_anchors_",MUT,"_unpairedMINUS_plus.bed")),row.names=F,quote=F,sep="\t",col.names=F)
bedminus <- unpaired_minus[,c('V1', 'V2', 'V3', 'Loopname','V5','V6')]
write.table(bedminus, file.path(diffanchorsn2t,"AnchorsOverlapPairs",paste0(cat,"_",FILT,"_anchors_",MUT,"_unpairedMINUS_minus.bed")),row.names=F,quote=F,sep="\t",col.names=F)
}}
```


# Statistics for paired vs unpaired loop anchors
```{r}
#number of paired and unpaired loopanchors by set in dataframe for stats to plot

pairedloopstats<-data.frame()
for (cat in allcat) {
for (MUT in MUTs) {
  for (FILT in Filts){
pairedloopstats[paste0(cat,"_",FILT),paste0(MUT,".paired")]<-nrow(subset(loopstatsbed[[paste0(cat,"_",FILT,"_",MUT,"_plus")]],loopstatsbed[[paste0(cat,"_",FILT,"_",MUT,"_plus")]]$Loopname %in% loopstatsbed[[paste0(cat,"_",FILT,"_",MUT,"_minus")]]$Loopname))
pairedloopstats[paste0(cat,"_",FILT),paste0(MUT,".unpaired")]<-nrow(subset(loopstatsbed[[paste0(cat,"_",FILT,"_",MUT,"_plus")]],!(loopstatsbed[[paste0(cat,"_",FILT,"_",MUT,"_plus")]]$Loopname %in% loopstatsbed[[paste0(cat,"_",FILT,"_",MUT,"_minus")]]$Loopname)))
}}}
#transform dataframe and relevel to desired order in plot
data.m2 <- melt(as.matrix(pairedloopstats), id.vars=rownames(pairedloopstats))
colnames(data.m2)<-c("feature","conditiontype","value")
i<-c(1:nrow(data.m2))
data.m2$group[i]<-sapply(data.m2$conditiontype[i],function(x) strsplit(as.character(x), "\\.")[[1]][1])
data.m2$type[i]<-sapply(data.m2$conditiontype[i],function(x) strsplit(as.character(x), "\\.")[[1]][2])
data.m2$direction[i]<-sapply(data.m2$feature[i],function(x) strsplit(as.character(x), "_")[[1]][2])
data.m2$group <- factor(data.m2$group , levels = c("SA2mut","RAD21mut"))
data.m2$conditiontype <- factor(data.m2$conditiontype , levels = c("SA2mut.paired","SA2mut.unpaired","RAD21mut.paired","RAD21mut.unpaired"))
data.m2$feature <- factor(data.m2$feature , levels = c("totalDiffAnchors_up.FC1","CTCFover_up.FC1","RAD21over_up.FC1","RAD21andCTCF_up.FC1","CTCFwoRAD21_up.FC1","MED12over_up.FC1","H3K27acOverlap_up.FC1","CohesinAssEnhancers.Overlap_up.FC1","CohesinAssEnhancersCTCF.overlap_up.FC1","totalDiffAnchors_down.FC1","CTCFover_down.FC1","RAD21over_down.FC1","RAD21andCTCF_down.FC1","CTCFwoRAD21_down.FC1","MED12over_down.FC1","H3K27acOverlap_down.FC1","CohesinAssEnhancers.Overlap_down.FC1","CohesinAssEnhancersCTCF.overlap_down.FC1"))


labs<-c("totalDiffAnchors","CTCF","RAD21","RAD21+CTCF","CTCFwoRAD21","MED12","H3K27ac","H3K27ac+RAD21","H3K27ac+RAD21+CTCF")

#create stacked (paired+unpaired) barplot grouped by up/down and Mutation group (using facets)

pdf(file=file.path(diffanchorsn2t,paste0("CohAML",".paired.Diffloop.ChIP.OverlapSummary.pdf")))
ggplot(data.m2) +
  geom_bar(aes(feature,value,fill = conditiontype), position = "stack", stat="identity") +
    facet_wrap(~group+direction,drop=T,nrow=2,scales="free_x") + theme(panel.spacing = unit(1, "lines"))+ 
    scale_fill_manual(values = c("SA2mut.paired" = "seagreen3","SA2mut.unpaired" = "lightgreen", "RAD21mut.paired" = "mediumvioletred","RAD21mut.unpaired" = "violet")) +
    scale_x_discrete(labels=labs) +
    xlab("feature") + ylab("frequency") +
    ggtitle("paired diff. Loop anchor associated features") +
    theme(
  axis.text.x = element_text(angle=90,hjust=0.95,vjust=0.2),
  axis.title=element_text(size=20,face="bold"), plot.title = element_text(size = 12, face = "bold"), 
  panel.border = element_blank(),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.background = element_blank(),
  axis.line = element_line(colour = "black"),)
dev.off() 
```


