--
title: "RAD21 diffpeak motif enrichment plots - HPSCs"
author: "Alexander Fischer"
date: "04 2023"
output: html
---

# set directories and define function to read motif results
```{r}
#libraries
library(XML)
library(ggplot2)
library(data.table)
library(cowplot)
library(grImport2)
library(grConvert)
library(grid)

#paths
PROJDIR<-"/misc/data/analysis/project_cohesin/"
MOTIFDIR<-file.path(PROJDIR,"CD34/ChIP_KD_analysis/Cohesin_CTCF_MED12/motifs/RAD21revised")


FIGDIR<-file.path(PROJDIR,"CD34/ChIP_KD_analysis/Cohesin_CTCF_MED12/figures/RAD21revised")
dir.create(file.path(FIGDIR,"TOPmotifBarplots"))


#variables
cats<-c("RAD21.Peaks_edgeR")
filts<-c(".2foldDown",".2foldUp")
cats2<-c("RAD21.Peaks_edgeR.model.")
filts2<-c("2folddown.","2foldup.")

```

# read homer html results into list
```{r}
STAGKD<-c("SA2KD","SA1KD")
#read homer html results into list
Motiflist<-vector(mode = "list", length = 0)
for (KD in STAGKD){
for (filt in filts){
  for (type in cats){   
file <-file.path(MOTIFDIR,paste0(KD,"vsCTRL.",type,filt,"/final/homerResults.html"))
info <- file.info(file)
if (is.na(info$size) == FALSE) { #check if file if empty 
html<-readHTMLTable(file)
htmldf<-as.data.frame(html)
colnames(htmldf)<-c("RANK","MOTIF","Pvalue","LogPvalue","percententTargets","percentBackground","STD","Match","motifFile")
#calculate enrichment
htmldf$logEnrichment<-log2((as.numeric(sub("%", "", htmldf$percententTargets))/100)/(as.numeric(sub("%", "", htmldf$percentBackground))/100))
htmldf$TARvsBG<-(as.numeric(sub("%", "", htmldf$percententTargets))/100)/(as.numeric(sub("%", "", htmldf$percentBackground))/100)
#set levels for RANK
htmldf$RANK<-factor(htmldf$RANK,levels=as.numeric(htmldf$RANK))
#extract motifname/Family to display
i<-c(1:nrow(htmldf))
htmldf$NameFamily[i]<-sapply(htmldf$Match[i],function(x) strsplit(as.character(x), "/")[[1]][1]) 
htmldf$MotifFamily[i]<-sapply(htmldf$NameFamily[i],function(x) strsplit(as.character(x), "[()]")[[1]][2]) 
htmldf$MotifName[i]<-sapply(htmldf$Match[i],function(x) strsplit(as.character(x), "[()]")[[1]][1]) 
htmldf$MotifName[i]<-sapply(htmldf$MotifName[i],function(x) strsplit(as.character(x), "/")[[1]][1]) 
#individual corrections
htmldf[which(htmldf$MotifName=="PB0058.1_Sfpi1_1"),"MotifName"]<-"PU.1"
htmldf[which(htmldf$MotifName=="PU.1"),"MotifFamily"]<-"ETS"
htmldf[which(htmldf$MotifName=="PU.1"),"NameFamily"]<-"PU.1(ETS)"
htmldf[which(htmldf$MotifName=="SPIC"),"MotifFamily"]<-"ETS"
htmldf[which(htmldf$MotifName=="IKZF1"),"MotifFamily"]<-"IKAROS Zf"
htmldf[which(htmldf$MotifName=="CTCF"|htmldf$MotifName=="CTCFL"|htmldf$MotifName=="BORIS"),"MotifFamily"]<-"CTCF"
htmldf[htmldf$MotifName %like% "BHLH","MotifFamily" ]<-"bHLH"
htmldf[htmldf$MotifName %like% "Gata","MotifFamily" ]<-"GATA"
htmldf[htmldf$MotifName %like% "RAP2","MotifFamily" ]<-"ERF"
htmldf[htmldf$MotifName %like% "ELF","MotifFamily" ]<-"ETS"
htmldf[htmldf$MotifName %like% "CEBP","MotifFamily" ]<-"CEBP"
#add to list
Motiflist[[paste0(KD,"vsCTRL.",type,filt)]]<-htmldf
}
}}}
```



# function for plotting
```{r}
MotifBarFunc<-function(htmldf,Top=5, xmax=NA,forcepvalcoord=NA,fillcol="seagreen",motifname="MotifName",sizeName=12){
#check if enough entries for Top, if not use as much as available
if (nrow(htmldf)< Top){
plotdf<-htmldf[1:nrow(htmldf),]
message(paste0("only ", nrow(htmldf), " entries"))
}else{
plotdf<-htmldf[1:Top,]
}
plotdf$RANK<-factor(plotdf$RANK,levels=c(Top:1)) ##rank needs to be inverted for this..!
maxval<-max(plotdf$logEnrichment) #cooridinate/position to display pvalue
if (is.na(forcepvalcoord)){
pvalcoord<-maxval+0.1*maxval}
else
{pvalcoord<-forcepvalcoord}
ggplot(plotdf) +
  geom_bar(aes(as.numeric(logEnrichment),RANK,fill = "filly",alpha=0.8), stat="identity") +
  scale_fill_manual(values = c("filly"=fillcol))+
    xlab("log2 enrichment") + ylab("") +
    ggtitle("") +
    geom_text(data=plotdf,aes(x=0,y=RANK,label=get(motifname)),size=sizeName,hjust=0,,fontface = "bold") +
    geom_text(data=plotdf,aes(x=pvalcoord,y=RANK,label=paste0("P < ",Pvalue)),size=sizeName*0.75,fontface = "italic") +
    coord_cartesian(xlim = c(0,xmax))+
  scale_x_continuous(expand = expansion(mult = c(0, 0))) + 
  theme(
  legend.position="none",
  axis.text.x = element_text(angle=0,hjust=0.95,vjust=0.2,size=20,face="bold"),
  axis.line.x = element_line(colour = "black"),
  axis.text.y = element_blank(), axis.line.y= element_blank(),axis.ticks.y= element_blank(),
  axis.title.x=element_text(size=20,face="bold"), 
  plot.title = element_text(size = 12, face = "bold"), 
  panel.border = element_blank(),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.background = element_blank(),
)
}
```


# integrate motifs images
## read in images (top 10 for all types)
```{r}
#import of svg files
##all diff peaks SA2KD
motifgroblist<-list()
for (filt in filts){
  for (type in cats){   
imagedir<-file.path(MOTIFDIR,paste0("SA2KDvsCTRL.",type,filt,"/motifimages"))
dir.create(imagedir)
for (i in c(1:10)){
convertPicture(file.path(MOTIFDIR,paste0("SA2KDvsCTRL.",type,filt,"/final/homerResults"),paste0("motif",i,".logo.svg")),file.path(imagedir,paste0("motif",i,".logo.svg")))
motifpic<-readPicture(file.path(imagedir,paste0("motif",i,".logo.svg")))
motifgroblist[[paste0("SA2KDvsCTRL.",type,filt,"motif",i)]]<-gTree(children=gList(pictureGrob(motifpic)))
  }  
}}
##all diff peaks SA1KD only "up"
imagedir<-file.path(MOTIFDIR,paste0("SA1KDvsCTRL.",type,".2foldUp","/motifimages"))
dir.create(imagedir)
for (i in c(1:2)){
convertPicture(file.path(MOTIFDIR,paste0("SA1KDvsCTRL.",type,".2foldUp","/final/homerResults"),paste0("motif",i,".logo.svg")),file.path(imagedir,paste0("motif",i,".logo.svg")))
motifpic<-readPicture(file.path(imagedir,paste0("motif",i,".logo.svg")))
motifgroblist[[paste0("SA1KDvsCTRL.",type,".2foldUp","motif",i)]]<-gTree(children=gList(pictureGrob(motifpic)))
  }  
```



## generate top 5 motifplots including svg logo for all-diff.peak sets
```{r}
##all in one
plotlistall<-list()
for (filt in filts){
  for (type in cats){  
p1<-MotifBarFunc(Motiflist[[paste0("SA2KDvsCTRL.",type,filt)]],Top=5,xmax=6,forcepvalcoord=4,fillcol="grey",motifname="MotifName",sizeName=6)
plotlistall[[paste0(filt,type)]]<-p1+ 
annotation_custom(motifgroblist[[paste0("SA2KDvsCTRL.",type,filt,"motif1")]],xmin = -6.5, xmax = 0.5, ymin = 4.5, ymax = 5.5)+
annotation_custom(motifgroblist[[paste0("SA2KDvsCTRL.",type,filt,"motif2")]],xmin = -6.5, xmax = 1.2, ymin = 3.5, ymax = 4.5)+
annotation_custom(motifgroblist[[paste0("SA2KDvsCTRL.",type,filt,"motif3")]],xmin = -6.5, xmax = 1.2, ymin = 2.5, ymax = 3.5)+
annotation_custom(motifgroblist[[paste0("SA2KDvsCTRL.",type,filt,"motif4")]],xmin = -6.5, xmax = 1.2, ymin = 1.5, ymax = 2.5)+
annotation_custom(motifgroblist[[paste0("SA2KDvsCTRL.",type,filt,"motif5")]],xmin = -6.5, xmax = 1.2, ymin = 0.5, ymax = 1.5)+
coord_cartesian(xlim = c(-6.5,6.5),clip="off")+
theme(plot.margin = unit(c(1,2,2,1), "cm"),axis.title.x=element_text(size=20,face="bold",hjust=0.8))
}}

pdf(file=file.path(FIGDIR,"TOPmotifBarplots","Top5motifs_SA2KD.RAD21diffpeaks.svgs.pdf"),height=8,width=7)
plot_grid(plotlist=plotlistall,ncol=1,labels=names(plotlistall))
dev.off()
```

## final motifplots filtered for motifs that are redundant
```{r}
type= "RAD21.Peaks_edgeR"
plotlistsel<-list()
###SA2 KD
#up top1 skipping redundant CTCF motifs
filt=".2foldUp"
p1<-MotifBarFunc(Motiflist[[paste0("SA2KDvsCTRL.",type,filt)]],Top=1,xmax=6,forcepvalcoord=6,fillcol="seagreen",motifname="NameFamily",sizeName=6)
plotlistsel[[paste0(filt,type)]]<-p1+ 
annotation_custom(motifgroblist[[paste0("SA2KDvsCTRL.",type,filt,"motif1")]],xmin = -6.5, xmax = 0.5, ymin = 0.5, ymax = 1.5)+
coord_cartesian(xlim = c(-6.5,6.5),clip="off")+
theme(plot.margin = unit(c(1,2,2,1), "cm"),axis.title.x=element_text(size=20,face="bold",hjust=0.8))

#down with CTCF top3 skipping redundant CTCF motifs
filt=".2foldDown"
selectmotifs<-Motiflist[[paste0("SA2KDvsCTRL.",type,filt)]][c(1:3,6:8),]
selectmotifs$RANK<-factor(c(1,2,3,4,5,6)) #important: redifine Ranks for plotting levels!
p1<-MotifBarFunc(selectmotifs,Top=6,xmax=6,forcepvalcoord=6,fillcol="firebrick",motifname="NameFamily",sizeName=6)
plotlistsel[[paste0(filt,type)]]<-p1+ 
annotation_custom(motifgroblist[[paste0("SA2KDvsCTRL.",type,filt,"motif1")]],xmin = -6.5, xmax = 0.5, ymin = 5.5, ymax = 6.5)+
annotation_custom(motifgroblist[[paste0("SA2KDvsCTRL.",type,filt,"motif2")]],xmin = -6.5, xmax = 0.5, ymin = 4.5, ymax = 5.5)+
annotation_custom(motifgroblist[[paste0("SA2KDvsCTRL.",type,filt,"motif3")]],xmin = -6.5, xmax = 0.5, ymin = 3.5, ymax = 4.5)+
annotation_custom(motifgroblist[[paste0("SA2KDvsCTRL.",type,filt,"motif6")]],xmin = -6.5, xmax = 0.5, ymin = 2.5, ymax = 3.5)+
annotation_custom(motifgroblist[[paste0("SA2KDvsCTRL.",type,filt,"motif7")]],xmin = -6.5, xmax = 0.5, ymin = 1.5, ymax = 2.5)+
annotation_custom(motifgroblist[[paste0("SA2KDvsCTRL.",type,filt,"motif8")]],xmin = -6.5, xmax = 0.5, ymin = 0.5, ymax = 1.5)+
coord_cartesian(xlim = c(-6.5,6.5),clip="off")+
theme(plot.margin = unit(c(1,2,2,1), "cm"),axis.title.x=element_text(size=20,face="bold",hjust=0.8))


#plot
pdf(file=file.path(FIGDIR,"TOPmotifBarplots","Top_sel.motifs_SA2KD.RAD21diffpeaks.svgs.pdf"),height=6,width=7)
plot_grid(plotlist=plotlistsel,ncol=1,labels=names(plotlistsel),rel_heights=c(3.5,5.5))
dev.off()

#save rds of plotlist
saveRDS(plotlistsel,file=file.path(FIGDIR,"TOPmotifBarplots","plotlistsel.HSPC.RAD21diff.motifbarplts.rdata"))

###SA1 KD
#up top1 skipping redundant CTCF motifs
filt=".2foldUp"
selectmotifs<-Motiflist[[paste0("SA1KDvsCTRL.",type,filt)]][c(2),]
selectmotifs$RANK<-factor(c(1)) #important: redifine Ranks for plotting levels!
p1<-MotifBarFunc(Motiflist[[paste0("SA1KDvsCTRL.",type,filt)]],Top=1,xmax=6,forcepvalcoord=6,fillcol="darkgoldenrod3",motifname="NameFamily",sizeName=6)
SA1KD_UP_motifs_plot<-p1+ 
annotation_custom(motifgroblist[[paste0("SA1KDvsCTRL.",type,filt,"motif2")]],xmin = -6.5, xmax = 0.5, ymin = 0.5, ymax = 1.5)+
coord_cartesian(xlim = c(-6.5,6.5),clip="off")+
theme(plot.margin = unit(c(1,2,2,1), "cm"),axis.title.x=element_text(size=20,face="bold",hjust=0.8))

plotlistsel2<-plotlistsel
plotlistsel2[["SA1KD_2folUP"]]<-SA1KD_UP_motifs_plot

pdf(file=file.path(FIGDIR,"TOPmotifBarplots","Top_sel.motifs_SA2KD_SA1KD.RAD21diffpeaks.svgs.pdf"),height=9,width=7)
plot_grid(plotlist=plotlistsel2,ncol=1,labels=names(plotlistsel),rel_heights=c(4,7,4))
dev.off()
```