--
title: "RAD21 diffpeak motif enrichment plots - AML"
author: "Alexander Fischer"
date: "04 2023"
output: html
---

# load libraries and set paths
```{r}
#libraries
library(XML)
library(ggplot2)
library(data.table)
library(cowplot)
library(grImport2)
library(grConvert)
library(grid)

#paths
PROJDIR<-"/misc/data/analysis/project_cohesin"
MOTIFDIR<-file.path(PROJDIR,"Cohesin_AML/ChIP_analysis/RAD21revised/motifs")
MOTIFDIRbyCTCF<-file.path(MOTIFDIR,"byCTCFoverlap")

FIGDIR<-file.path(PROJDIR,"Cohesin_AML/ChIP_analysis/RAD21revised/figures")
FIGDIRCD34<-file.path(PROJDIR,"CD34/ChIP_KD_analysis/Cohesin_CTCF_MED12/figures/RAD21revised")
dir.create(file.path(FIGDIR,"TOPmotifBarplots"))
dir.create(file.path(FIGDIR,"TOPmotifBarplots","byCTCFoverlap"))

#variables
CTCFcats=c("withCTCF", "noCTCF")
filts<-c("2folddown","2foldup")
MUTS<-c("SA2mut","RAD21mut")
```



# read homer html results into list
```{r}
###for CTCF split results
Motiflist<-vector(mode = "list", length = 0)
###for CTCF split results: only applies to SA2mut!
for (filt in filts){
type= "stringentPeaks_DESEQ.model."
      for (CTCF in CTCFcats){  
file <-file.path(MOTIFDIRbyCTCF,paste0("SA2mutvsCTRL.RAD21.",type,filt,".",CTCF,"/final/homerResults.html"))
html<-readHTMLTable(file)
htmldf<-as.data.frame(html)
colnames(htmldf)<-c("RANK","MOTIF","Pvalue","LogPvalue","percententTargets","percentBackground","STD","Match","motifFile")
#calculate enrichment
htmldf$logEnrichment<-log2((as.numeric(sub("%", "", htmldf$percententTargets))/100)/(as.numeric(sub("%", "", htmldf$percentBackground))/100))
htmldf$TARvsBG<-(as.numeric(sub("%", "", htmldf$percententTargets))/100)/(as.numeric(sub("%", "", htmldf$percentBackground))/100)
#set levels for RANK
htmldf$RANK<-factor(htmldf$RANK,levels=as.numeric(htmldf$RANK))
#extract motifname/Family to display
i<-c(1:nrow(htmldf))
htmldf$NameFamily[i]<-sapply(htmldf$Match[i],function(x) strsplit(as.character(x), "/")[[1]][1]) 
htmldf$MotifFamily[i]<-sapply(htmldf$NameFamily[i],function(x) strsplit(as.character(x), "[()]")[[1]][2]) 
htmldf$MotifName[i]<-sapply(htmldf$Match[i],function(x) strsplit(as.character(x), "[()]")[[1]][1]) 
htmldf$MotifName[i]<-sapply(htmldf$MotifName[i],function(x) strsplit(as.character(x), "/")[[1]][1]) 
#individual corrections
htmldf[which(htmldf$MotifName=="PB0058.1_Sfpi1_1"),"MotifName"]<-"PU.1"
htmldf[which(htmldf$MotifName=="PU.1"),"MotifFamily"]<-"ETS"
htmldf[which(htmldf$MotifName=="PU.1"),"NameFamily"]<-"PU.1(ETS)"
htmldf[which(htmldf$MotifName=="SPIC"),"MotifFamily"]<-"ETS"
htmldf[which(htmldf$MotifName=="IKZF1"),"MotifFamily"]<-"IKAROS Zf"
htmldf[which(htmldf$MotifName=="CTCF"|htmldf$MotifName=="CTCFL"|htmldf$MotifName=="BORIS"),"MotifFamily"]<-"CTCF"
htmldf[htmldf$MotifName %like% "BHLH","MotifFamily" ]<-"bHLH"
htmldf[htmldf$MotifName %like% "Gata","MotifFamily" ]<-"GATA"
htmldf[htmldf$MotifName %like% "RAP2","MotifFamily" ]<-"ERF"
#add to list
Motiflist[[paste0("SA2mutvsCTRL.RAD21.",type,filt,CTCF)]]<-htmldf
}}
```


# function for plotting 
```{r}
#function for plotting 
MotifBarFunc<-function(htmldf,Top=5, xmax=NA,forcepvalcoord=NA,fillcol="seagreen",motifname="MotifName",sizeName=12){
plotdf<-htmldf[1:Top,]
plotdf$RANK<-factor(plotdf$RANK,levels=c(Top:1)) ##rank needs to be inverted for this..!
maxval<-max(plotdf$logEnrichment) #cooridinate/position to display pvalue
if (is.na(forcepvalcoord)){
pvalcoord<-maxval+0.1*maxval}
else
{pvalcoord<-forcepvalcoord}
ggplot(plotdf) +
  geom_bar(aes(as.numeric(logEnrichment),RANK,fill = "filly",alpha=0.8), stat="identity") +
  scale_fill_manual(values = c("filly"=fillcol))+
    xlab("log2 enrichment") + ylab("") +
    ggtitle("") +
    geom_text(data=plotdf,aes(x=0,y=RANK,label=get(motifname)),size=sizeName,hjust=0,,fontface = "bold") +
    geom_text(data=plotdf,aes(x=pvalcoord,y=RANK,label=paste0("P < ",Pvalue)),size=sizeName*0.75,fontface = "italic") +
    coord_cartesian(xlim = c(0,xmax))+
  scale_x_continuous(expand = expansion(mult = c(0, 0))) + 
  theme(
  legend.position="none",
  axis.text.x = element_text(angle=0,hjust=0.95,vjust=0.2,size=20,face="bold"),
  axis.line.x = element_line(colour = "black"),
  axis.text.y = element_blank(), axis.line.y= element_blank(),axis.ticks.y= element_blank(),
  axis.title.x=element_text(size=20,face="bold"), 
  plot.title = element_text(size = 12, face = "bold"), 
  panel.border = element_blank(),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.background = element_blank(),
)
}
```

## test plots
```{r}
###look at all plots by CTCF subset
      for (CTCF in CTCFcats){
        type= "stringentPeaks_DESEQ.model."
pdf(file=file.path(FIGDIR,"TOPmotifBarplots","byCTCFoverlap",paste0("Top6.SA2mutvsCTRL.RAD21.",type,"2foldup.",CTCF,".pdf")))
plot(MotifBarFunc(Motiflist[[paste0("SA2mutvsCTRL.RAD21.",type,"2foldup.",CTCF)]],Top=6,xmax=6.5,forcepvalcoord=5.7,fillcol="seagreen",motifname="NameFamily"))
dev.off()
pdf(file=file.path(FIGDIR,"TOPmotifBarplots","byCTCFoverlap",paste0("Top6.SA2mutvsCTRL.RAD21.",type,"2folddown.",CTCF,".pdf")))
plot(MotifBarFunc(Motiflist[[paste0("SA2mutvsCTRL.RAD21.",type,"2folddown.",CTCF)]],Top=6,xmax=6.5,forcepvalcoord=5.7,fillcol="firebrick",motifname="NameFamily"))
dev.off()
}

```

# integrate motifs images
## read in images (top 10 for all types)
```{r}
##diffpeaks by CTCF subset
motifgroblistbyCTCF<-list()
for (filt in filts){
#  for (type in cats){
type= "stringentPeaks_DESEQ.model."
for (CTCF in CTCFcats){  
imagedir<-file.path(MOTIFDIRbyCTCF,paste0("SA2mutvsCTRL.RAD21.",type,filt,".",CTCF,"/motifimages"))
dir.create(imagedir)
 for (i in c(1:10)){
convertPicture(file.path(MOTIFDIRbyCTCF,paste0("SA2mutvsCTRL.RAD21.",type,filt,".",CTCF,"/final/homerResults"),paste0("motif",i,".logo.svg")),file.path(imagedir,paste0("motif",i,".logo.svg")))
motifpic<-readPicture(file.path(imagedir,paste0("motif",i,".logo.svg")))
motifgroblistbyCTCF[[paste0("SA2mutvsCTRL.RAD21.",type,filt,CTCF,"motif",i)]]<-gTree(children=gList(pictureGrob(motifpic)))
}
}}
```

## generate top 5 motifplots including svg logo for all-diff.peak sets
```{r}
##all in one STAG2 mut ##by CTCF category
plotlistbyCTCF<-list()
for (filt in filts){
    for (CTCF in CTCFcats){
#  for (type in cats){
type= "stringentPeaks_DESEQ.model."
p1<-MotifBarFunc(Motiflist[[paste0("SA2mutvsCTRL.RAD21.",type,filt,CTCF)]],Top=5,xmax=6,forcepvalcoord=5,fillcol="grey",motifname="MotifName",sizeName=6)
plotlistbyCTCF[[paste0(filt,type,CTCF)]]<-p1+ 
annotation_custom(motifgroblistbyCTCF[[paste0("SA2mutvsCTRL.RAD21.",type,filt,CTCF,"motif1")]],xmin = -6.5, xmax = 0.5, ymin = 4.5, ymax = 5.5)+
annotation_custom(motifgroblistbyCTCF[[paste0("SA2mutvsCTRL.RAD21.",type,filt,CTCF,"motif2")]],xmin = -6.5, xmax = 1.2, ymin = 3.5, ymax = 4.5)+
annotation_custom(motifgroblistbyCTCF[[paste0("SA2mutvsCTRL.RAD21.",type,filt,CTCF,"motif3")]],xmin = -6.5, xmax = 1.2, ymin = 2.5, ymax = 3.5)+
annotation_custom(motifgroblistbyCTCF[[paste0("SA2mutvsCTRL.RAD21.",type,filt,CTCF,"motif4")]],xmin = -6.5, xmax = 1.2, ymin = 1.5, ymax = 2.5)+
annotation_custom(motifgroblistbyCTCF[[paste0("SA2mutvsCTRL.RAD21.",type,filt,CTCF,"motif5")]],xmin = -6.5, xmax = 1.2, ymin = 0.5, ymax = 1.5)+
coord_cartesian(xlim = c(-6.5,6.5),clip="off")+
theme(plot.margin = unit(c(1,2,2,1), "cm"),axis.title.x=element_text(size=20,face="bold",hjust=0.8))
}}


pdf(file=file.path(FIGDIR,"TOPmotifBarplots","byCTCFoverlap","Top5motifs_SA2mut.stringentRAD21diffpeaks.by.CTCF.svgs.pdf"),height=16,width=7)
plot_grid(plotlist=plotlistbyCTCF,ncol=1,labels=names(plotlistbyCTCF))
dev.off()
```

## final motifplots by CTCF individually filtered for motifs that are redundant
```{r}
type= "stringentPeaks_DESEQ.model."
plotlistbyCTCFsel<-list()
#up with CTCF top3 skipping redundant CTCF motifs
filt="2foldup"
CTCF="withCTCF"
selectmotifs<-Motiflist[[paste0("SA2mutvsCTRL.RAD21.",type,filt,CTCF)]][c(1,2,5),]
selectmotifs$RANK<-factor(c(1,2,3)) #important: redifine Ranks for plotting levels!
p1<-MotifBarFunc(selectmotifs,Top=3,xmax=6,forcepvalcoord=6,fillcol="seagreen",motifname="NameFamily",sizeName=6)
plotlistbyCTCFsel[[paste0(filt,type,CTCF)]]<-p1+ 
annotation_custom(motifgroblistbyCTCF[[paste0("SA2mutvsCTRL.RAD21.",type,filt,CTCF,"motif1")]],xmin = -6.5, xmax = 0.5, ymin = 2.5, ymax = 3.5)+
annotation_custom(motifgroblistbyCTCF[[paste0("SA2mutvsCTRL.RAD21.",type,filt,CTCF,"motif2")]],xmin = -6.5, xmax = 1.2, ymin = 1.5, ymax = 2.5)+
annotation_custom(motifgroblistbyCTCF[[paste0("SA2mutvsCTRL.RAD21.",type,filt,CTCF,"motif5")]],xmin = -6.5, xmax = 1.2, ymin = 0.5, ymax = 1.5)+
coord_cartesian(xlim = c(-6.5,6.5),clip="off")+
theme(plot.margin = unit(c(1,2,2,1), "cm"),axis.title.x=element_text(size=20,face="bold",hjust=0.8))
#up no CTCF top3
CTCF="noCTCF"
p1<-MotifBarFunc(Motiflist[[paste0("SA2mutvsCTRL.RAD21.",type,filt,CTCF)]],Top=3,xmax=6,forcepvalcoord=6,fillcol="seagreen",motifname="NameFamily",sizeName=6)
plotlistbyCTCFsel[[paste0(filt,type,CTCF)]]<-p1+ 
annotation_custom(motifgroblistbyCTCF[[paste0("SA2mutvsCTRL.RAD21.",type,filt,CTCF,"motif1")]],xmin = -6.5, xmax = 0.5, ymin = 2.5, ymax = 3.5)+
annotation_custom(motifgroblistbyCTCF[[paste0("SA2mutvsCTRL.RAD21.",type,filt,CTCF,"motif2")]],xmin = -6.5, xmax = 1.2, ymin = 1.5, ymax = 2.5)+
annotation_custom(motifgroblistbyCTCF[[paste0("SA2mutvsCTRL.RAD21.",type,filt,CTCF,"motif3")]],xmin = -6.5, xmax = 1.2, ymin = 0.5, ymax = 1.5)+
coord_cartesian(xlim = c(-6.5,6.5),clip="off")+
theme(plot.margin = unit(c(1,2,2,1), "cm"),axis.title.x=element_text(size=20,face="bold",hjust=0.8))


#down with CTCF top3 skipping redundant CTCF motifs
filt="2folddown"
CTCF="withCTCF"
selectmotifs<-Motiflist[[paste0("SA2mutvsCTRL.RAD21.",type,filt,CTCF)]][c(1,2,5),]
selectmotifs$RANK<-factor(c(1,2,3)) #important: redifine Ranks for plotting levels!
p1<-MotifBarFunc(selectmotifs,Top=3,xmax=6,forcepvalcoord=6,fillcol="firebrick",motifname="NameFamily",sizeName=6)
plotlistbyCTCFsel[[paste0(filt,type,CTCF)]]<-p1+ 
annotation_custom(motifgroblistbyCTCF[[paste0("SA2mutvsCTRL.RAD21.",type,filt,CTCF,"motif1")]],xmin = -6.5, xmax = 0.5, ymin = 2.5, ymax = 3.5)+
annotation_custom(motifgroblistbyCTCF[[paste0("SA2mutvsCTRL.RAD21.",type,filt,CTCF,"motif2")]],xmin = -6.5, xmax = 1.2, ymin = 1.5, ymax = 2.5)+
annotation_custom(motifgroblistbyCTCF[[paste0("SA2mutvsCTRL.RAD21.",type,filt,CTCF,"motif5")]],xmin = -6.5, xmax = 1.2, ymin = 0.5, ymax = 1.5)+
coord_cartesian(xlim = c(-6.5,6.5),clip="off")+
theme(plot.margin = unit(c(1,2,2,1), "cm"),axis.title.x=element_text(size=20,face="bold",hjust=0.8))

#down no CTCF top3 skipping redundant ETS motifs
CTCF="noCTCF"
selectmotifs<-Motiflist[[paste0("SA2mutvsCTRL.RAD21.",type,filt,CTCF)]][c(1,2,4),]
selectmotifs$RANK<-factor(c(1,2,3)) #important: redifine Ranks for plotting levels!
p1<-MotifBarFunc(selectmotifs,Top=3,xmax=6,forcepvalcoord=6,fillcol="firebrick",motifname="NameFamily",sizeName=6)
plotlistbyCTCFsel[[paste0(filt,type,CTCF)]]<-p1+ 
annotation_custom(motifgroblistbyCTCF[[paste0("SA2mutvsCTRL.RAD21.",type,filt,CTCF,"motif1")]],xmin = -6.5, xmax = 0.5, ymin = 2.5, ymax = 3.5)+
annotation_custom(motifgroblistbyCTCF[[paste0("SA2mutvsCTRL.RAD21.",type,filt,CTCF,"motif2")]],xmin = -6.5, xmax = 1.2, ymin = 1.5, ymax = 2.5)+
annotation_custom(motifgroblistbyCTCF[[paste0("SA2mutvsCTRL.RAD21.",type,filt,CTCF,"motif4")]],xmin = -6.5, xmax = 1.2, ymin = 0.5, ymax = 1.5)+
coord_cartesian(xlim = c(-6.5,6.5),clip="off")+
theme(plot.margin = unit(c(1,2,2,1), "cm"),axis.title.x=element_text(size=20,face="bold",hjust=0.8))

pdf(file=file.path(FIGDIR,"TOPmotifBarplots","byCTCFoverlap","Top3_sel.motifs_SA2mut.stringentRAD21diffpeaks.by.CTCF.svgs.pdf"),height=14,width=8)
plot_grid(plotlist=plotlistbyCTCFsel,ncol=1,labels=names(plotlistbyCTCFsel))
dev.off()
```

## combined plot with HSPC diffpeak motifs
```{r}
plotlistHSPCs<-readRDS(file.path(FIGDIRCD34,"TOPmotifBarplots","plotlistsel.HSPC.RAD21diff.motifbarplts.rdata"))
comblist<-c(plotlistbyCTCFsel,plotlistHSPCs)

pdf(file=file.path(FIGDIR,"TOPmotifBarplots","byCTCFoverlap","Top3_sel.motifs_SA2mut.SA2KD.stringentRAD21diffpeaks.by.CTCF.svgs.pdf"),height=24,width=8)
plot_grid(plotlist=comblist,ncol=1,labels=names(comblist),rel_heights=c(8,8,8,8,5.5,12))
dev.off()
```


