---
title: "QC_p069"
author: "Inma Hernandez"
date: "1/4/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Merge demultiplexed objects


```{r,echo=FALSE}
rm(list = ls())
library(Seurat)
library(ggpubr)
library(tidyverse)
library(plyr)
library(ggplot2)
library(cowplot)
library(ggExtra)
library(harmony)
library(WriteXLS)
library(Nebulosa)
library(SCpubr)

path_to_sample<- str_c(
  "/Users/Inma/Documents/mnt/r26rci/projects/p069_af_10x/experiments/scRNA_scCSF/analysis/demultiplexed_rds_objects/CD34_NCKO/")
files_to_load <- list.files(path_to_sample)
#files_to_load <- files_to_load[1:2]
gem_ids <- files_to_load %>%
  str_remove("GEX_CD34_") %>%
  str_remove("_NCKO_library_run_1to3.rds")
files_to_load <- str_c(path_to_sample, files_to_load, sep = "")
path_to_results<- str_c(
  "/Users/Inma/Documents/mnt/r26rci/projects/p069_af_10x/experiments/scRNA_scCSF/analysis/demultiplexed_rds_objects/CD34_NCKO_results/")

#purrr::map() is a function for applying a function to each element of a list. 
#https://jennybc.github.io/purrr-tutorial/ls01_map-name-position-shortcuts.html
NCKO_list <- purrr::map(files_to_load, readRDS)
names(NCKO_list) <- gem_ids
path_to_metadata <- "/Users/Inma/Documents/mnt/r26rci/projects/p069_af_10x/metainfo/metadata_CD34_NCKO.xls"

path_to_save <- str_c(path_to_sample, "NCKO_merged.rds", sep = "")
path_to_save_markers <- str_c(path_to_sample, "markers_NCKO.xls", sep = "")

path_to_save <- str_c(path_to_results, "AML_healthy_merged.rds", sep = "")
path_to_results_Cluster <- str_c(path_to_results, "CD34_NCKO_integrated_FinalObject.rds", sep = "")

for (i in 1:length(NCKO_list)) {
  print(gem_ids[i])
  NCKO_list[[i]]$orig.ident <- gem_ids[i]
  NCKO_list[[i]]$group <- NCKO_list[[i]]$hash.ID
  NCKO_list[[i]]$group <- revalue(NCKO_list[[i]]$group, c("Sample1" = "NC", "Sample2"= "STAG2KO"))
  NCKO_list[[i]]$ID <- paste(NCKO_list[[i]]$group,NCKO_list[[i]]$orig.ident, sep = "_")
  print(table(NCKO_list[[i]]$group))
  Idents(NCKO_list[[i]]) <- "group"
  NCKO_list[[i]] <- subset(NCKO_list[[i]], idents = c("NC","STAG2KO"))
}

NCKO_hashed  <- merge(NCKO_list[[1]],
                       y = NCKO_list[2: length(NCKO_list)],
                       add.cell.ids = names(NCKO_list))



#Filtering by MT percent, number of features and number of counts

NCKO_hashed[["percent.mt"]] <- PercentageFeatureSet(NCKO_hashed, pattern = "^MT-")

VlnPlot(NCKO_hashed, 
        c("nFeature_RNA", "nCount_RNA", "percent.mt"), 
        ncol = 3, group.by = "ID", pt.size = -1)

```


## 2. Quality plots


```{r,echo=FALSE}
#plot common distribution number of counts, number of features
mid <- 5
p  = ggplot(NCKO_hashed@meta.data, aes(x = nCount_RNA, y = nFeature_RNA)) + geom_point(aes(color = percent.mt), size = 1) +
scale_color_gradient2(midpoint = mid, low = "black", mid = "grey", high ="red") + scale_y_continuous(trans = "log10") + scale_x_continuous(trans = "log10") + theme_cowplot(12) + geom_vline(xintercept = 750, linetype = 3) + geom_hline(yintercept = 250, linetype = 3) 
ggMarginal(p, type = "densigram", size = 10)
p

#plot individual distribution number of counts, number of features

ggplot(NCKO_hashed@meta.data, aes(x = nCount_RNA, y = nFeature_RNA)) + geom_point(aes(color = percent.mt), size = 1) +
scale_color_gradient2(midpoint = mid, low = "black", mid = "grey", high ="red") + scale_y_continuous(trans = "log10") + scale_x_continuous(trans = "log10") + theme_cowplot(12) + facet_wrap( ~ID, ncol = 2) + geom_vline(xintercept = 750, linetype = 3) + geom_hline(yintercept = 250, linetype = 3)

#plot number of genes

n_genes_hist1 <- NCKO_hashed@meta.data %>%
  ggplot(aes(nFeature_RNA)) +
    geom_histogram(bins = 100) +
    labs(x = "Number of Detected Genes", y = "Number of Cells") +
    # scale_x_log10() +
    theme_pubr()
n_genes_hist2 <- NCKO_hashed@meta.data %>%
  ggplot(aes(nFeature_RNA)) +
    geom_histogram(bins = 100) +
    scale_x_continuous(limits = c(0, 1000)) +
    labs(x = "Number of Detected Genes", y = "Number of Cells") +
    theme_pubr()

ggarrange(plotlist = list(n_genes_hist1, n_genes_hist2), ncol = 2)



n_genes_hist1 <- NCKO_hashed@meta.data %>%
  ggplot(aes(nCount_RNA)) +
    geom_histogram(bins = 100) +
    labs(x =  "Library Size (log10(total UMI))", y = "Number of Cells") +
    # scale_x_log10() +
    theme_pubr()
n_genes_hist2 <- NCKO_hashed@meta.data %>%
  ggplot(aes(nCount_RNA)) +
    geom_histogram(bins = 100) +
    scale_x_continuous(limits = c(0, 1000)) +
    labs(x =  "Library Size (log10(total UMI))", y = "Number of Cells") +
    theme_pubr()

ggarrange(plotlist = list(n_genes_hist1, n_genes_hist2), ncol = 2)


## Subset empty droplets and lysed cells

min_lib_size <- 750
min_n_genes <- 250
max_pct_mt <- 5
min_cells <- 5

metadata_before_qc <- NCKO_hashed@meta.data
is_low_quality <- 
  NCKO_hashed$nCount_RNA < min_lib_size |
  NCKO_hashed$nFeature_RNA < min_n_genes |
  NCKO_hashed$percent.mt > max_pct_mt
table(is_low_quality)
NCKO_hashed$keep_cells <- !is_low_quality
Idents(NCKO_hashed) <- "keep_cells"
table(NCKO_hashed$keep_cells)
NCKO_hashed <- subset(NCKO_hashed, idents = TRUE)
metadata_after_qc <- NCKO_hashed@meta.data

saveRDS(NCKO_hashed, file = path_to_save)
```

## 2. Normalization and integration


```{r,echo=FALSE}

#Clusters <- readRDS(path_to_save)

Clusters <- NCKO_hashed
Clusters <- NormalizeData(
  Clusters,
  normalization.method = "LogNormalize",
  scale.factor = 1e4
)

Clusters[["RNA"]]@data[1:10, 1:10]

Clusters <- Clusters %>%
  FindVariableFeatures(nfeatures = 3000) %>%
  ScaleData() %>% 
  RunPCA()

Clusters <- RunHarmony(Clusters, group.by.vars = "orig.ident")

Clusters <- Clusters %>% 
  RunUMAP(reduction = "harmony", dims = 1:50) %>% 
  FindNeighbors(reduction = "harmony", dims = 1:30) %>% 
  FindClusters(resolution = 0.5) %>% 
  identity()

Clusters@active.assay <- "RNA"
markers <- FindAllMarkers(Clusters, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.5)
markers <- markers %>% filter(p_val_adj < 0.05)
WriteXLS(markers, ExcelFileName = path_to_save_markers)

DimPlot(Clusters, split.by = "ID", label = TRUE)
DimPlot(Clusters, split.by = "group", label = TRUE)

markers %>%
    group_by(cluster) %>%
    top_n(n = 5, wt = avg_log2FC) -> top10
#DoHeatmap(Clusters, features = top10$gene) + NoLegend()



Idents(Clusters) <- "seurat_clusters"
new.cluster.ids <- c("HSC","MACDC","NEUT","MAST","MONO","BCELL","GMP","MAST","NEUT","ERY","CLP","GMP","MACDC")
names(new.cluster.ids) <- levels(Clusters)
Clusters <- RenameIdents(Clusters, new.cluster.ids)
Clusters[["Subclusters"]] <- Idents(Clusters)
Idents(Clusters) <- "Subclusters"
DimPlot(Clusters, label = TRUE)


cluster_cols <- c("HSC" = "#ECA809",                    # Prussian Blue.
                  "NEUT" = "#043362",                    # Prussian Blue.
                  "MAST" = "#009FF5",                     # Carolina Blue.
                  "MONO" = "#BC5210",            # Burnt Orange.
                  "BCELL" = "#e9d8a6",                  # Celadon Green.
                  "GMP" = "#7EB356",                   # Bud Green.
                  "ERY" = "#AC70FF",                   # Medium Purple.
                  "CLP" = "#63412C",                   # Van Dyke Brown.
                  "MACDC" = "#D6D6D6")

do_DimPlot(Clusters, group.by = "Subclusters", colors.use = cluster_cols)
library(magrittr)
do_BarPlot(Clusters,var.to.plot = "group", group.by = "Subclusters", position = "fill", horizontal =  FALSE)

Clusters@active.assay <- "RNA"
Idents(Clusters) <- "Subclusters"
saveRDS(Clusters, path_to_results_Cluster)


markers <- FindAllMarkers(Clusters, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.5)
markers <- markers %>% filter(p_val_adj < 0.05)
saveRDS(markers, file = "/Users/Inma/Documents/mnt/r26rci/projects/p069_af_10x/experiments/scRNA_scCSF/analysis/demultiplexed_rds_objects/CD34_NCKO/markers_annotat.rds")

DimPlot(Clusters, split.by = "ID", label = TRUE, repel = TRUE)
DimPlot(Clusters, split.by = "group", label = TRUE, repel = TRUE)

```


## Markers per cell type
```{r,echo=FALSE, fig.width=16, fig.height=12}
markers %>%
    group_by(cluster) %>%
    top_n(n = 5, wt = avg_log2FC) -> top10
DoHeatmap(Clusters, features = top10$gene) + NoLegend() + 
    theme(text = element_text(size = 20))

```

## HSC outstanding markers
```{r,echo=FALSE}
p3 <- plot_density(Clusters, c("MEF2C","SMC2","CDK1","S100B", "HMGB3", "FKBP4"))
p3 + plot_layout(ncol = 2)
```

## CLP outstanding markers
```{r,echo=FALSE}
p3 <- plot_density(Clusters, c("CD34","SOCS2","HOXA9","MEF2C"))
p3 + plot_layout(ncol = 2)
```

## GMP outstanding markers
```{r,echo=FALSE}
p3 <- plot_density(Clusters, c("CYTL1", "TPSAB1", "CFD", "CLEC11A", "GYPC"))
p3 + plot_layout(ncol = 2)
```

## MACDC outstanding markers
```{r,echo=FALSE}
p3 <- plot_density(Clusters, c("CD14","IRF8","NFKB1","CD4","SPI1", "CD86", "CLEC10A"))
p3 + plot_layout(ncol = 2)
```

## NEUT outstanding markers
```{r,echo=FALSE}
p3 <- plot_density(Clusters, c("CSF3R", "CLEC12A", "MYO1F", "FCN1", "MYB"))
p3 + plot_layout(ncol = 2)
```


## MAST outstanding markers
```{r,echo=FALSE}
p3 <- plot_density(Clusters, c("KIT","CD33","BTK","FCER1A","CCL2"))
p3 + plot_layout(ncol = 2)
```

## MONO outstanding markers
```{r,echo=FALSE}
p3 <- plot_density(Clusters, c("S100A8","S100A9","HLA-A","HLA-B","CD14","SPI1"))
p3 + plot_layout(ncol = 2)
```

## BCELL outstanding markers
```{r,echo=FALSE}
p3 <- plot_density(Clusters, c("SOX4","JCHAIN","HOXA9"))
p3 + plot_layout(ncol = 2)
```

## ERY outstanding markers
```{r,echo=FALSE}
p3 <- plot_density(Clusters, c("HBG1","HBB", "HBA1"))
p3 + plot_layout(ncol = 2)
```


## Barplot 
```{r,echo=FALSE, fig.width=12, fig.height=10}
ggplot(Clusters@meta.data, aes(x=group, fill=Subclusters)) + geom_bar(position = "fill") + scale_fill_manual(values=c("#BC3C29FF", "#0072B5FF", "#E18727FF","#20854EFF", "#7876B1FF", "#6F99ADFF","#FFDC91FF" ,"#EE4C97FF", "grey"))  + 
    theme(text = element_text(size = 20))
```


```{r,echo=FALSE, fig.width=12, fig.height=10}
ggplot(Clusters@meta.data, aes(x=ID, fill=Subclusters)) + geom_bar(position = "fill") + scale_fill_manual(values=c("#BC3C29FF", "#0072B5FF", "#E18727FF","#20854EFF", "#7876B1FF", "#6F99ADFF","#FFDC91FF" ,"#EE4C97FF", "grey"))  + 
    theme(text = element_text(size = 16))
```

## Projection of bulk RNA DE genes
```{r,echo=FALSE, fig.width=18, fig.height=18}
SA2KO = c("CD22","MMP9","ADAM12","SLC45A3","SPATS2L","PDE4DIP","CCR5","PHLDA2","PKLR")

p3 <- plot_density(Clusters, SA2KO)
p3 + plot_layout(ncol = 3)
```


```{r,echo=FALSE, fig.width=18, fig.height=18}
CONTROL = c("IGFBP5","PDE5A","ANK3","IGFBP5", "RAB6B","CALD1","MMRN1","CDHR1","SIDT1")

p4 <- plot_density(Clusters, CONTROL)
p4 + plot_layout(ncol = 3)
```

## genes of interest
```{r,echo=FALSE, fig.width=18, fig.height=18}
genes = c("ITGA9","PROM1",  "NPR3", "DACT1", "SMYD3")

p4 <- plot_density(Clusters, genes)
p4 + plot_layout(ncol = 3)
```
