---
title: "RNAseqCohesinKD_inCD34"
author: "Alexander Fischer"
date: "18 05 2020"
output: html_document



---
#rbioc_3-12
# Loading libraries needed for data processing and analysis
```{r, echo = FALSE, include = FALSE}
library(edgeR)
library(GGally)
library(ggplot2)
library(ggrepel)
library(gplots)
library(RColorBrewer)
library(reshape2)
library(sqldf)
library(ggpubr)
library(Rtsne)
library(dplyr)
library(readr)
library(tidyr)
library(stringr)
library(gridExtra)
```
# Defining path variables at the start
```{r}
DIRDATA="/misc/data/"
MAPLOGDIR<-file.path(DIRDATA,"processedData/mapping/RNA/GRCh38/RNAseq/CD34/logs")
PROJDIR=file.path(DIRDATA,"analysis/project_cohesin")
WORKDIR=file.path(PROJDIR,"CD34/RNAseq")
METADIR=file.path(WORKDIR,"Metadata")
#dir.create(METADIR)
RPGDIR=file.path(WORKDIR,"ReadsPerGene") #input directory with the .txt files containing the reads per gene information
dir.create(file.path(WORKDIR,"Analysis"))
ANALDIR=file.path(WORKDIR,"Analysis/Resulttables") #output of analysis results

FIGDIR=file.path(WORKDIR,"Analysis/Plots") #output of figure directories
ANALn="RNAseq_HSPCs_CohesinKD" #Name the particular analysis you are about to conduct

#path to the ShortTranscriptID for a fully functional annotation of the RCT
STIDp="/misc/software/ngs/genome/sequence/GRCh38.PRI_p10/STAR_transcriptIDshort.txt" 

##Directory for external Datasets for Comparison
ANALDIRAML<-file.path(PROJDIR,"Cohesin_AML/RNAseq/Analysis/Resulttables/RNAseq_Cohesin_AML_AF3")

dir.create(ANALDIR)
dir.create(file.path(ANALDIR,ANALn))
dir.create(file.path(FIGDIR,ANALn))
dir.create(file.path(FIGDIR,ANALn,"Clustering"))
dir.create(file.path(FIGDIR,ANALn,"GSEA"))
```
# Reading in metadata
```{r}
#reading in the metadata file as a pilot
metad<-read.table(file.path(METADIR,"Metadata_RNAseq_HSPCs_cohesin_KD.txt"),sep = "\t",header=T)
head(metad)
#consistency check
dim(metad)
```
# RCT Assembly in R
## generating short transcript ID variable
```{r}
#character vector containing the short transcript ID with consistency check
stid<-read.delim(STIDp,skip = 4,header = FALSE)[,1]
#consistency check
length(stid)
head(stid)
```
## get reads per gene deposited in RPGDIR
```{r}
#reading in all file names using an input directory only containing the files of interest..the variable "reads per gene list"
rpgl<-list.files(path = RPGDIR,pattern = "ReadsPerGene.txt",ignore.case = FALSE) #reading in the filenames as a list
length(rpgl) #this should be equal to the number of samples you want to analyse
#a sorting vector according to RNAseq_ID, later this will preserve the order when matching metadata information
sortvec<-sapply(as.character(metad$Sample_ID), function(x) grep(x,rpgl))
sortvec
length(sortvec)
#create an empty matrix as data frame, which can be written in in the following for-loop
counts<-as.data.frame(matrix(),row.names = NULL)
#in this forloop the read counts of the samples in the right order are extracted and appended to the counts data.frame
for (i in sortvec) {
rpgf<-read.delim(file.path(RPGDIR,rpgl[i]),sep = "\t",header = FALSE,check.names = TRUE,skip = 4,row.names = NULL)
counts<-cbind(counts,rpgf[,4])
}
counts<-counts[,(-1)] #leaving out the empty column resulting from generating the "empty" matrix
colnames(counts)<-metad$Sample_ID
rownames(counts)<-stid
ncol(counts) #all samples there?
head(counts)
```
## write output file for RCT
```{r}
write.table(counts,file=file.path(ANALDIR,ANALn,"RNAseq_HSPCs_cohesin_KD_counts_raw.txt"),sep = "\t",col.names=TRUE, quote=FALSE)
counts<-read.table(file=file.path(ANALDIR,ANALn,"RNAseq_HSPCs_cohesin_KD_counts_raw.txt"))
```

# Define group Variables
```{r}
#define factors and set levels
SI_target<-factor(as.character(metad$siRNA_target),levels=c("CTRL","SA1","SA2","SA1_SA2","RAD21")) #cave: level adjustment
siRNA12<-factor(as.character(metad$siRNA12),levels = c("Untr","siLuc","siLuc2x","Mock","SA1_2259","SA1_4094","poolSA1","SA2_452", "SA2_529","SA2_549","SA2_1252","poolSA2","SA1_2259_SA2_529","SA1_2259_SA2_549","SA1_2259_SA2_1252","SA1_4094_SA2_452","SA1_4094_SA2_529","SA1_4094_SA2_549","SA1_4094_SA2_1252","poolSA1_poolSA2","RAD21_467","RAD21_2031","RAD21pool","MED12_3408","MED12_4356","MED12pool")) #cave: level adjustment
donorb<-factor(as.character(metad$donor)) #no level adjustment necessary
labelID<-factor(as.character(metad$siRNA12_batch),levels = c("Untr_6","Untr_9","Untr_13", "siLuc_6","siLuc_7","siLuc_9","siLuc_13","siLuc2x_9","siLuc2x_13","Mock_6","Mock_7","Mock_9","Mock_13","SA1_2259_6","SA1_2259_9","SA1_2259_13","SA1_4094_6","SA1_4094_9","SA1_4094_13","poolSA1_9","poolSA1_13","SA2_452_6","SA2_452_7","SA2_529_7","SA2_529_9","SA2_529_13","SA2_549_7","SA2_549_9","SA2_549_13","SA2_1252_6","SA2_1252_7","SA2_1252_9","SA2_1252_13","SA2_1352_7","poolSA2_9","poolSA2_13","SA1_2259_SA2_529_9","SA1_2259_SA2_529_13","SA1_2259_SA2_549_9","SA1_2259_SA2_549_13","SA1_2259_SA2_1252_6","SA1_2259_SA2_1252_9","SA1_2259_SA2_1252_13","SA1_4094_SA2_529_9","SA1_4094_SA2_529_13","SA1_4094_SA2_549_9","SA1_4094_SA2_549_13","SA1_4094_SA2_1252_9","SA1_4094_SA2_1252_13","SA1_4094_SA2_452_6","poolSA1_SA2_9","poolSA1_SA2_13","RAD21_467_6","RAD21_467_9","RAD21_467_13","RAD21_2031_6","RAD21_2031_9","RAD21_2031_13","RAD21pool_9","RAD21pool_13","MED12_3408_9","MED12_3408_13","MED12_4356_9","MED12_4356_13","MED12pool_9","MED12pool_13"))

#Creater a genes data frame to add to the dglist object
#genens<-strsplit2(rownames(counts),"$",fixed = TRUE)[,2] #extract gene names from the ShortTranscriptID containing the gene names
genes.df<-as.data.frame(strsplit2(stid,"[$]"))
colnames(genes.df)<-c("EnsemblID","GeneSymbol","Length","GeneType")
genes.df$EnsemblID<-as.character(genes.df$EnsemblID)
genes.df$GeneSymbol<-as.character(genes.df$GeneSymbol)
genes.df$Length<-as.numeric(as.character(genes.df$Length))
genes.df$GeneType<-as.character(genes.df$GeneType)
head(genes.df)
```

# Generation and Filtering of DGEList object with all samples
```{r}
dgel <- DGEList(counts = counts, group = SI_target, genes = genes.df)
keep <- rowSums(cpm(dgel)>1) >= 3 
dgel <- dgel[keep, , keep.lib.sizes=FALSE]
dgel <- calcNormFactors(dgel)
summary(keep)
dgel$samples
```

# generation of normalized counts
```{r}
d.log.cpm <- cpm(dgel, prior.count = 2, log = TRUE)
d.log.rpkm <- rpkm(dgel, prior.count = 2, normalized.lib.sizes = TRUE, log = TRUE)
#batch correction of counts for donor
design_plots <- model.matrix(~0+SI_target)
d.corr.log.rpkm<-removeBatchEffect(d.log.rpkm,batch=donorb,design=design_plots)
write.table(d.corr.log.rpkm, quote=FALSE, sep = "\t", file=file.path(ANALDIR,ANALn,paste0("RNAseq.HSPCs.cohesinKD.corr.log.rpkm.txt")))
d.corr.log.cpm<-removeBatchEffect(d.log.cpm,batch=donorb,design=design_plots)
write.table(d.corr.log.cpm, quote=FALSE, sep = "\t", file=file.path(ANALDIR,ANALn,paste0("RNAseq.HSPCs.cohesinKD.corr.log.cpm.txt")))
```

# Clustering Analysis
## function for PCA
```{r}
#function for PCA plot
pca_func3<-function(data,group,labels){
  prin_comp.d <- prcomp(data, scale. = FALSE)
  std_dev.d <- prin_comp.d$sdev 
  pr_var.d <- std_dev.d^2
  prop_varex.d <- round(100*pr_var.d/sum(pr_var.d), digits=1)
  embedding <- as.data.frame(prin_comp.d$rotation)
  embedding$Class <- as.factor(group)
  p1 <- ggplot(data=embedding) +
    aes(x=PC1, y=PC2)+
      geom_point(size=rel(8), aes(colour=Class)) +
      scale_colour_manual(values = c("CTRL" = "firebrick1", "SA1" = "darkgoldenrod2", "SA2" = "lightgreen", "SA1_SA2" = "lightskyblue2", "RAD21" = "mediumorchid3"),
      name="HSPC-group",
      labels=c("CTRL","STAG1 KD","STAG2 KD","STAG1+2 KD","RAD21 KD")) +
      geom_text_repel(aes(label=labels, size=20), col="black", segment.colour="black", segment.size=0.4, min.segment.length=0.9, size=5.5, point.padding=.5,segment.alpha=0.5, alpha=1, size=20) +
      ggtitle("PCA: RNAseq HSPCs")  +
      xlab(paste("PC2: ",prop_varex.d[2],"% of variance")) + ylab(paste("PC1: ",prop_varex.d[1],"% of variance")) +
theme_light() +
    theme(plot.tag=element_text(size = 12*2.0, face = "bold"),
          plot.title = element_text(size = rel(2.4), face = "plain", hjust = 0.5),
          legend.text = element_text(colour="black", size = rel(1.2), face = "plain"),
          legend.title = element_text(colour="black", size = rel(1.2), face = "bold"),
          legend.box.just = "top",
          axis.text = element_text(colour = "black", size = rel(1.5), face = "plain"),
          axis.title = element_text(colour = "black",size = rel(2),face = "plain"),
          panel.grid = element_blank(),
          panel.border = element_rect(colour = "black",size=2),
          axis.ticks = element_line(colour="black"),
          aspect.ratio = 1.0, #plot panel aspect ratio for squared == 1.0! :)
          legend.position = c(0.21,0.51),legend.background = element_rect(fill = "transparent"),legend.box.background = element_rect(colour = "darkgrey")
        )
  plot(p1)
}
#variables for PCA
nolabs=c("")
```
## function for UMAP
```{r}
##define a umap function for variable seeds and inputdata
umapfunc<-function(data,seed=42,legpos="right",pchsize=rel(8),labels=nolabs,groupvec=SI_target,donorvec=donorb,x1=NA,x2=NA,y1=NA,y2=NA){
set.seed(seed = seed)
  #appply umap algorithm on transposed data matrix, then extract layout for plotting
  umap.dat<-umap::umap(t(as.matrix(data)))
  umap.dat<-data.frame(umap.dat$layout)
  ##add KD group and donor IDs to umap.dat
  umap.dat$group<-groupvec
  umap.dat$donor<-donorvec
  umap.p<-ggplot(data =umap.dat)+
    aes(x = X1, y = X2)+
    xlab("UMAP2") +
    ylab("UMAP1") +
    scale_x_continuous(limits = c(x1,x2)) +
    scale_y_continuous(limits = c(y1,y2))+    
    geom_point(size=pchsize,aes(color=group))+
      scale_color_manual(values = c("CTRL" = "firebrick1", "SA1" = "darkgoldenrod2", "SA2" = "lightgreen", "SA1_SA2" = "lightskyblue2", "RAD21" = "mediumorchid3"),
      name="HSPC-group",
      labels=c("CTRL","STAG1 KD","STAG2 KD","STAG1+2 KD","RAD21 KD")) +
    geom_text_repel(aes(label=labels),size=4,segment.size=0.2,min.segment.length=0.0,point.padding=.05,segment.alpha=0.5,max.overlaps = 50,force = 50, show.legend = FALSE) +
    labs(title = paste0("UMAP: RNAseq HSPCs seed", seed))+
    theme_light(base_size=12) +
    theme(plot.tag=element_text(size = 12*2.0, face = "bold"),
          plot.title = element_text(size = rel(2.4), face = "plain", hjust = 0.5),
          plot.title.position = "panel",
          legend.text = element_text(colour="black", size = rel(1.2), face = "plain"),
          legend.title = element_text(colour="black", size = rel(1.2), face = "bold"),
          legend.box.just = "top",
          axis.text = element_text(colour = "black", size = rel(1.5), face = "plain"),
          axis.title = element_text(colour = "black",size = rel(2),face = "plain"),
          panel.grid = element_blank(),
          panel.border = element_rect(colour = "black",size=2),
          axis.ticks = element_line(colour="black"),
          aspect.ratio = 1.0,
          legend.position = legpos,legend.background = element_rect(fill = "transparent"),legend.box.background = element_rect(colour = "darkgrey")
        )
}      
```


## plot PCA and UMAP
```{r}
#PCA
#without batch corr.
pdf(file = file.path(FIGDIR,ANALn,"Clustering","PCA.donorIDlabels.pdf"),height = 8,width = 12)
pca_func3(d.log.cpm,SI_target,donorb)
dev.off()
#with batch corr
pdf(file = file.path(FIGDIR,ANALn,"Clustering","PCA_wBC.donorIDlabels.pdf"),height = 8,width = 12)
pca_func3(d.corr.log.cpm,SI_target,donorb)
dev.off()
pdf(file = file.path(FIGDIR,ANALn,"Clustering","PCA_wBC.nolabels.pdf"),height = 8,width = 12)
pca_func3(d.corr.log.cpm,SI_target,nolabs)
dev.off()

#UMAP - with batch corr.
pdf(file = file.path(FIGDIR,ANALn,"Clustering","UMAP_wBC.nolab.noleg.logcpm.pdf"),height = 8,width = 12)
plot(umapfunc(data=d.corr.log.cpm ,seed=142,pchsize=rel(8),legpos="none"))
dev.off()
```


# DEGs by siRNA Target
## Calculate dispersion and fit the model for design
```{r}
#estimate dispersion
design_DEGs_target2 <- model.matrix(~0+SI_target+donorb)
dgel<- estimateDisp(dgel,design_DEGs_target2,robust = TRUE)
dgel$common.dispersion #Output:  0.01595838
#Visualize dispersion
pdf(file= file.path(FIGDIR,ANALn,"dispersion.bcv.plot.HSPCs.KDvsCTRL.pdf"), height=8, width=12)
plotBCV(dgel)
dev.off()
#save the dglist oject 
saveRDS(dgel,file=file.path(ANALDIR,ANALn,"dgel.rdata"))
#Fitting genewise glms
f4<-glmQLFit(dgel,design_DEGs_target2)
saveRDS(f4,file=file.path(ANALDIR,ANALn,"fit.f4.rdata"))
```
## Contrasts: siRNA-Target vs CTRL (siLuc+Mock)
```{r}
#comparison siRNAs vs CTRL(siCTRL)
Targets <- levels(SI_target)[2:length(levels(SI_target))]
for (i in Targets) {
##define variable names
#important variables specifiying the contrast and paths/names of outputfiles
comppair = paste0(paste0("SI_target",i),"-SI_targetCTRL")
tablepath = file.path(ANALDIR,ANALn,paste0(paste0("qstat_",i),"vs.CTRL.glm.txt"))
DEGsumpath = file.path(ANALDIR,ANALn,paste0(paste0("DEGsummary_",i),".vs.CTRL.txt"))

#make Contrasts
con <- makeContrasts(comppair, levels = design_DEGs_target2)
qlf <-glmQLFTest(f4,contrast = con)
qstat <- topTags(qlf, n = Inf)
sumDEG <- summary(qdt <- decideTestsDGE(qlf))
sumDEG 
write.table(sumDEG, file = DEGsumpath, sep = "\t", col.names=NA, quote=FALSE)
write.table(qstat, file = tablepath, sep = "\t", col.names=NA, quote=FALSE)
#add FPKM to qstat table
qstatmod <- as.data.frame(qstat)
qstatmod$logFPKM <- qstatmod$logCPM - log2(as.numeric(as.character(qstatmod$Length)) * 0.001)
write.table(qstatmod, file.path(ANALDIR,ANALn,paste0(paste0("qstat_",i),"vs.CTRL.glm.extended.txt")), sep = "\t", col.names=NA, quote=FALSE)
#create smearplot
qisDE <- as.logical(qdt)
qDEnames <- rownames(dgel)[qisDE]
pdf(file= file.path(FIGDIR,ANALn,paste0("qlf.results.smearplot.",i,"KDvsCTRL.pdf")), height=8, width=12)
plotSmear(qlf, de.tags=qDEnames)
dev.off()
}
```
## Filtering of DEGs
```{r}
## read in the generated qstat tables as varibles for filtering, pvalues to show etc
qstat_SA2vs.CTRL <- read.table(file=file.path(ANALDIR,ANALn,"qstat_SA2vs.CTRL.glm.extended.txt"),header = TRUE,row.name = 1,sep = "",dec = ".")
qstat_SA1vs.CTRL <- read.table(file=file.path(ANALDIR,ANALn,"qstat_SA1vs.CTRL.glm.extended.txt"),header = TRUE,row.name = 1,sep = "",dec = ".")
qstat_SA1_SA2vs.CTRL <- read.table(file=file.path(ANALDIR,ANALn,"qstat_SA1_SA2vs.CTRL.glm.extended.txt"),header = TRUE,row.name = 1,sep = "",dec = ".")
qstat_RAD21vs.CTRL <- read.table(file=file.path(ANALDIR,ANALn,"qstat_RAD21vs.CTRL.glm.extended.txt"),header = TRUE,row.name = 1,sep = "",dec = ".")
#save them in a list
qslist<-list(qstat_SA2vs.CTRL,qstat_SA1vs.CTRL,qstat_SA1_SA2vs.CTRL,qstat_RAD21vs.CTRL)
names(qslist)<-c("qstat_SA2vs.CTRL","qstat_SA1vs.CTRL","qstat_SA1_SA2vs.CTRL","qstat_RAD21vs.CTRL")

#filter all comparisons for DEGs and add to qslist
for (i in Targets) {
qslist[[paste0("DEGs_FC2up.",i)]]<-filter(qslist[[paste0("qstat_",i,"vs.CTRL")]],logFC>(1) & FDR <0.05 & logCPM>1)
qslist[[paste0("DEGs_FC2down.",i)]]<-filter(qslist[[paste0("qstat_",i,"vs.CTRL")]],logFC<(-1) & FDR <0.05 & logCPM>1)
qslist[[paste0("DEGs_FC1.5up.",i)]]<-filter(qslist[[paste0("qstat_",i,"vs.CTRL")]],logFC>(0.585) & FDR <0.05 & logCPM>1)
qslist[[paste0("DEGs_FC1.5down.",i)]]<-filter(qslist[[paste0("qstat_",i,"vs.CTRL")]],logFC<(-0.585) & FDR <0.05 & logCPM>1)
qslist[[paste0("DEGs_FC1.5up.nocpmfilt",i)]]<-filter(qslist[[paste0("qstat_",i,"vs.CTRL")]],logFC>(0.585) & FDR <0.05)
qslist[[paste0("DEGs_FC1.5down.nocpmfilt",i)]]<-filter(qslist[[paste0("qstat_",i,"vs.CTRL")]],logFC<(-0.585) & FDR <0.05)
}
#filter SA2KD DEGs in separate objects as they will be used very frequently
SA2KDdownDEGScpm1 <- filter(qstat_SA2vs.CTRL, qstat_SA2vs.CTRL$logFC <= (-0.585) & qstat_SA2vs.CTRL$FDR <= 0.05 & qstat_SA2vs.CTRL$logCPM > 1) #546
SA2KDupDEGScpm1 <- filter(qstat_SA2vs.CTRL, qstat_SA2vs.CTRL$logFC >= (0.585) & qstat_SA2vs.CTRL$FDR <= 0.05 & qstat_SA2vs.CTRL$logCPM > 1) #163

#get FC values for cohesin genes
kdtargetFC<-data.frame()
kdtargetgenes<-c("STAG1","STAG2","RAD21")
for (i in Targets) {
for (gene in kdtargetgenes) {
kdtargetFC[gene,i]<-qslist[[paste0("qstat_",i,"vs.CTRL")]][qslist[[paste0("qstat_",i,"vs.CTRL")]]$GeneSymbol==gene,"logFC"]
}}
kdtargetFC
#              SA1        SA2     SA1_SA2      RAD21
#STAG1 -1.53668653  0.3088120 -0.46764022  1.2478082
#STAG2  0.06936171 -1.6643468 -1.57161458  0.2657389
#RAD21  0.02354872  0.1504337 -0.04886098 -1.8896551

#count DEGS in a dataframe
DEGstats=data.frame()
for (i in Targets) {
  DEGstats["DEGs_FC2up",i]<-nrow(filter(qslist[[paste0("qstat_",i,"vs.CTRL")]],logFC>(1) & FDR <0.05 & logCPM>1))
  DEGstats["DEGs_FC2down",i]<-nrow(filter(qslist[[paste0("qstat_",i,"vs.CTRL")]],logFC<(-1) & FDR <0.05 & logCPM>1))
  DEGstats["DEGs_FC1.5up",i]<-nrow(filter(qslist[[paste0("qstat_",i,"vs.CTRL")]],logFC>(0.585) & FDR <0.05 & logCPM>1))
  DEGstats["DEGs_FC1.5down",i]<-nrow(filter(qslist[[paste0("qstat_",i,"vs.CTRL")]],logFC<(-0.585) & FDR <0.05 & logCPM>1))
  DEGstats["DEGs_FC1.5up.nocpmfilt",i]<-nrow(filter(qslist[[paste0("qstat_",i,"vs.CTRL")]],logFC>(0.585) & FDR <0.05))
  DEGstats["DEGs_FC1.5down.nocpmfilt",i]<-nrow(filter(qslist[[paste0("qstat_",i,"vs.CTRL")]],logFC<(-0.585) & FDR <0.05))
}
DEGstats
#                         SA1 SA2 SA1_SA2 RAD21
#DEGs_FC2up                 1  45     185   380
#DEGs_FC2down              18 213     468   298
#DEGs_FC1.5up              10 163     513   884
#DEGs_FC1.5down            81 546     865   731
#DEGs_FC1.5up.nocpmfilt    17 326     895  1396
#DEGs_FC1.5down.nocpmfilt 125 839    1342  1196
```


# Volcanoplots: KD vs CTRL results
## Sorting of qstat results table and defining colouring categories
```{r}
#use qstat list to prepare data for volcano plot
for (i in Targets) {
  #order and filter data by FDR and create new dfs sorted by logFC decreasing or increasing
  qslist[[paste0("qstat_",i,"vs.CTRL")]]<-qslist[[paste0("qstat_",i,"vs.CTRL")]][order(qslist[[paste0("qstat_",i,"vs.CTRL")]]$FDR, decreasing = FALSE), ]
  qslist[[paste0("qstat_",i,"vs.CTRL.FDR_filt")]]<-filter(qslist[[paste0("qstat_",i,"vs.CTRL")]],(logFC>(1)| logFC<(-1))& FDR <0.05)
  qslist[[paste0("qstat_",i,"vs.CTRL_ordered_FCdec_FDR_filt")]]<-qslist[[paste0("qstat_",i,"vs.CTRL.FDR_filt")]][order(qslist[[paste0("qstat_",i,"vs.CTRL.FDR_filt")]]$logFC, decreasing = TRUE), ]
  qslist[[paste0("qstat_",i,"vs.CTRL_ordered_FCinc_FDR_filt")]]<-qslist[[paste0("qstat_",i,"vs.CTRL.FDR_filt")]][order(qslist[[paste0("qstat_",i,"vs.CTRL.FDR_filt")]]$logFC, decreasing = FALSE), ]
  #assign up / down / not sig categories in new column 
  qslist[[paste0("qstat_",i,"vs.CTRL")]]$Volccolor<- "NA"
  qslist[[paste0("qstat_",i,"vs.CTRL")]]$Volccolor[qslist[[paste0("qstat_",i,"vs.CTRL")]]$logFC < -0.585] <- "down" 
  qslist[[paste0("qstat_",i,"vs.CTRL")]]$Volccolor[qslist[[paste0("qstat_",i,"vs.CTRL")]]$logFC > 0.585] <- "up" 
  qslist[[paste0("qstat_",i,"vs.CTRL")]]$Volccolor[qslist[[paste0("qstat_",i,"vs.CTRL")]]$Volccolor=="NA"] <- "notsig"
  qslist[[paste0("qstat_",i,"vs.CTRL")]]$Volccolor[qslist[[paste0("qstat_",i,"vs.CTRL")]]$FDR>0.05] <- "notsig"
}
#add desired colors for upregulated genes depending on KD to list
qslist[[paste0("UPcolor.SA2vs.CTRL")]]<-"lightgreen"
qslist[[paste0("UPcolor.SA1vs.CTRL")]]<-"darkgoldenrod2"
qslist[[paste0("UPcolor.SA1_SA2vs.CTRL")]]<-"lightskyblue2"
qslist[[paste0("UPcolor.RAD21vs.CTRL")]]<-"mediumorchid3"
```

## generate Volcanos for all comparions showing topFC genes
```{r}
dir.create(file.path(FIGDIR,ANALn,"Volcanoplots"))
Volclist<-list()

for (i in Targets) {
Volclist[[i]]<-ggplot()+
  geom_point(data=qslist[[paste0("qstat_",i,"vs.CTRL")]], aes(x=logFC, y=-log10(FDR), colour=Volccolor)) +
  scale_color_manual(values = c("notsig" = "darkgrey", "up" = qslist[[paste0("UPcolor.",i,"vs.CTRL")]], "down" = "firebrick2")) +
  ggtitle(paste0(i," KD vs. CTRL-HSPCs")) +
  xlab("log2 fold change") +
  ylab("-log10 FDR") +
  scale_x_continuous(limits = c(-7,10)) +
  scale_y_continuous(limits = c(0,25))+    
  theme(legend.position = "none",
        plot.background = element_rect(fill="white",colour = "white",linetype = 3),
        panel.background = element_rect(fill = "white",color = "black"),
        panel.border = element_rect(fill = NA, colour = "black",size=2),
        panel.grid.major = element_line(size = 0.5, linetype = 'solid',colour = NULL),
        panel.grid.minor = element_line(size = 0.25, linetype = 'dashed',colour = NULL),
        plot.title = element_text(size = rel(2.4), hjust = 0.5),axis.text = element_text(colour = "black", size = rel(1.5), face = "plain"),
        axis.title = element_text(size = rel(2)))+
geom_text_repel(aes(x = qslist[[paste0("qstat_",i,"vs.CTRL_ordered_FCinc_FDR_filt")]]$logFC[1:10], y = -log10(qslist[[paste0("qstat_",i,"vs.CTRL_ordered_FCinc_FDR_filt")]]$FDR[1:10])), data =    head(qslist[[paste0("qstat_",i,"vs.CTRL_ordered_FCinc_FDR_filt")]], 10), label = qslist[[paste0("qstat_",i,"vs.CTRL_ordered_FCinc_FDR_filt")]]$GeneSymbol[1:10],cex=5,max.overlaps=30) +
geom_text_repel(aes(x = qslist[[paste0("qstat_",i,"vs.CTRL_ordered_FCdec_FDR_filt")]]$logFC[1:10], y = -log10(qslist[[paste0("qstat_",i,"vs.CTRL_ordered_FCdec_FDR_filt")]]$FDR[1:10])), data =    head(qslist[[paste0("qstat_",i,"vs.CTRL_ordered_FCdec_FDR_filt")]], 10), label = qslist[[paste0("qstat_",i,"vs.CTRL_ordered_FCdec_FDR_filt")]]$GeneSymbol[1:10],cex=5,max.overlaps=30) +
geom_text_repel(aes(x = qslist[[paste0("qstat_",i,"vs.CTRL")]]$logFC[1:10], y = -log10(qslist[[paste0("qstat_",i,"vs.CTRL")]]$FDR[1:10])), data =    head(qslist[[paste0("qstat_",i,"vs.CTRL")]], 10), label = qslist[[paste0("qstat_",i,"vs.CTRL")]]$GeneSymbol[1:10],cex=5,max.overlaps=30) +
       geom_vline(xintercept = -0.585, linetype = 'dashed') +  geom_vline(xintercept = 0.585,linetype = 'dashed') +
       geom_hline(yintercept = -log10(0.05),linetype = 'dashed')
}

#save all
for (i in Targets) {
ggsave(plot=Volclist[[i]],file=file.path(FIGDIR,ANALn,"Volcanoplots",paste0("Volcano_",i,"KD.vs.CTRL.topGenes.pdf")),height=6,width=6)
}



#adjust plot windows (x/y lims) for SA1 and SA2 KD plots
i<-"SA1"
pdf(file = file.path(FIGDIR,ANALn,"Volcanoplots",paste0("Volcano_",i,"KD.vs.CTRL.topGenes.pdf")), width = 6, height = 6)
ggplot()+
  geom_point(data=qslist[[paste0("qstat_",i,"vs.CTRL")]], aes(x=logFC, y=-log10(FDR), colour=Volccolor)) +
  scale_color_manual(values = c("notsig" = "darkgrey", "up" = qslist[[paste0("UPcolor.",i,"vs.CTRL")]], "down" = "firebrick2")) +
  ggtitle(paste0(i," KD vs. CTRL-HSPCs")) +
  xlab("log2 fold change") +
  ylab("-log10 FDR") +
  scale_x_continuous(limits = c(-3,3)) +
  scale_y_continuous(limits = c(0,15))+    
  theme(legend.position = "none",
        plot.background = element_rect(fill="white",colour = "white",linetype = 3),
        panel.background = element_rect(fill = "white",color = "black"),
        panel.border = element_rect(fill = NA, colour = "black",size=2),
        panel.grid.major = element_line(size = 0.5, linetype = 'solid',colour = NULL),
        panel.grid.minor = element_line(size = 0.25, linetype = 'dashed',colour = NULL),
        plot.title = element_text(size = rel(2.4), hjust = 0.5),axis.text = element_text(colour = "black", size = rel(1.5), face = "plain"),
        axis.title = element_text(size = rel(2)))+
geom_text_repel(aes(x = qslist[[paste0("qstat_",i,"vs.CTRL_ordered_FCinc_FDR_filt")]]$logFC[1:10], y = -log10(qslist[[paste0("qstat_",i,"vs.CTRL_ordered_FCinc_FDR_filt")]]$FDR[1:10])), data =    head(qslist[[paste0("qstat_",i,"vs.CTRL_ordered_FCinc_FDR_filt")]], 10), label = qslist[[paste0("qstat_",i,"vs.CTRL_ordered_FCinc_FDR_filt")]]$GeneSymbol[1:10],cex=5,max.overlaps=30) +
geom_text_repel(aes(x = qslist[[paste0("qstat_",i,"vs.CTRL_ordered_FCdec_FDR_filt")]]$logFC[1:10], y = -log10(qslist[[paste0("qstat_",i,"vs.CTRL_ordered_FCdec_FDR_filt")]]$FDR[1:10])), data =    head(qslist[[paste0("qstat_",i,"vs.CTRL_ordered_FCdec_FDR_filt")]], 10), label = qslist[[paste0("qstat_",i,"vs.CTRL_ordered_FCdec_FDR_filt")]]$GeneSymbol[1:10],cex=5,max.overlaps=30) +
geom_text_repel(aes(x = qslist[[paste0("qstat_",i,"vs.CTRL")]]$logFC[1:10], y = -log10(qslist[[paste0("qstat_",i,"vs.CTRL")]]$FDR[1:10])), data =    head(qslist[[paste0("qstat_",i,"vs.CTRL")]], 10), label = qslist[[paste0("qstat_",i,"vs.CTRL")]]$GeneSymbol[1:10],cex=5,max.overlaps=30) +
       geom_vline(xintercept = -0.585, linetype = 'dashed') +  geom_vline(xintercept = 0.585,linetype = 'dashed') +
       geom_hline(yintercept = -log10(0.05),linetype = 'dashed')
dev.off()

i<-"SA2"
SA2volc<-
ggplot()+
  geom_point(data=qslist[[paste0("qstat_",i,"vs.CTRL")]], aes(x=logFC, y=-log10(FDR), colour=Volccolor)) +
  scale_color_manual(values = c("notsig" = "darkgrey", "up" = qslist[[paste0("UPcolor.",i,"vs.CTRL")]], "down" = "firebrick2")) +
  ggtitle(paste0(i," KD vs. CTRL-HSPCs")) +
  xlab("log2 fold change") +
  ylab("-log10 FDR") +
  scale_x_continuous(limits = c(-6,4)) +
#  scale_y_continuous(limits = c(0,15))+    
  theme(legend.position = "none",
        plot.background = element_rect(fill="white",colour = "white",linetype = 3),
        panel.background = element_rect(fill = "white",color = "black"),
        panel.border = element_rect(fill = NA, colour = "black",size=2),
        panel.grid.major = element_line(size = 0.5, linetype = 'solid',colour = NULL),
        panel.grid.minor = element_line(size = 0.25, linetype = 'dashed',colour = NULL),
        plot.title = element_text(size = rel(2.4), hjust = 0.5),axis.text = element_text(colour = "black", size = rel(1.5), face = "plain"),
        axis.title = element_text(size = rel(2)))+
geom_text_repel(aes(x = qslist[[paste0("qstat_",i,"vs.CTRL_ordered_FCinc_FDR_filt")]]$logFC[1:10], y = -log10(qslist[[paste0("qstat_",i,"vs.CTRL_ordered_FCinc_FDR_filt")]]$FDR[1:10])), data =    head(qslist[[paste0("qstat_",i,"vs.CTRL_ordered_FCinc_FDR_filt")]], 10), label = qslist[[paste0("qstat_",i,"vs.CTRL_ordered_FCinc_FDR_filt")]]$GeneSymbol[1:10],cex=5,max.overlaps=30,min.segment.length=0) +
geom_text_repel(aes(x = qslist[[paste0("qstat_",i,"vs.CTRL_ordered_FCdec_FDR_filt")]]$logFC[1:10], y = -log10(qslist[[paste0("qstat_",i,"vs.CTRL_ordered_FCdec_FDR_filt")]]$FDR[1:10])), data =    head(qslist[[paste0("qstat_",i,"vs.CTRL_ordered_FCdec_FDR_filt")]], 10), label = qslist[[paste0("qstat_",i,"vs.CTRL_ordered_FCdec_FDR_filt")]]$GeneSymbol[1:10],cex=5,max.overlaps=30,min.segment.length=0) +
geom_text_repel(aes(x = qslist[[paste0("qstat_",i,"vs.CTRL")]]$logFC[1:10], y = -log10(qslist[[paste0("qstat_",i,"vs.CTRL")]]$FDR[1:10])), data =    head(qslist[[paste0("qstat_",i,"vs.CTRL")]], 10), label = qslist[[paste0("qstat_",i,"vs.CTRL")]]$GeneSymbol[1:10],cex=5,max.overlaps=30,min.segment.length=0) +
       geom_vline(xintercept = -0.585, linetype = 'dashed') +  geom_vline(xintercept = 0.585,linetype = 'dashed') +
       geom_hline(yintercept = -log10(0.05),linetype = 'dashed')


pdf(file = file.path(FIGDIR,ANALn,"Volcanoplots",paste0("Volcano_",i,"KD.vs.CTRL.topGenes.2.pdf")), width = 6, height = 6)
plot(SA2volc)
dev.off()
pdf(file = file.path(FIGDIR,ANALn,"Volcanoplots",paste0("Volcano_",i,"KD.vs.CTRL.topGenes.pdf")), width = 6, height = 9) #other size format
plot(SA2volc)
dev.off()
```



# DEG overlap between KDs
```{r}
dir.create(file.path(FIGDIR,ANALn,"Venn_diagrams"))
##collect DEGs in two lists for up and down:all vs all
KDs_down_DEGs<-list(qslist[["DEGs_FC1.5down.SA2"]]$GeneSymbol,qslist[["DEGs_FC1.5down.SA1"]]$GeneSymbol,qslist[["DEGs_FC1.5down.SA1_SA2"]]$GeneSymbol,qslist[["DEGs_FC1.5down.RAD21"]]$GeneSymbol)
names(KDs_down_DEGs)<-c("SA2KD","SA1KD","SA1_SA2KD","RAD21KD")
KDs_up_DEGs<-list(qslist[["DEGs_FC1.5up.SA2"]]$GeneSymbol,qslist[["DEGs_FC1.5up.SA1"]]$GeneSymbol,qslist[["DEGs_FC1.5up.SA1_SA2"]]$GeneSymbol,qslist[["DEGs_FC1.5up.RAD21"]]$GeneSymbol)
names(KDs_up_DEGs)<-c("SA2KD","SA1KD","SA1_SA2KD","RAD21KD")
##generate Venn objects:all vs all
venn_KDs_down_DEGs<-venn(KDs_down_DEGs)
venn_KDs_up_DEGs<-venn(KDs_up_DEGs)

##save venn plots all vs all
pdf(file = file.path(FIGDIR,ANALn,"Venn_diagrams","Venn_DEGs_down_allKDs.pdf"), width = 5, height = 5)
venn(KDs_down_DEGs)
dev.off()
pdf(file = file.path(FIGDIR,ANALn,"Venn_diagrams","Venn_DEGs_up_allKDs.pdf"), width = 5, height = 5)
venn(KDs_up_DEGs)
dev.off()
##ID of overlapping DEGs
DownOverl.allvsall<-attr(venn_KDs_down_DEGs,"intersections")$`SA2KD:SA1KD:SA1_SA2KD:RAD21KD`
DownOverl.allvsall
UpOverl.allvsall<-attr(venn_KDs_up_DEGs,"intersections")$`SA2KD:SA1KD:SA1_SA2KD:RAD21KD`
UpOverl.allvsall
```


# Expression of GOIs
## mean rpkms for GEX barplots
```{r}
##mean values log rpkms for sel2
dat.st2<-stack(as.data.frame(d.corr.log.rpkm))   ##stack data to a single column
dat.st2$genename<- rep(rownames(d.corr.log.rpkm), times = ncol(d.corr.log.rpkm))  ##add genenames as extra col
dat.st2$type=rep(SI_target, each=length(d.corr.log.rpkm[,1]))    ##add group identity as extra col
library(plyr)
mean.d.corr.log.rpkm<-ddply(dat.st2, .(type,genename),
summarize,
mean= round(mean(values),2),
sd=round(sd(values),2))   ##get means and standard deviation by group
```

##  GENEX barplot function showing the individual values of the repliecates as dots and displaying significance level of FDR vs CTRL
```{r}
collabvec<-c("CTRL\nHSPCs","STAG2\nKD","STAG1\nKD","STAG1+2\nKD","RAD21\nKD")
GENEX_func_means_dots<-function(GOI,y1=NA,y2=NA,bottom="",legloc="none",pointsize=rel(2),col.labs=collabvec){
GOIsearchterm<-paste0("\\$",GOI,"\\$")  
GOIdataPoints<-dat.st2[grep(GOIsearchterm,dat.st2$genename),]
GOIdataPoints$sd<-0
colnames(GOIdataPoints)<-c("value","label","genename","group","sd")
rownames(GOIdataPoints)<-GOIdataPoints$label
#average data of all patients
GOIdata<-mean.d.corr.log.rpkm[grep(GOIsearchterm,mean.d.corr.log.rpkm$genename),]
GOIdata$label<-c("CTRL-average","STAG1KD-average","STAG2KD-average","STAG1+2KD-average","RAD21KD-average")
col_order <- c("mean","label","genename","type","sd")
GOIdata2 <- GOIdata[, col_order]
rownames(GOIdata2)<-GOIdata2$label
colnames(GOIdata2)<-c("value","label","genename","group","sd")
GOIdatM<-rbind(GOIdata2,GOIdataPoints) ##merge the two
GOIdatM$group<-factor(GOIdatM$group,levels=c("CTRL","SA2","SA1","SA1_SA2","RAD21"))
pvaldataSA2<-qstat_SA2vs.CTRL[grep(GOIsearchterm,rownames(qstat_SA2vs.CTRL)),]
pvalSA2<-pvaldataSA2$FDR
pvaldataRAD21<-qstat_RAD21vs.CTRL[grep(GOIsearchterm,rownames(qstat_RAD21vs.CTRL)),]
pvalRAD21<-pvaldataRAD21$FDR
pvaldataSA1<-qstat_SA1vs.CTRL[grep(GOIsearchterm,rownames(qstat_SA1vs.CTRL)),]
pvalSA1<-pvaldataSA1$FDR
pvaldataSADKO<-qstat_SA1_SA2vs.CTRL[grep(GOIsearchterm,rownames(qstat_SA1_SA2vs.CTRL)),]
pvalSADKO<-pvaldataSADKO$FDR
#symbol is assigned to significance level for SA2 and RAD21 group
if (pvalSA2 < 0.001) { siglvlSA2 <- "***" } else if (pvalSA2 < 0.01) { siglvlSA2 <- "**" } else if (pvalSA2 < 0.05) { siglvlSA2 <- "*" } else { siglvlSA2 <- "ns" }
if (pvalRAD21 < 0.001) { siglvlRAD21 <- "***" } else if (pvalRAD21 < 0.01) { siglvlRAD21 <- "**" } else if (pvalRAD21 < 0.05) { siglvlRAD21 <- "*" } else { siglvlRAD21 <- "ns" }
if (pvalSA1 < 0.001) { siglvlSA1 <- "***" } else if (pvalSA1 < 0.01) { siglvlSA1 <- "**" } else if (pvalSA1 < 0.05) { siglvlSA1 <- "*" } else { siglvlSA1 <- "ns" }
if (pvalSADKO < 0.001) { siglvlSADKO <- "***" } else if (pvalSADKO < 0.01) { siglvlSADKO <- "**" } else if (pvalSADKO < 0.05) { siglvlSADKO <- "*" } else { siglvlSADKO <- "ns" }
#y coordinate of signifcance level symbol is calculated for SA2 and RAD21 group
maxval<-max(GOIdatM$value)
if (maxval > 0) {
siglvlSTAG2coord_y <- (maxval + 0.05*maxval)
siglvlSTAG1coord_y <- (siglvlSTAG2coord_y + 0.05*maxval)
siglvlSTAG12coord_y <- (siglvlSTAG1coord_y + 0.05*maxval)
siglvlRAD21coord_y <- (siglvlSTAG12coord_y + 0.05*maxval)}
else {
siglvlSTAG2coord_y <- 0.15
siglvlSTAG1coord_y <- 0.3
siglvlSTAG12coord_y <- 0.45
siglvlRAD21coord_y <- 0.6}
#pdf(file = file.path(FIGDIR,ANALn,"wBC","test.pdf"),height = 5,width = 7)
ggplot(GOIdatM,aes(x=group,y=value))+
geom_col(data=GOIdatM[1:5,],aes(x=group,y=value, fill=group),alpha=0.8)+
scale_fill_manual(values = c("CTRL" = "firebrick", "SA2" = "lightgreen","SA1" = "darkgoldenrod","SA1_SA2" = "steelblue3", "RAD21" = "mediumvioletred")) +
scale_x_discrete(labels=col.labs) +
geom_errorbar(data=GOIdatM[1:5,],aes(ymin=value[1:5]-sd[1:5], ymax=value[1:5]+sd[1:5]), width=.2,position=position_dodge(.9)) +
geom_jitter(data = GOIdatM[6:nrow(GOIdatM),], aes(x=group,y=value,fill=group,color=group),alpha=0.6,size=pointsize, width = 0.15)+
scale_color_manual(values = c("CTRL" = "lightsalmon2", "SA2" = "seagreen3","SA1" = "darkgoldenrod1","SA1_SA2" = "steelblue1", "RAD21" = "violet")) +
theme(
  legend.position=legloc,
  axis.text.x=element_text(angle=0,size=rel(2),face="bold"),
  axis.text.y=element_text(size=rel(3.5)),
  axis.title=element_text(size=rel(2),face="bold"),
  plot.title = element_text(size = rel(3.5), face = "bold",hjust=0.5),
  panel.border = element_blank(),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.background = element_blank(),
  axis.line = element_line(colour = "black"),) +
    coord_cartesian(ylim = c(y1,y2))+
    xlab(bottom) + ylab("log rpkm") +
    ggtitle(GOI) + 
annotate("text", x=1.5, y=(siglvlSTAG2coord_y+0.015*siglvlSTAG2coord_y), label= siglvlSA2, size=rel(8),fontface="bold") +
annotate("text", x=2, y=(siglvlSTAG1coord_y+0.015*siglvlSTAG1coord_y), label= siglvlSA1, size=rel(8),fontface="bold") +
annotate("text", x=2.5, y=(siglvlSTAG12coord_y+0.015*siglvlSTAG12coord_y), label= siglvlSADKO, size=rel(8),fontface="bold") +
annotate("text", x=3, y=(siglvlRAD21coord_y+0.015*siglvlRAD21coord_y), label= siglvlRAD21, size=rel(8),fontface="bold") +
annotate("segment", x = 1, xend = 2, y = siglvlSTAG2coord_y, yend = siglvlSTAG2coord_y)+
annotate("segment", x = 1, xend = 3, y = siglvlSTAG1coord_y, yend = siglvlSTAG1coord_y)+
annotate("segment", x = 1, xend = 4, y = siglvlSTAG12coord_y, yend = siglvlSTAG12coord_y)+
annotate("segment", x = 1, xend = 5, y = siglvlRAD21coord_y, yend = siglvlRAD21coord_y)
}
```


## plot expression of GOIs
```{r}
dir.create(file.path(FIGDIR,ANALn,"GEXbarplots"))
#all into one grid by type of gene set:
CohComp=c("STAG1","STAG2","RAD21","SMC1A","SMC3","NIPBL","WAPL","ESCO2","MAU2","PDS5A","SMC1B")
CohCompCore=c("STAG2","STAG1","RAD21","SMC1A","SMC3")

PlotsCoh_CD34 = lapply(CohCompCore, GENEX_func_means_dots, y1=-0,y2=7.1)
pdf(file = file.path(FIGDIR,ANALn,"GEXbarplots","Cohesin_Core_component_expr_meandots_rpkm.siglvl.pdf"),height = 10 ,width = 20)
grid.arrange(grobs=PlotsCoh_CD34,ncol=4,top="RNA expression of Cohesin Components")
dev.off()

#without column labs, all in one row, fixed limits
collablvecempty<-c(rep("",5))
PlotsCoh_CD34 = lapply(CohCompCore, GENEX_func_means_dots, y1=-0,y2=7.1,pointsize=rel(3),col.labs=collablvecempty)
pdf(file = file.path(FIGDIR,ANALn,"GEXbarplots","Cohesin_Core_component_expr_meandots_rpkm.siglvl.nolab.pdf"),height = 10,width = 35)
grid.arrange(grobs=PlotsCoh_CD34,ncol=5)
dev.off()

#differention makers of interest
##, ITGAM=CD11b, ITGA2B=CD41 
##HSC markers (PROM1=CD133, KIT=CD117)
diffmarkersHSC=c("KIT","CD34","PROM1","ABCG2","CD93","ETV6","STAT5","PTEN","GATA2","MCL1","GFI1","BMI1","HOXB4","THY1","MME","FLT3","CDCP1","CXCR4")
##myeloid markers (ANPEP=CD13, ITGAM=CD11b,FUT4=CD15,FCGR1A=CD64,FCGR3A=CD16,ITGB2=CD18, CEACAM8=CD66b,FCGR2A=CD32,PTPRC=CD45,SELL=CD62L)
diffmarkersMy=c("MPO","ANPEP","CD33","ITGAM","FUT4","FCGR1A","FCGR3A","FCGR2A","ITGB2","CEACAM8","PTPRC","CD44","SELL")
# Monocytic markers CD14, CD36, CD64, CD4, CD38, CD11c=ITGAX
diffmarkersMO=c("CD14","CD36","CD4","CD38","ITGAX")
#Megakaryocytic markers 	CD41 (glycoprotein IIb/IIIa=ITGA2B), CD61 (glycoprotein IIIa=ITGB3)
diffmarkersMEG=c("ITGA2B","ITGB3")
#Erythroid markers 	CD235a (glycophorin A = GYPA), CD71=TFRC, CD36 
diffmarkersERY=c("GYPA","TFRC")
diffmarkersall=c(diffmarkersHSC,diffmarkersMy,diffmarkersMO,diffmarkersMEG,diffmarkersERY)
diffmarkersall<-diffmarkersall[diffmarkersall %in% qstat_SA2vs.CTRL$GeneSymbol]
diffmarkersplots=lapply(diffmarkersall, GENEX_func_means_dots)
pdf(file = file.path(FIGDIR,ANALn,"GEXbarplots","Diffmarkers2_meandots_rpkm.siglvl.pdf"),height = 40,width = 20)
grid.arrange(grobs=diffmarkersplots,ncol=5,top="RNA expression of important differentiation markers")
dev.off()

##panel with sig. HBG and ery genes only with adapted scales
HBAplots=lapply(c("HBA1","HBA2","HBG1","HBG2","HBQ1","TFRC"), GENEX_func_means_dots,y1=-0.5,y2=10)
HBZplot=lapply(c("HBZ","GYPA"), GENEX_func_means_dots,y1=-3,y2=2.5)
sigHBplots<-c(HBAplots,HBZplot)
pdf(file = file.path(FIGDIR,ANALn,"GEXbarplots","sig.HGBgenes_meandots_rpkm.siglvl.pdf"),height = 10,width = 20)
grid.arrange(grobs=sigHBplots,ncol=5,top="RNA expression of sig. ery/hemoglobin genes")
dev.off()
##panel with sig. HSC and Myeloid genes only with adapted scales
topdiffHSCMY=lapply(c("CD34","CD14","CD36","FCGR1A","CEACAM8","PTPRC","CD44"), GENEX_func_means_dots,y1=-0,y2=7)
pdf(file = file.path(FIGDIR,ANALn,"GEXbarplots","sig.HSC_MYE_genes_meandots_rpkm.siglvl.pdf"),height = 10,width = 20)
grid.arrange(grobs=topdiffHSCMY,ncol=5,top="RNA expression of sig. HSC/myeloid genes")
dev.off()
##panel with sig.  Myeloid genes only with adapted scales
topdiffMY=lapply(c("CD36","FCGR1A","CEACAM8","CD1C"), GENEX_func_means_dots,y1=-1.5,y2=5)
pdf(file = file.path(FIGDIR,ANALn,"GEXbarplots","sig.MYE_genes_meandots_rpkm.siglvl.2.pdf"),height = 5,width = 20)
grid.arrange(grobs=topdiffMY,ncol=4)
dev.off()
```

# Gene set enrichment analyses: STAG2 mut AML signature genes
## Reading in qstat results for the signatures and filter for DEGs to test
```{r}
##Read qstat results / geneslists for the signatures to test
#SA2mut AML specific (SA2mut vs CTRL AMLs)
listSA2mut2<-read.table(file.path(ANALDIRAML,"STAG2pat_vs_CTRL/qstat_STAG2.vs.CTRL.glm.extended.txt"), header=T, sep="\t",row.names=1)
listRAD21mut2<-read.table(file.path(ANALDIRAML,"RAD21pat_vs_CTRL/qstat_RAD21.vs.CTRL.glm.txt"), header=T, sep="\t",row.names=1)
##subset for up and downregulated genes
#subset SA2mut AML
genelist_SA2mut_UP2 <- subset(listSA2mut2,(logFC > .585 & FDR < 0.05 & logCPM > 1)) #439   
genelist_SA2mut_DOWN2 <- subset(listSA2mut2,(logFC < -.585 & FDR < 0.05 & logCPM > 1)) #212
#subset RAD21mut AML (less stringent criteria)
genelist_RAD21mut_UP2 <- subset(listRAD21mut2,(logFC > .385 & FDR < 0.05)) #65  
genelist_RAD21mut_DOWN2 <- subset(listRAD21mut2,(logFC < -.385 & FDR < 0.05)) #12
```

## gene set enrichment using fry
### prepeare indices and contrasts
```{r}
#create indices for the fit
indSA2mut_UP2 <- rownames(f4) %in% rownames(genelist_SA2mut_UP2) #387
indSA2mut_DOWN2 <- rownames(f4) %in% rownames(genelist_SA2mut_DOWN2) #186
indSA2mut<-list(indSA2mut_UP2,indSA2mut_DOWN2)
names(indSA2mut) <- c("SA2mut_up","SA2mut_down")

indRAD21mut_UP2 <- rownames(f4) %in% rownames(genelist_RAD21mut_UP2) #59
indRAD21mut_DOWN2 <- rownames(f4) %in% rownames(genelist_RAD21mut_DOWN2) #8
indRAD21mut<-list(indRAD21mut_UP2,indRAD21mut_DOWN2)
names(indRAD21mut) <- c("RAD21mut_up","RAD21mut_down")

##list of all contrasts
conSA2 <- makeContrasts(SI_targetSA2 - SI_targetCTRL, levels=design_DEGs_target2)
conSA1 <- makeContrasts(SI_targetSA1 - SI_targetCTRL, levels=design_DEGs_target2)
conSA12 <- makeContrasts(SI_targetSA1_SA2 - SI_targetCTRL, levels=design_DEGs_target2)
conRAD21 <- makeContrasts(SI_targetRAD21 - SI_targetCTRL, levels=design_DEGs_target2)
contrastslist<-list(conSA2,conSA1,conSA12,conRAD21)
names(contrastslist)<-c("SA2","SA1","SA1_SA2","RAD21")
```

### fry all comparisons for AML signatures
```{r}
###fry and save in list
selfdef_sigs<-list(indSA2mut,indRAD21mut)
names(selfdef_sigs)<-c("SA2mut","RAD21mut")
GSEAlist_allKDs<-list()
for (sig in names(selfdef_sigs)){
for (i in names(contrastslist)){
GSEAlist_allKDs[[paste0(i,sig)]]<- fry(dgel, index=selfdef_sigs[[sig]], design=design_DEGs_target2, contrast=contrastslist[[i]])
}}

###summarize by signature
GSEA_res_allKDs<-list()
for (dir in c("up","down")){
for (sig in names(selfdef_sigs)){
df<-data.frame()
for (KD in names(contrastslist)){
df[KD,"PValue"]<-GSEAlist_allKDs[[paste0(KD,sig)]][paste0(sig,"_",dir),"PValue"]
df[KD,"FDR"]<-GSEAlist_allKDs[[paste0(KD,sig)]][paste0(sig,"_",dir),"FDR"]
df[KD,"NGenes"]<-GSEAlist_allKDs[[paste0(KD,sig)]][paste0(sig,"_",dir),"NGenes"]
df[KD,"Direction"]<-GSEAlist_allKDs[[paste0(KD,sig)]][paste0(sig,"_",dir),"Direction"]
}
GSEA_res_allKDs[[paste0(sig,"_",dir)]]<-df
}}
#look at enrichments
GSEA_res_allKDs[[paste0("SA2mut","_",dir)]]
#save list
saveRDS(GSEA_res_allKDs,file=file.path(FIGDIR,ANALn,"GSEA","GSEA_CohesinMutAML_signature_results.allKDs.rds"))
```

## barcodeplot to visualize enrichement of STAG2mut signature enrichent in the STAG2 KD
```{r}
#prepare input dataframes for fold-changes and GSEA stats to display
resSA2 <- glmQLFTest(f4, contrast=conSA2)
resSA1 <- glmQLFTest(f4, contrast=conSA1)
resSA12 <- glmQLFTest(f4, contrast=conSA12)
resRAD <- glmQLFTest(f4, contrast=conRAD21)

gseSA2mutinSA2KD<-GSEAlist_allKDs[[paste0("SA2","SA2mut")]]
gseSA2mutinSA1KD<-GSEAlist_allKDs[[paste0("SA1","SA2mut")]]

##plot results SA2AML signature in SA2 KD
pdf(file=file.path(FIGDIR,ANALn,"GSEA","GSEA_SA2mutDEGs_SA2KD.barcodeplot.pdf"),width=4,height=5)
barcodeplot(resSA2$table$logFC, 
            index=indSA2mut_UP2, 
            index2=indSA2mut_DOWN2, 
            labels=c("CTRL","STAG2KD"), 
            xlab = bquote(log[2]*"FC in RNAseq"), 
            main="", 
            col.bars=c("seagreen2", "firebrick1"))
par(new=TRUE)
plot.new( )
plot.window( xlim=c(-5,5), ylim=c(-5,5) )
text(3,5,bquote('green bars: genes upregulated in SA2mut AML '~'('*.(gseSA2mutinSA2KD["SA2mut_up","NGenes"])*')'), adj = c(1,.5),cex=0.5)
text(3,4.5,bquote(italic(P)[adj.]*"<"*.(gseSA2mutinSA2KD["SA2mut_up","FDR"])*'('*.(gseSA2mutinSA2KD["SA2mut_up","Direction"])*')'), adj = c(1,.5),cex=1)
text(-4,-5,bquote('red bars: genes downregulated in SA2mut AML  '~'('*.(gseSA2mutinSA2KD["SA2mut_down","NGenes"])*')'), adj = c(0,.5),cex=0.5)
text(-4,-4.5,bquote(italic(P)[adj.]*"<"*.(gseSA2mutinSA2KD["SA2mut_down","FDR"])*'('*.(gseSA2mutinSA2KD["SA2mut_down","Direction"])*')'), adj = c(0,.5),cex=1)
dev.off()


###same for SA1 KD 
pdf(file=file.path(FIGDIR,ANALn,"GSEA","GSEA_SA2mutDEGs_SA1KD.barcodeplot.pdf"),width=4,height=5)
barcodeplot(resSA1$table$logFC, 
            index=indSA2mut_UP2, 
            index2=indSA2mut_DOWN2, 
            labels=c("CTRL","STAG1KD"), 
            xlab = bquote(log[2]*"FC in RNAseq"), 
            main="", 
            col.bars=c("darkgoldenrod3", "firebrick1"))
par(new=TRUE)
plot.new( )
plot.window( xlim=c(-5,5), ylim=c(-5,5) )
text(3,5,bquote('green bars: genes upregulated in SA2mut AML '~'('*.(gseSA2mutinSA1KD["SA2mut_up","NGenes"])*')'), adj = c(1,.5),cex=0.5)
text(3,4.5,bquote(italic(P)[adj.]*"<"*.(gseSA2mutinSA1KD["SA2mut_up","FDR"])*'('*.(gseSA2mutinSA1KD["SA2mut_up","Direction"])*')'), adj = c(1,.5),cex=1)
text(-4,-5,bquote('red bars: genes downregulated in SA2mut AML  '~'('*.(gseSA2mutinSA1KD["SA2mut_down","NGenes"])*')'), adj = c(0,.5),cex=0.5)
text(-4,-4.5,bquote(italic(P)[adj.]*"<"*.(gseSA2mutinSA1KD["SA2mut_down","FDR"])*'('*.(gseSA2mutinSA1KD["SA2mut_down","Direction"])*')'), adj = c(0,.5),cex=1)
dev.off()
```


# GO anlysis/GSEA using camera and the msigdb database
## get signatures of interest, index and calculate GSEA in all contrasts
```{r}
library(msigdbr)

#get genesets/GO terms form msigdb database
Cameralist<-vector(mode = "list", length = 0)
signatures<-c("H","C2","C3","C5","C6","C7","C8")
for (sig in signatures){
Cameralist[[sig]] <- msigdbr(species = "Homo sapiens", category = sig)
#transform into list format required for indexing
Cameralist[[sig]] <- split(x = Cameralist[[sig]]$gene_symbol, f = Cameralist[[sig]]$gs_name)
#indexing in dgelist object
Cameralist[[paste0("index_",sig)]] <- ids2indices(Cameralist[[sig]],id=dgel$genes$GeneSymbol)
}

#GSEA using camera and all indices
for (cond in names(contrastslist)){
for (sig in signatures){
Cameralist[[paste0(cond,"_camera_index_",sig)]] <- camera(dgel, index = Cameralist[[paste0("index_",sig)]],design = design_DEGs_target2,contrast=contrastslist[[cond]])
}}

#show excemplary results: C2 signatures in  STAG2 KD 
head(Cameralist[["SA2_camera_index_C2"]],10)
```

## look at individual interesting signatures of interest: BROWN/JAATINEN
```{r}
###C2 signatures
reslist2<-list(resSA2,resSA1,resSA12,resRAD)
names(reslist2)<-names(contrastslist)
dir.create(file.path(FIGDIR,ANALn,"GSEA","msigdb.genesig.enrichment"))
##For two-way C2 signatures:  JAATINEN / RAMALHO
twowaysigs<-c("JAATINEN_HEMATOPOIETIC_STEM_CELL","BROWN_MYELOID_CELL_DEVELOPMENT")

#plot signatures in all KD vs CTRL contrasts
for (KD in names(reslist2)){
  for (sig in twowaysigs){
pdf(file=file.path(FIGDIR,ANALn,"GSEA","msigdb.genesig.enrichment",paste0("GSEA.",sig,".",KD,"KD.barcodeplot.pdf")), height=6, width=8)
barcodeplot(reslist2[[KD]]$table$logFC,
            index=as.integer(unlist(Cameralist[["index_C2"]][paste0(sig,"_UP")])), 
            index2=as.integer(unlist(Cameralist[["index_C2"]][paste0(sig,"_DN")])), 
            labels=c("CTRL",paste0(KD," KD")), 
            xlab = bquote(log[2]*"FC in RNAseq"),
            main=paste0("GSEA: ",sig, " in \n", KD, " KD vs CTRL-HSPCs"),
            col.bars=c("lightsalmon3", "darkblue"))
par(new=TRUE)
plot.new( )
plot.window( xlim=c(-5,5), ylim=c(-5,5) )
text(3,5,bquote(''*.(sig)*  ~' UP ' ~'('*.(Cameralist[[paste0(KD,"_camera_index_C2")]][paste0(sig,"_UP"),"NGenes"])*')'), adj = c(1,.5))
text(3,4.5,bquote(italic(P)[adj.]*"<"*.(Cameralist[[paste0(KD,"_camera_index_C2")]][paste0(sig,"_UP"),"FDR"])*'('*.(Cameralist[[paste0(KD,"_camera_index_C2")]][paste0(sig,"_UP"),"Direction"])*')'), adj = c(1,.5),cex=1.5)
text(-4,-5,bquote(''*.(sig)* ~' DOWN ' ~'('*.(Cameralist[[paste0(KD,"_camera_index_C2")]][paste0(sig,"_DN"),"NGenes"])*')'), adj = c(0,.5))
text(-4,-4.5,bquote(italic(P)[adj.]*"<"*.(Cameralist[[paste0(KD,"_camera_index_C2")]][paste0(sig,"_DN"),"FDR"])*'('*.(Cameralist[[paste0(KD,"_camera_index_C2")]][paste0(sig,"_DN"),"Direction"])*')'), adj = c(0,.5),cex=1.5)
dev.off()
}}
```