---
title: "RAD21 enrichment at diffloop anchors and STAG1/2 dominant positions in CD34 cohesin KD"
author: "Alexander Fischer"
date: "03 2022"
output: html_document
---

#rbioc_3-12
# Load libraries set paths
```{r setup, include=FALSE}
DIR_DATA="/misc/data"
#directories for HIC associated RAD21 peaks
INPUTDIR=file.path(DIR_DATA,"analysis/project_cohesin/CD34/ChIP_KD_analysis/Cohesin_CTCF_MED12")
WORKDIR_CD34=file.path(DIR_DATA,"analysis/project_cohesin/CD34/HiC_KDs/DeepSeq_Analysis")
FIGURESDIR=file.path(WORKDIR_CD34,"figures")

LOOPDIR=file.path(WORKDIR_CD34,"loops","withoutXY")
diffanchorsn2t=file.path(LOOPDIR,"differentialanchorsn2t")
overlapsplitdir=file.path(diffanchorsn2t,"splitbyoverlap")

##loop subtypes 
EPpairs=file.path(diffanchorsn2t,"AnchorsOverlapPairs/TSSoverlap/RAD21centred") #regulatoy
STRUCANCHS=file.path(diffanchorsn2t,"ChIPoverlap/structural_Loops") #structural

#directories for SA1/2 dominant site associated peaks
WORDIR_CD34_2=file.path(INPUTDIR,"diffPeaks/SA1vsSA2")
FIGURESDIR_dom=file.path(INPUTDIR,"figures/GSEA")

#libraries
library(ggplot2)
library(edgeR)
library(dplyr)
library(cowplot)

#variables
KDs<-c("SA2KD","SA1KD")
Filts<-c("up.FC1", "down.FC1")
```
# read in data and set variables
```{r}
#read in data
data <- read.delim(file.path(INPUTDIR,"CD34_RAD21.peaks.ann.Rinput.txt"), row.names="ID")
#set variables
group <- factor(c(rep("CTRL",10), rep("SA1KD",5), rep("SA2KD",6), rep("RAD21KD",4)))
batch <- factor(c(14,14,17,18,19,21,22,24,27,28,14,17,21,27,28,14,17,20,21,22,28,18,22,27,28))
```

# edge R fit
```{r}
d <- DGEList(counts=data,group=group)
d <- calcNormFactors(d)
design <- model.matrix(~0 + group + batch)
rownames(design) <- colnames(d)
d <- estimateDisp(d, design, robust=TRUE)
d$common.dispersion
fit <- glmQLFit(d, design)
#contrasts
conSA1 <- makeContrasts(groupSA1KD - groupCTRL, levels=design)
conSA2 <- makeContrasts(groupSA2KD - groupCTRL, levels=design)
#resulttables
res1 <- glmQLFTest(fit, contrast=conSA1)
res2 <- glmQLFTest(fit, contrast=conSA2)
```

# RAD21 Peak set Enrichment analyses in STAG1 and STAG2 KD HSPCs vs CTRL HSPCs

## PSEA of SA1/2 dominant site associated RAD21 peaks in STAG KDs
```{r}
#read in RAD21 peak names with STAG1 or STAG2 domination
SA12dom<-vector(mode = "list")
SA12dom[["SA1dom"]]<-row.names(read.delim(file.path(WORDIR_CD34_2,"SA2vsSA1.allCD34.RAD21.peaks_edgeR.2folddown.txt"), header=T, sep="\t",row.names=1))
SA12dom[["SA2dom"]]<-row.names(read.delim(file.path(WORDIR_CD34_2,"SA2vsSA1.allCD34.RAD21.peaks_edgeR.2foldup.txt"), header=T, sep="\t",row.names=1))
SA12dom[["common"]]<-row.names(read.delim(file.path(WORDIR_CD34_2,"SA2vsSA1.allCD34.RAD21.peaks_edgeR.common.txt"), header=T, sep="\t",row.names=1))

#set index
cats<-c("SA1dom","SA2dom","common")
for (cat in cats) {
SA12dom[[paste0("ind_",cat)]]<- rownames(fit) %in% SA12dom[[cat]]
}
#index list for complete set
SA12dom[[paste0("ind_all")]]<- list(SA12dom[["ind_SA1dom"]],SA12dom[["ind_SA2dom"]],SA12dom[["ind_common"]])
names(SA12dom[[paste0("ind_all")]])<-cats

#fry for all three types at once in STAG2 or STAG1 KD contrast
##STAG2 KD
fr_SA2indDOM <- fry(d, index=SA12dom[[paste0("ind_all")]], design=design, contrast=conSA2)
fr_SA2indDOM
write.table(fr_SA2indDOM,file=file.path(FIGURESDIR_dom,"PSEA.RAD21.at.STAGpreference.pos.SA2KDvsCTRL.txt"),sep="\t",quote=FALSE)
##STAG1 KD
fr_SA1indDOM <- fry(d, index=SA12dom[[paste0("ind_all")]], design=design, contrast=conSA1)
fr_SA1indDOM
write.table(fr_SA1indDOM,file=file.path(FIGURESDIR_dom,"PSEA.RAD21.at.STAGpreference.pos.SA1KDvsCTRL.txt"),sep="\t",quote=FALSE)



#barcode plots
pdf(file=file.path(FIGURESDIR_dom,"edgeR_barcodeplot_RAD21.at.STAGpreference.pos.SA1KDvsCTRL.pdf"), height=6, width=6)
barcodeplot(res1$table$logFC, 
            index=SA12dom[["ind_SA2dom"]] ,
            index2=SA12dom[["ind_SA1dom"]],
            labels=c("CTRL-HSPCs","STA1GKD"),
            xlab = bquote(log[2]*"FC in RAD21 ChIPseq"),
            cex.axis = 0.5, 
            main=paste0("RAD21 at STAG dominant positions"), cex.main=0.5,
            col.bars=c("#12cb65","#835f04"))
par(new=TRUE)
plot.new( )
plot.window( xlim=c(-5,5), ylim=c(-5,5) )
text(5,5,bquote('green bars: RAD21 peaks at SA2 dominant positions  '~'('*.(fr_SA1indDOM["SA2dom","NGenes"])*')'), adj = c(1,.3),cex=1)
text(5,4.5,bquote(italic(P)[adj.]*"<"*.(fr_SA1indDOM["SA2dom","FDR"])*'('*.(fr_SA1indDOM["SA2dom","Direction"])*')'), adj = c(1,.5),cex=1.5)
text(-5,-5,bquote('golden bars: RAD21 peaks at SA1 dominant positions  '~'('*.(fr_SA1indDOM["SA1dom","NGenes"])*')'), adj = c(0,.5),cex=1)
text(-5,-4.5,bquote(italic(P)[adj.]*"<"*.(fr_SA1indDOM["SA1dom","FDR"])*'('*.(fr_SA1indDOM["SA1dom","Direction"])*')'), adj = c(0,.5),cex=1.5)
dev.off()

pdf(file=file.path(FIGURESDIR_dom,"edgeR_barcodeplot_RAD21.at.STAGpreference.pos.SA2KDvsCTRL.pdf"), height=6, width=6)
barcodeplot(res2$table$logFC, 
            index=SA12dom[["ind_SA2dom"]] ,
            index2=SA12dom[["ind_SA1dom"]],
            labels=c("CTRL-HSPCs","STA2GKD"),
            xlab = bquote(log[2]*"FC in RAD21 ChIPseq"),
            cex.axis = 0.5, 
            main=paste0("RAD21 at STAG dominant positions"), cex.main=0.5,
            col.bars=c("#12cb65","#835f04"))
par(new=TRUE)
plot.new( )
plot.window( xlim=c(-5,5), ylim=c(-5,5) )
text(5,5,bquote('green bars: RAD21 peaks at SA2 dominant positions  '~'('*.(fr_SA2indDOM["SA2dom","NGenes"])*')'), adj = c(1,.3),cex=1)
text(5,4.5,bquote(italic(P)[adj.]*"<"*.(fr_SA2indDOM["SA2dom","FDR"])*'('*.(fr_SA2indDOM["SA2dom","Direction"])*')'), adj = c(1,.5),cex=1.5)
text(-5,-5,bquote('golden bars: RAD21 peaks at SA1 dominant positions  '~'('*.(fr_SA2indDOM["SA1dom","NGenes"])*')'), adj = c(0,.5),cex=1)
text(-5,-4.5,bquote(italic(P)[adj.]*"<"*.(fr_SA2indDOM["SA1dom","FDR"])*'('*.(fr_SA2indDOM["SA1dom","Direction"])*')'), adj = c(0,.5),cex=1.5)
dev.off()
```

## PSEA of SA1/2 dominant site associated RAD21 peaks divided by CTCF status in STAG KDs
```{r}
#read in RAD21 peak names with STAG1 or STAG2 domination subset by CTCF overlap
###using same peaksets as above but separated by CTCF###
SA12domBYCTCF<-vector(mode = "list")
SA12domBYCTCF[["SA1dom_noCTCF"]]<-row.names(read.delim(file.path(WORDIR_CD34_2,"subsetByCTCF","SA1dom_noCTCF.txt"), header=T, sep="\t",row.names=1))
SA12domBYCTCF[["SA1dom_withCTCF"]]<-row.names(read.delim(file.path(WORDIR_CD34_2,"subsetByCTCF","SA1dom_withCTCF.txt"), header=T, sep="\t",row.names=1))
SA12domBYCTCF[["SA2dom_noCTCF"]]<-row.names(read.delim(file.path(WORDIR_CD34_2,"subsetByCTCF","SA2dom_noCTCF.txt"), header=T, sep="\t",row.names=1))
SA12domBYCTCF[["SA2dom_withCTCF"]]<-row.names(read.delim(file.path(WORDIR_CD34_2,"subsetByCTCF","SA2dom_withCTCF.txt"), header=T, sep="\t",row.names=1))
SA12domBYCTCF[["common_noCTCF"]]<-row.names(read.delim(file.path(WORDIR_CD34_2,"subsetByCTCF","common_noCTCF.txt"), header=T, sep="\t",row.names=1))
SA12domBYCTCF[["common_withCTCF"]]<-row.names(read.delim(file.path(WORDIR_CD34_2,"subsetByCTCF","common_withCTCF.txt"), header=T, sep="\t",row.names=1))


#set index
cats<-c("SA1dom_noCTCF","SA1dom_withCTCF","SA2dom_noCTCF","SA2dom_withCTCF","common_noCTCF","common_withCTCF")
for (cat in cats) {
SA12domBYCTCF[[paste0("ind_",cat)]]<- rownames(fit) %in% SA12domBYCTCF[[cat]]
}
#index list for complete set
SA12domBYCTCF[[paste0("ind_all")]]<- list(SA12domBYCTCF[["ind_SA1dom_noCTCF"]],SA12domBYCTCF[["ind_SA1dom_withCTCF"]],SA12domBYCTCF[["ind_SA2dom_noCTCF"]],SA12domBYCTCF[["ind_SA2dom_withCTCF"]],SA12domBYCTCF[["ind_common_noCTCF"]],SA12domBYCTCF[["ind_common_withCTCF"]])
names(SA12domBYCTCF[[paste0("ind_all")]])<-cats

#fry for all three types at once in STAG2 or STAG1 KD contrast
fr_SA2indDOM_byCTCF <- fry(d, index=SA12domBYCTCF[[paste0("ind_all")]], design=design, contrast=conSA2)
fr_SA2indDOM_byCTCF
#                NGenes Direction       PValue          FDR PValue.Mixed
#SA1dom_withCTCF    926        Up 5.428555e-09 3.257133e-08 2.343432e-11
#SA1dom_noCTCF     1128      Down 2.251957e-04 6.755870e-04 2.228318e-06
#common_noCTCF    27244      Down 6.629236e-04 1.325847e-03 3.726199e-07
#SA2dom_noCTCF      349      Down 9.272551e-04 1.390883e-03 3.259361e-06
#common_withCTCF  48246        Up 1.416997e-03 1.700396e-03 6.326759e-10
#SA2dom_withCTCF    968      Down 4.889823e-03 4.889823e-03 4.617335e-07

fr_SA1indDOM_byCTCF <- fry(d, index=SA12domBYCTCF[[paste0("ind_all")]], design=design, contrast=conSA1)
fr_SA1indDOM_byCTCF
#                NGenes Direction      PValue         FDR PValue.Mixed
#SA1dom_withCTCF    926      Down 0.001583374 0.009500246 4.723769e-05
#SA2dom_withCTCF    968        Up 0.005218957 0.015656871 1.482349e-03
#common_withCTCF  48246        Up 0.130130544 0.214611527 1.444558e-02
#common_noCTCF    27244      Down 0.183670846 0.214611527 6.162534e-03
#SA2dom_noCTCF      349      Down 0.191049195 0.214611527 2.056545e-01
#SA1dom_noCTCF     1128      Down 0.214611527 0.214611527 1.567535e-02

dir.create(file.path(FIGURESDIR_dom,"byCTCFsubset"))
#barcode plots showing STAG1 vs STAG2 dominant pos divided by CTCF subset in STAG1 or STAG2 KD contrasts
CTCFcats=c("noCTCF","withCTCF")
KDs=c("STAG1KD","STAG2KD")
reslist<-list(res1,res2)
names(reslist)<-KDs
frylist<-list(fr_SA1indDOM_byCTCF,fr_SA2indDOM_byCTCF)
names(frylist)<-KDs
for (CTCF in CTCFcats){
  for (KD in KDs){
pdf(file=file.path(FIGURESDIR_dom,"byCTCFsubset",paste0("edgeR_barcodeplot_RAD21",CTCF,".STAGdom.",KD,".pdf")), height=6, width=6)
barcodeplot(reslist[[KD]]$table$logFC, 
            index=SA12domBYCTCF[[paste0("ind_SA2dom_",CTCF)]] ,
            index2=SA12domBYCTCF[[paste0("ind_SA1dom_",CTCF)]],
            labels=c("CTRL-HSPCs",KD),
            xlab = bquote(log[2]*"FC in RAD21 ChIPseq"),
            cex.axis = 0.5, 
            main=paste0(KD, ": RAD21 at STAG dominant positions ", CTCF), cex.main=0.5,
            col.bars=c("#12cb65","#835f04"))
par(new=TRUE)
plot.new( )
plot.window( xlim=c(-5,5), ylim=c(-5,5) )
text(5,5,bquote('green bars: RAD21 peaks at SA2 dominant positions  '~'('*.(frylist[[KD]][paste0("SA2dom_",CTCF),"NGenes"])*')'), adj = c(1,.3),cex=1)
text(5,4.5,bquote(italic(P)[adj.]*"<"*.(frylist[[KD]][paste0("SA2dom_",CTCF),"FDR"])*'('*.(frylist[[KD]][paste0("SA2dom_",CTCF),"Direction"])*')'), adj = c(1,.5),cex=1.5)
text(-5,-5,bquote('golden bars: RAD21 peaks at SA1 dominant positions  '~'('*.(frylist[[KD]][paste0("SA1dom_",CTCF),"NGenes"])*')'), adj = c(0,.5),cex=1)
text(-5,-4.5,bquote(italic(P)[adj.]*"<"*.(frylist[[KD]][paste0("SA1dom_",CTCF),"FDR"])*'('*.(frylist[[KD]][paste0("SA1dom_",CTCF),"Direction"])*')'), adj = c(0,.5),cex=1.5)
dev.off()
}}
#barcode plots showing common pos divided by CTCF subset in STAG1 or STAG2 KD contrasts
for (KD in KDs){
pdf(file=file.path(FIGURESDIR_dom,"byCTCFsubset",paste0("edgeR_barcodeplot_RAD21.common.byCTCFstatus.",KD,".pdf")), height=6, width=6)
barcodeplot(reslist[[KD]]$table$logFC, 
            index=SA12domBYCTCF[[paste0("ind_common_withCTCF")]] ,
            index2=SA12domBYCTCF[[paste0("ind_common_noCTCF")]],
            labels=c("CTRL-HSPCs",KD),
            xlab = bquote(log[2]*"FC in RAD21 ChIPseq"),
            cex.axis = 0.5, 
            main=paste0(KD, ": RAD21 at STAG1/2 common positions "), cex.main=0.5,
            col.bars=c("#3737dfc0","#313131"))
par(new=TRUE)
plot.new( )
plot.window( xlim=c(-5,5), ylim=c(-5,5) )
text(5,5,bquote('RAD21 peaks at common sites with CTCF '~'('*.(frylist[[KD]]["common_withCTCF","NGenes"])*')'), adj = c(1,.3),cex=1)
text(5,4.5,bquote(italic(P)[adj.]*"<"*.(frylist[[KD]]["common_withCTCF","FDR"])*'('*.(frylist[[KD]]["common_withCTCF","Direction"])*')'), adj = c(1,.5),cex=1.5)
text(-5,-5,bquote('RAD21 peaks at common sites without CTCF  '~'('*.(frylist[[KD]]["common_noCTCF","NGenes"])*')'), adj = c(0,.5),cex=1)
text(-5,-4.5,bquote(italic(P)[adj.]*"<"*.(frylist[[KD]]["common_noCTCF","FDR"])*'('*.(frylist[[KD]]["common_noCTCF","Direction"])*')'), adj = c(0,.5),cex=1.5)
dev.off()
}
```

### summarize Padj values in barplot for STAG dominant sites by CTCF status
```{r}
##negative log FDR (positive value)
fr_SA2indDOM_byCTCF$logFDR<--log(fr_SA2indDOM_byCTCF$FDR)
fr_SA1indDOM_byCTCF$logFDR<--log(fr_SA1indDOM_byCTCF$FDR)
#for down direction use log FDR (negativ value)
fr_SA2indDOM_byCTCF$logFDR[fr_SA2indDOM_byCTCF$Direction == "Down"] <- log(subset(fr_SA2indDOM_byCTCF,Direction=="Down")[,"FDR"])
fr_SA1indDOM_byCTCF$logFDR[fr_SA1indDOM_byCTCF$Direction == "Down"] <- log(subset(fr_SA1indDOM_byCTCF,Direction=="Down")[,"FDR"])
#define categories
fr_SA2indDOM_byCTCF$STAGcat<-strsplit2(rownames(fr_SA2indDOM_byCTCF),"_",fixed = TRUE)[,1]
fr_SA2indDOM_byCTCF$CTCFcat<-strsplit2(rownames(fr_SA2indDOM_byCTCF),"_",fixed = TRUE)[,2]
fr_SA1indDOM_byCTCF$STAGcat<-strsplit2(rownames(fr_SA1indDOM_byCTCF),"_",fixed = TRUE)[,1]
fr_SA1indDOM_byCTCF$CTCFcat<-strsplit2(rownames(fr_SA1indDOM_byCTCF),"_",fixed = TRUE)[,2]
#set levels
fr_SA2indDOM_byCTCF$STAGcat<-factor(fr_SA2indDOM_byCTCF$STAGcat,levels=c("SA2dom","SA1dom","common"))
fr_SA2indDOM_byCTCF$CTCFcat<-factor(fr_SA2indDOM_byCTCF$CTCFcat,levels=c("withCTCF","noCTCF"))
fr_SA1indDOM_byCTCF$STAGcat<-factor(fr_SA1indDOM_byCTCF$STAGcat,levels=c("SA2dom","SA1dom","common"))
fr_SA1indDOM_byCTCF$CTCFcat<-factor(fr_SA1indDOM_byCTCF$CTCFcat,levels=c("withCTCF","noCTCF"))


##plot for STAG2KD
#
p1<-ggplot(fr_SA2indDOM_byCTCF, aes(x= logFDR, y= rownames(fr_SA2indDOM_byCTCF),fill = STAGcat)) +
geom_bar(aes(colour=CTCFcat),size=1.5, width=0.8, position = position_dodge(width=0.5), stat="identity") +
  scale_fill_manual(values = c("SA2dom"="#12cb65","SA1dom"="darkgoldenrod3","common"="grey"))+
  scale_color_manual(values = c("noCTCF"="firebrick1","withCTCF"="darkblue"))+
  scale_x_continuous(limits = c(-18,18),labels=abs,breaks = seq(-18,18,by=6)) + 
  xlab("- log2 FDR of RAD21 PSEA") + ylab(" ") + 
  ggtitle("STAG2KD vs CTRL") +
  theme(
  axis.text.x = element_text(angle=0,vjust=0.2,size=16),
  legend.text = element_text(angle=0,vjust=0.2,size=16),
  axis.text.y = element_text(angle=0,hjust=0.95,vjust=0.2,size=0),
  legend.position="none", legend.title=element_blank(),
  axis.title=element_text(size=16,face="bold"), plot.title = element_text(size = 20, face = "bold",hjust = 0.5), 
  panel.border = element_blank(),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.background = element_blank(),
  axis.line.y = element_blank(),
  axis.ticks.y = element_blank(),
  axis.line.x = element_line(colour = "black"),)+
geom_vline(xintercept = -log2(0.05),linetype = 'dashed') + 
geom_vline(xintercept = log2(0.05),linetype = 'dashed',color="black")+
geom_text(aes(x=-12,y="common_withCTCF",label="Down"),size=6,vjust=-0.1,fontface = "bold")+
geom_text(aes(x=12,y="common_withCTCF",label="Up"),size=6,vjust=-0.1,fontface = "bold")

pdf(file=file.path(FIGURESDIR_dom,"byCTCFsubset",paste0("RAD21.PSEA.summary.byCTCF.STAG2KD.pdf")), height=3, width=10)
plot(p1)
dev.off()

#if no y labels desired use:
#axis.text.y = element_blank(), axis.line.y= element_blank(),axis.ticks.y= element_blank()

##plot for STAG1KD

p2<-ggplot(fr_SA1indDOM_byCTCF, aes(x= logFDR, y= rownames(fr_SA1indDOM_byCTCF),fill = STAGcat)) +
geom_bar(aes(colour=CTCFcat), size=1.5,width=0.8, position = position_dodge(width=0.5), stat="identity") +
  scale_fill_manual(values = c("SA2dom"="#12cb65","SA1dom"="darkgoldenrod3","common"="grey"))+
  scale_color_manual(values = c("noCTCF"="firebrick1","withCTCF"="darkblue"))+
  scale_x_continuous(limits = c(-18,18),labels=abs,breaks = seq(-18,18,by=6)) + 
  xlab("- log2 FDR of RAD21 PSEA") + ylab(" ") + 
  ggtitle("STAG1KD vs CTRL") +
  theme(
  axis.text.x = element_text(angle=0,vjust=0.2,size=16),
  axis.text.y = element_text(angle=0,hjust=0.95,vjust=0.2,size=0),
  legend.text = element_text(angle=0,vjust=0.2,size=16),
  legend.position="none", legend.title=element_blank(),
  axis.title=element_text(size=16,face="bold"), plot.title = element_text(size = 20, face = "bold",hjust = 0.5), 
  panel.border = element_blank(),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.background = element_blank(),
  axis.line.y = element_blank(),
  axis.ticks.y = element_blank(),
  axis.line.x = element_line(colour = "black"),)+
geom_vline(xintercept = -log2(0.05),linetype = 'dashed') + 
geom_vline(xintercept = log2(0.05),linetype = 'dashed',color="black")+
geom_text(aes(x=-12,y="common_withCTCF",label="Down"),size=6,vjust=-0.1,fontface = "bold")+
geom_text(aes(x=12,y="common_withCTCF",label="Up"),size=6,vjust=-0.1,fontface = "bold")

pdf(file=file.path(FIGURESDIR_dom,"byCTCFsubset",paste0("RAD21.PSEA.summary.byCTCF.STAG1KD.pdf")), height=3, width=10)
plot(p2)
dev.off()

#both combined
##extract legend
legend_b <- get_legend(
  p1 + 
    guides(color = guide_legend(nrow = 1)) +
    theme(legend.position = "bottom")
)

##plot both plots in grid and add one legend
pgrid <- plot_grid(p1, p2, ncol = 2)
pdf(file=file.path(FIGURESDIR_dom,"byCTCFsubset",paste0("RAD21.PSEA.summary.byCTCF.STAG1_2_KD.pdf")), height=4, width=12)
plot_grid(pgrid,legend_b,ncol=1,rel_heights=c(1.2,0.2))
dev.off()
```

# PSEA of RAD21 peaks at differential loop subtypes in STAG KDs (- without XY chromosomes; based on DESEQ2 n2t analyis)
### all structural-anchor associated RAD21 peaks
```{r}
Filts<-c("up.FC1", "down.FC1")
KDs<-c("SA2KD","SA1KD")

conslist<-list(SA2KD=conSA2,SA1KD=conSA1)
colorlist<-list(SA2KD="seagreen2",SA1KD="darkgoldenrod")
reslist2<-list(SA1KD=res1$table$logFC,SA2KD=res2$table$logFC)




RAD21strucAnch<-list()
for (KD in KDs) {
  for (FILT in Filts){
path<-file.path(STRUCANCHS,paste0(KD,"vsCTRL.loopAnchors.red.", FILT, ".structural.RAD21.centred.bed"))
##bedfiles
RAD21strucAnch[[paste0(FILT,"_",KD)]]<-read.table(path, header=F, sep="\t")
##indices
RAD21strucAnch[[paste0("ind_struct_",KD,FILT)]]<-rownames(fit) %in% RAD21strucAnch[[paste0(FILT,"_",KD)]]$V4
  }
}

##combined struc. Anchs. up + down
for (KD in KDs) {
#feature names
RAD21strucAnch[[paste0("struct.all.",KD)]]<-list(RAD21strucAnch[[paste0("up.FC1_",KD)]]$V4,RAD21strucAnch[[paste0("down.FC1_",KD)]]$V4)
names(RAD21strucAnch[[paste0("struct.all.",KD)]])<-c("up.FC1_structuralLoops","down.FC1_structuralLoops")
#indices
RAD21strucAnch[[paste0("ind_struct.all.",KD)]]<-list(RAD21strucAnch[[paste0("ind_struct_",KD,"up.FC1")]],RAD21strucAnch[[paste0("ind_struct_",KD,"down.FC1")]])
names(RAD21strucAnch[[paste0("ind_struct.all.",KD)]])<-c("up.FC1_structuralLoops","down.FC1_structuralLoops")
}

PSEAlist<-list()
for (KD in KDs) {
  PSEAlist[[paste0(KD,"fr_struct.all")]] <- fry(d, index=RAD21strucAnch[[paste0("ind_struct.all.",KD)]], design=design, contrast=conslist[[KD]])
}
PSEAlist
#$SA2KDfr_struct.all
#                         NGenes Direction       PValue          FDR
#up.FC1_structuralLoops     1042        Up 3.641680e-07 7.283361e-07
#down.FC1_structuralLoops    447        Up 1.372823e-03 1.372823e-03
#$SA1KDfr_struct.all
#                         NGenes Direction    PValue      FDR PValue.Mixed
#down.FC1_structuralLoops    967      Down 0.8811266 0.973456  0.026239765
#up.FC1_structuralLoops       69        Up 0.9734560 0.973456  0.007355809

dir.create(file.path(FIGURESDIR,"GSEA","diffStructloops"))
#barcodeplots fry results: structural loops
for (KD in KDs) {
type="structuralLoops"
text1<-paste0(colorlist[[KD]],' bars: RAD21 peaks in ',type,' at strengthened loop anchors ')
text2<-paste0('red bars: RAD21 peaks in ',type,' at weakended loop anchors ')
pdf(file=file.path(FIGURESDIR,"GSEA","diffStructloops",paste0("edgeR.PSEA.",KD,".RAD21peaks.diffloop.FC1.barcodeplot.",type,".pdf")), height=6, width=6)
barcodeplot(reslist2[[KD]], 
            index=RAD21strucAnch[[paste0("ind_struct_",KD,"up.FC1")]] ,
            index2=RAD21strucAnch[[paste0("ind_struct_",KD,"down.FC1")]],
            labels=c("CTRL",KD),
            xlab = bquote(log[2]*"FC in RAD21 ChIPseq"),
            cex.axis = 0.5, 
            main=paste0("RAD21 at differential structural loop anchors"), cex.main=0.5,
            col.bars=c(colorlist[[KD]],"firebrick1"))
par(new=TRUE)
plot.new( )
plot.window( xlim=c(-5,5), ylim=c(-5,5) )
text(5,5,bquote(.(text1)~'('*.(PSEAlist[[paste0(KD,"fr_struct.all")]][paste0("up.FC1_",type),"NGenes"])*')'), adj = c(1,.3),cex=1)
text(5,4.5,bquote(italic(P)[adj.]*"<"*.(PSEAlist[[paste0(KD,"fr_struct.all")]][paste0("up.FC1_",type),"FDR"])*'('*.(PSEAlist[[paste0(KD,"fr_struct.all")]][paste0("up.FC1_",type),"Direction"])*')'), adj = c(1,.5),cex=1.5)
text(-5,-5,bquote(.(text2)~'('*.(PSEAlist[[paste0(KD,"fr_struct.all")]][paste0("down.FC1_",type),"NGenes"])*')'), adj = c(0,.5),cex=1)
text(-5,-4.5,bquote(italic(P)[adj.]*"<"*.(PSEAlist[[paste0(KD,"fr_struct.all")]][paste0("down.FC1_",type),"FDR"])*'('*.(PSEAlist[[paste0(KD,"fr_struct.all")]][paste0("down.FC1_",type),"Direction"])*')'), adj = c(0,.5),cex=1.5)
dev.off()
}
```


### all diff. regulatory loop anchor RAD21 peaks without XY chrom
```{r}
RAD21EPanch2<-list()

for (KD in KDs) {
  for (FILT in Filts){
path<-file.path(diffanchorsn2t,"H3K27overlap",paste0(KD,"vsCTRL.loopAnchors.red.", FILT, ".H3K27accentred.RAD21centred.bed"))
RAD21EPanch2[[paste0(FILT,"_",KD)]]<-read.table(path, header=F, sep="\t")
RAD21EPanch2[[paste0("ind",FILT,"_",KD)]]<- rownames(fit) %in% RAD21EPanch2[[paste0(FILT,"_",KD)]]$V4
  }
}
for (KD in KDs) {
RAD21EPanch2[[paste0("ind_",KD)]]<-list(RAD21EPanch2[[paste0("ind","up.FC1","_",KD)]],RAD21EPanch2[[paste0("ind","down.FC1","_",KD)]])
names(RAD21EPanch2[[paste0("ind_",KD)]])<-c("up.FC1","down.FC1")
}



#SA2 vs CTRL fry
fr_SA2indEP <- fry(d, index=SA2indEP, design=design, contrast=conSA2)
gse_SA2indEP <- fr_SA2indEP[names(SA2indEP),] %>% tibble::rownames_to_column("Set")
gse_SA2indEP
#            Set NGenes Direction      PValue         FDR PValue.Mixed
#   up.FC1_SA2KD    220        Up 0.482395977 0.482395977 5.970206e-11
# down.FC1_SA2KD    185      Down 0.004991882 0.009983764 1.152832e-07


for (KD in KDs) {
  PSEAlist[[paste0(KD,"fr_EP.all")]] <- fry(d, index=RAD21EPanch2[[paste0("ind_",KD)]], design=design, contrast=conslist[[KD]])
}
PSEAlist
#$SA2KDfr_EP.all
#         NGenes Direction     PValue        FDR PValue.Mixed    FDR.Mixed
#down.FC1    144      Down 0.01031072 0.02062144 7.305477e-08 7.305477e-08
#up.FC1      193        Up 0.25265325 0.25265325 6.096301e-11 1.219260e-10

#$SA1KDfr_EP.all
#         NGenes Direction    PValue       FDR PValue.Mixed FDR.Mixed
#up.FC1       31      Down 0.8070268 0.9311209   0.56434099 0.5643410
#down.FC1    193      Down 0.9311209 0.9311209   0.07462831 0.1492566



dir.create(file.path(FIGURESDIR,"GSEA","diffRegulatoryloops"))
#barcodeplots fry results: structural loops
for (KD in KDs) {
type="regulatoryLoops"
text1<-paste0('Strenghtend regulatory anchor: RAD21 peaks ')
text2<-paste0('Weakened regulatory anchor: RAD21 peaks ')
pdf(file=file.path(FIGURESDIR,"GSEA","diffRegulatoryloops",paste0("edgeR.PSEA.",KD,".RAD21peaks.diffloop.FC1.barcodeplot.",type,".pdf")), height=4, width=5)
barcodeplot(reslist2[[KD]], 
            index=RAD21EPanch2[[paste0("ind","up.FC1","_",KD)]] ,
            index2=RAD21EPanch2[[paste0("ind","down.FC1","_",KD)]],
            labels=c("CTRL",KD),
            xlab = bquote(log[2]*"FC in RAD21 ChIPseq"),
            cex.axis = 1, 
            main=paste0(""), cex.main=0.5,
            col.bars=c(colorlist[[KD]],"firebrick1"))
par(new=TRUE)
plot.new( )
plot.window( xlim=c(-5,5), ylim=c(-5,5) )
text(5,5,bquote(.(text1)~'('*.(PSEAlist[[paste0(KD,"fr_EP.all")]][paste0("up.FC1"),"NGenes"])*')'), adj = c(1,.3),cex=0.5)
text(5,2.8,bquote(italic(P)[adj.]*"<"*.(PSEAlist[[paste0(KD,"fr_EP.all")]][paste0("up.FC1"),"FDR"])*'('*.(PSEAlist[[paste0(KD,"fr_EP.all")]][paste0("up.FC1"),"Direction"])*')'), adj = c(1,.5),cex=1.5)
text(-5,-5,bquote(.(text2)~'('*.(PSEAlist[[paste0(KD,"fr_EP.all")]][paste0("down.FC1"),"NGenes"])*')'), adj = c(0,.5),cex=0.5)
text(-5,-2.8,bquote(italic(P)[adj.]*"<"*.(PSEAlist[[paste0(KD,"fr_EP.all")]][paste0("down.FC1"),"FDR"])*'('*.(PSEAlist[[paste0(KD,"fr_EP.all")]][paste0("down.FC1"),"Direction"])*')'), adj = c(0,.5),cex=1.5)
dev.off()
}
```


