---
title: "Differential loop size distribution Analysis in Cohesin KD HSPCs"
author: "Alexander Fischer"
date: "02 2023"
output: html
---


# load libraries and set directories
```{r}
## libraries
library(reshape2)
library(ggplot2)
library(beanplot)
library(rstatix)

## paths
WORKDIR="/misc/data/analysis/project_cohesin/CD34/HiC_KDs/DeepSeq_Analysis"
LOOPDIR=file.path(WORKDIR,"loops/withoutXY")
FIGURESDIR=file.path(WORKDIR,"figures","Loopsize_Beanplots")

KDs<-c("SA2KD","SA1KD", "RAD21KD")
Filts<-c("up.FC1", "down.FC1")

## new dir for size plots
dir.create(FIGURESDIR)
```

# read loop analysis files
```{r}
SA2KDloops<-read.table(file.path(LOOPDIR,"SA2KDvsCTRL.merged.loop.scores.DESeq.norm2Total.loopcoords2.txt"),header=T,row.names=1)
SA1KDloops<-read.table(file.path(LOOPDIR,"SA1KDvsCTRL.merged.loop.scores.DESeq.norm2Total.loopcoords2.txt"),header=T,row.names=1)
RAD21KDloops<-read.table(file.path(LOOPDIR,"RAD21KDvsCTRL.merged.loop.scores.DESeq.norm2Total.loopcoords2.txt"),header=T,row.names=1)
LooplistCD34<-list(SA2KDloops,SA1KDloops,RAD21KDloops)
names(LooplistCD34)<-c("SA2KDloops","SA1KDloops","RAD21KDloops")
```


## calculate loop sizes and filter sig. loops
```{r}
for (KD in KDs){
LooplistCD34[[paste0(KD,"loops")]]$loopsize<-LooplistCD34[[paste0(KD,"loops")]]$end2-LooplistCD34[[paste0(KD,"loops")]]$start1
LooplistCD34[[paste0(KD,"loops_filt_strengthened")]]<-subset(LooplistCD34[[paste0(KD,"loops")]],padj<0.05 & log2FoldChange > 0.585)
LooplistCD34[[paste0(KD,"loops_filt_weakened")]]<-subset(LooplistCD34[[paste0(KD,"loops")]],padj<0.05 & log2FoldChange < -0.585)
LooplistCD34[[paste0(KD,"loops_filt_strengthened_Top500")]]<-head(subset(LooplistCD34[[paste0(KD,"loops")]],padj<0.05 & log2FoldChange > 0.585),500)
LooplistCD34[[paste0(KD,"loops_filt_weakened_Top500")]]<-head(subset(LooplistCD34[[paste0(KD,"loops")]],padj<0.05 & log2FoldChange < -0.585),500)
}
```

## report numbers for size categories
```{r}
changes<-c("strengthened","weakened")
sizedf<-data.frame()
for (KD in KDs){
   for (change in changes){ 
sizedf[paste0(change,"<500kb"),KD]<-nrow(subset(LooplistCD34[[paste0(KD,"loops_filt_",change)]],loopsize<500000))
sizedf[paste0(change,">500kb"),KD]<-nrow(subset(LooplistCD34[[paste0(KD,"loops_filt_",change)]],loopsize>500000))
sizedf[paste0(change,"total"),KD]<-nrow(subset(LooplistCD34[[paste0(KD,"loops_filt_",change)]]))
}}
sizedf
write.table(sizedf,file.path(LOOPDIR,"diffloopsizes.txt"))
```

## test up vs down sizes with wilcoxon test
```{r}
#1vs1 comparisons
for (KD in KDs){
LooplistCD34[[paste0(KD,"wilcoxonUPvsDown")]]<-wilcox.test(LooplistCD34[[paste0(KD,"loops_filt_strengthened")]]$loopsize,LooplistCD34[[paste0(KD,"loops_filt_weakened")]]$loopsize)
LooplistCD34[[paste0(KD,"wilcoxonUPvsDownTop500")]]<-wilcox.test(LooplistCD34[[paste0(KD,"loops_filt_strengthened_Top500")]]$loopsize,LooplistCD34[[paste0(KD,"loops_filt_weakened_Top500")]]$loopsize)
}


#SA1 vs. SA2 all comparisons all tests with BH adjustment
df1<-data.frame(size=LooplistCD34[[paste0("SA1KD","loops_filt_strengthened")]]$loopsize,group="SA1KD",change="strengthened",groupchange="SA1KD_strenghthened")
df2<-data.frame(size=LooplistCD34[[paste0("SA1KD","loops_filt_weakened")]]$loopsize,group="SA1KD",change="weakened",groupchange="SA1KD_weakened")
df3<-data.frame(size=LooplistCD34[[paste0("SA2KD","loops_filt_strengthened")]]$loopsize,group="SA2KD",change="strengthened",groupchange="SA2KD_strenghthened")
df4<-data.frame(size=LooplistCD34[[paste0("SA2KD","loops_filt_weakened")]]$loopsize,group="SA2KD",change="weakened",groupchange="SA2KD_weakened")
combdf<-rbind(df1,df2,df3,df4)

library(rstatix)
wilSA1SA2<-wilcox_test(formula=size ~ groupchange,data=combdf, paired = FALSE, p.adjust.method = "BH")
wilSA1SA2
write.table(wilSA1SA2,file.path(LOOPDIR,"SA1vsSA2_wilcoxon.BHadj.txt"),sep="\t",quote=F,row.names=F)



```

## generate beanplots comparing sizes of weakened and strengthened loops using all sign. loops
```{r}
LooplistCD34[["SA2KDcolors"]] <- c("springgreen1","springgreen3")
LooplistCD34[["SA1KDcolors"]] <- c("#f7bd37","darkgoldenrod3")
LooplistCD34[["RAD21KDcolors"]] <- c("mediumorchid1","mediumorchid3")
marks <- c(20000,50000,100000,200000, 500000,1000000,2000000, 5000000,10000000)

for (KD in KDs){
str<-LooplistCD34[[paste0(KD,"loops_filt_strengthened")]]$loopsize
weak<-LooplistCD34[[paste0(KD,"loops_filt_weakened")]]$loopsize
if (length(str)<length(weak)){
data<-data.frame(weakened=weak)
data$strenghtened<-c(str, rep(NA, nrow(data)-length(str)))  #add NAs to fill up missing entries other wise no dataframe
}
else {
weakenedNA<-c(weak, rep(NA, length(str)-length(weak))) #add NAs to fill up missing entries other wise no dataframe
data<-data.frame(weakened=weakenedNA,strengthened=str)
}
if (LooplistCD34[[paste0(KD,"wilcoxonUPvsDown")]]$p.value>0.05){
siglvl<-"ns"
}
else {
siglvl<-"****"
}
pdf(file=file.path(FIGURESDIR,paste0("Beanplot_LoopSizeDistribution.",KD,".vsCTRL_sig.n2t.pdf")), height=4, width=2.5)
beanplot(data,log="y", bw="nrd",what=c(0,1,0,0),axes = FALSE, col = "darkgrey", border = "darkgrey" ,overallline = "median", method="jitter", boxwex = 1.25, beanlinewd = 1, maxstripline = 0.8, ylim=c(20000,10000000), lwd=0.5)
par(new=TRUE)
boxplot(data,range=0,log="y", style="quantile",axes = FALSE, col = LooplistCD34[[paste0(KD,"colors")]], alpha=0.5, border = "black",overallline = "median", notch=TRUE, boxwex = 0.5, staplewex = 0.5, ylim=c(20000,10000000),lwd=0.6)
axis(1,padj=-0.8,family="Helvetica",cex.axis=0.5,at=1:2, labels=c("weakened","strenghtened"))
axis(2,padj=0.4,family="Helvetica",cex.axis=0.6,at=marks,labels=format(marks,scientific=FALSE),las=2,mgp=c(0,.6,0))
mtext("loop domain size",family="Helvetica",side=2,line=2,cex=1.2,padj=-1.2)
mtext(paste0("in ",KD),family="Helvetica",ps=12,side=1,line=2,cex=1.2,padj=0)
text(1.5,10000000,siglvl) #adding significance levels
segments(1,5000000,2,5000000)
dev.off()
}
```

## generate beanplots comparing sizes of weakened and strengthened loops using top 500 sign. loops
```{r}
for (KD in KDs){
str<-LooplistCD34[[paste0(KD,"loops_filt_strengthened_Top500")]]$loopsize
weak<-LooplistCD34[[paste0(KD,"loops_filt_weakened_Top500")]]$loopsize
if (length(str)<length(weak)){
data<-data.frame(weakened=weak)
data$strenghtened<-c(str, rep(NA, nrow(data)-length(str)))  #add NAs to fill up missing entries other wise no dataframe
}
else {
weakenedNA<-c(weak, rep(NA, length(str)-length(weak))) #add NAs to fill up missing entries other wise no dataframe
data<-data.frame(weakened=weakenedNA,strengthened=str)
}
if (LooplistCD34[[paste0(KD,"wilcoxonUPvsDownTop500")]]$p.value>0.05){
siglvl<-"ns"
}
else {
siglvl<-"****"
}
pdf(file=file.path(FIGURESDIR,paste0("Beanplot_LoopSizeDistribution.",KD,".vsCTRL_top500.sig.n2t.pdf")), height=4, width=2.5)
beanplot(data,log="y", bw="nrd",what=c(0,1,0,0),axes = FALSE, col = "darkgrey", border = "darkgrey" ,overallline = "median", method="jitter", boxwex = 1.25, beanlinewd = 1, maxstripline = 0.8, ylim=c(20000,10000000), lwd=0.5)
par(new=TRUE)
boxplot(data,range=0,log="y", style="quantile",axes = FALSE, col = LooplistCD34[[paste0(KD,"colors")]], alpha=0.5, border = "black",overallline = "median", notch=TRUE, boxwex = 0.5, staplewex = 0.5, ylim=c(20000,10000000),lwd=0.6)
axis(1,padj=-0.8,family="Helvetica",cex.axis=0.5,at=1:2, labels=c("weakened","strenghtened"))
axis(2,padj=0.4,family="Helvetica",cex.axis=0.6,at=marks,labels=format(marks,scientific=FALSE),las=2,mgp=c(0,.6,0))
mtext("loop domain size",family="Helvetica",side=2,line=2,cex=1.2,padj=-1.2)
mtext(paste0("in ",KD),family="Helvetica",ps=12,side=1,line=2,cex=1.2,padj=0)
text(1.5,10000000,"****") #adding significance levels
segments(1,5000000,2,5000000)
dev.off()
}
```


#### SA1 and SA2 KD combined plot
## generate beanplots comparing sizes of weakened and strengthened loops using all sign. loops in SA1 and SA2KD results
```{r}
weakSA1<-LooplistCD34[["SA1KDloops_filt_weakened"]]$loopsize #longest vector
strSA1<-c(LooplistCD34[["SA1KDloops_filt_strengthened"]]$loopsize,rep(NA, length(weakSA1)-length(LooplistCD34[["SA1KDloops_filt_strengthened"]]$loopsize)))
weakSA2<-c(LooplistCD34[["SA2KDloops_filt_weakened"]]$loopsize,rep(NA, length(weakSA1)-length(LooplistCD34[["SA2KDloops_filt_weakened"]]$loopsize)))
strSA2<-c(LooplistCD34[["SA2KDloops_filt_strengthened"]]$loopsize,rep(NA, length(weakSA1)-length(LooplistCD34[["SA2KDloops_filt_strengthened"]]$loopsize)))

data<-data.frame(weakSA2=weakSA2,strSA2=strSA2,weakSA1=weakSA1,strSA1=strSA1)
colvec<-c(LooplistCD34[["SA2KDcolors"]],LooplistCD34[["SA1KDcolors"]])

pdf(file=file.path(FIGURESDIR,paste0("Beanplot_LoopSizeDistribution.SA1vsSA2KD.vsCTRL_sig.n2t.pdf")), height=4, width=4)
beanplot(data,log="y", bw="nrd",what=c(0,1,0,0),axes = FALSE, col = "darkgrey", border = "darkgrey" ,overallline = "median", method="jitter", boxwex = 1.25, beanlinewd = 1, maxstripline = 0.8, ylim=c(20000,10000000), lwd=0.5)
par(new=TRUE)
boxplot(data,range=0,log="y", style="quantile",axes = FALSE, col = colvec, alpha=0.5, border = "black",overallline = "median", notch=TRUE, boxwex = 0.5, staplewex = 0.5, ylim=c(20000,10000000),lwd=0.6)
axis(1,padj=-0.8,family="Helvetica",cex.axis=0.5,at=1:2, labels=c("weakened","strenghtened"))
axis(1,padj=-0.8,family="Helvetica",cex.axis=0.5,at=3:4, labels=c("weakened","strenghtened"))
axis(2,padj=0.4,family="Helvetica",cex.axis=0.6,at=marks,labels=format(marks,scientific=FALSE),las=2,mgp=c(0,.6,0))
mtext("loop domain size",family="Helvetica",side=2,line=2,cex=1.2,padj=-1.2)
mtext(paste0("in STAG2 KD"),family="Helvetica",ps=12,side=1,at=1.5,line=2,cex=1.2,padj=0)
mtext(paste0("in STAG1 KD"),family="Helvetica",ps=12,side=1,at=3.5,line=2,cex=1.2,padj=0)
text(1.5,6000000,subset(wilSA1SA2,group1=="SA2KD_strenghthened" & group2=="SA2KD_weakened")$p.adj.signif) #adding significance levelstext(1.5,5200000,subset(wilSA1SA2,group1=="SA2KD_strenghthened" & group2=="SA2KD_weakened")$p.adj.signif) #adding significance levels
text(3,8000000,subset(wilSA1SA2,group1=="SA1KD_strenghthened" & group2=="SA2KD_strenghthened")$p.adj.signif) #adding significance levels
text(3.5,6000000,subset(wilSA1SA2,group1=="SA1KD_strenghthened" & group2=="SA1KD_weakened")$p.adj.signif) #adding significance levels
text(2,11000000,subset(wilSA1SA2,group1=="SA1KD_weakened" & group2=="SA2KD_weakened")$p.adj.signif) #adding significance levels
segments(1,5000000,2,5000000)
segments(2,7000000,4,7000000)
segments(3,5000000,4,5000000)
segments(1,9500000,3,9500000)
dev.off()

```


