---
title: "Differential Analyis of STAG1 vs STAG2 ChIPseq data in Cohesin-wildtype CTRL-AMLs"
author: "Alexander Fischer"
date: "04 2022"
output: html
---


# paths and libraries
```{r}
PROJDIR<-"/misc/data/analysis/project_cohesin"
WORKDIRCHIP<-file.path(PROJDIR,"Cohesin_AML/ChIP_analysis")
ANNDIR=file.path(WORKDIRCHIP,"annotation_tables")
FIGDIR=file.path(WORKDIRCHIP,"figures")
DIFFDIR=file.path(WORKDIRCHIP,"diffPeaks","SA1vsSA2")
DIFFDIRRAD21=file.path(WORKDIRCHIP,"diffPeaks")
STAGDOMCD34DIR<-file.path(PROJDIR,"CD34/ChIP_KD_analysis/Cohesin_CTCF_MED12/diffPeaks/SA1vsSA2")

library(edgeR)
library(ggplot2)
library(ggrepel)
```

# get data and define group vectors
```{r}
meta<-read.delim(file.path(WORKDIRCHIP,"Rad21_AML_meta.csv"),sep=",",header=TRUE)
matchedCTRLpat<-c("ctr_16911", "ctr_18136", "ctr_18519", "ctr_19405", "ctr_19416" ,"ctr_21047")
metamatched<-subset(meta,meta$shortID_vial %in% matchedCTRLpat)

datastr <- read.delim(file.path(ANNDIR,"CTRL.AML.matched.SA1.SA2.strRAD21.peaks.ann.tab.filt.txt"), row.names=1)
datareg <- read.delim(file.path(ANNDIR,"CTRL.AML.matched.SA1.SA2.RAD21.peaks.ann.tab.filt.txt"), row.names=1)

group<-factor(c(rep("SA1",6),rep("SA2",6)),levels=c("SA1","SA2"))
patient<-factor(c(rep(matchedCTRLpat,2)))
sex<-factor(rep(metamatched$sex,2))
NPM1<-factor(rep(metamatched$npm1,2))
FLT3ITD<-factor(rep(metamatched$flt3itd,2))
DNMT3<-factor(rep(metamatched$Dmut,2))
RUNX1<-factor(rep(metamatched$RUNX1,2))

colnames(data)<-paste0(group,patient)
```

# Analysis based on AML RAD21 peak-annotated data
## edgeR contrast SA1 vs SA2 - annotated to AML RAD21 peaks (stringent/regular sets)
```{r}
#Stringent peak annotated
##create dgelist object # include the patient ID as covariate for design
d <- DGEList(counts=data,group=group)
design <- model.matrix(~0 + group + patient)
rownames(design) <- colnames(d)
d <- calcNormFactors(d)
d <- estimateDisp(d, design, robust=TRUE)
fit <- glmQLFit(d, design)
## contrasts and qlf test
con <- makeContrasts(groupSA1-groupSA2, levels=design)
qlf <- glmQLFTest(fit, contrast=con)
qstat <- topTags(qlf, n=Inf)
write.table (qstat, file = file.path(DIFFDIR,"qstat.CTRL.AML.topQC.SA1vsSA2.strRAD21peak.ann.txt"), sep = "\t", quote=FALSE)
summary(qs <- decideTestsDGE(qlf, lfc=1))


#all RAD21 peak annotated
##create dgelist object # include the patient ID as covariate for design
d2 <- DGEList(counts=data,group=group)
rownames(design) <- colnames(d2)
d2 <- calcNormFactors(d2)
d2 <- estimateDisp(d2, design, robust=TRUE)
fit2 <- glmQLFit(d2, design)
## qlf test
qlf2 <- glmQLFTest(fit2, contrast=con)
qstat2 <- topTags(qlf2, n=Inf)
write.table (qstat2, file = file.path(DIFFDIR,"qstat.CTRL.AML.topQC.SA1vsSA2.RAD21.ann.txt"), sep = "\t", quote=FALSE)
summary(qs <- decideTestsDGE(qlf2, lfc=1))
```



## Filter results for STAG Dominant or common positions
```{r}
statslist<-list()
PTS=c("strRAD21","RAD21")
for (p in PTS){
PEAKcoords<-read.delim(file=file.path(ANNDIR,paste0("CTRL.AML.topQCpat.SA1.SA2.",p,".peaks.ann.filt.txt")),row.names=1)
PEAKcoords<-PEAKcoords[,c(1:4)]
SA2vsSA1edgeRresult<- read.delim(file=file.path(DIFFDIR,paste0("qstat.CTRL.AML.topQC.SA1vsSA2.",p,".ann.txt")),row.names=1)
#add peak coordinates
PEAKcoords<-subset(PEAKcoords, row.names(PEAKcoords) %in% row.names(SA2vsSA1edgeRresult))
SA2vsSA1edgeRresult2<-merge(PEAKcoords,SA2vsSA1edgeRresult,by=0)
row.names(SA2vsSA1edgeRresult2)<-SA2vsSA1edgeRresult2$Row.names
SA2vsSA1edgeRresult2<-SA2vsSA1edgeRresult2[,-1]
#subset for types of positions
SA1up<-SA2vsSA1edgeRresult2[which(SA2vsSA1edgeRresult2$FDR < 0.05 & SA2vsSA1edgeRresult2$logFC > (1) ), ] #these are the STAG1 dominatnt sites
write.table (SA1up,file = file.path(DIFFDIR,paste0("SA1vsSA2.CTRL.AML.RAD21.",p,".SA1dom.txt")), sep = "\t", row.names=T, quote=FALSE)
SA1down<-SA2vsSA1edgeRresult2[ which(SA2vsSA1edgeRresult2$FDR < 0.05 & SA2vsSA1edgeRresult2$logFC < (-1) ), ] #these are the STAG2 dominatnt sites
write.table (SA1down,file = file.path(DIFFDIR,paste0("SA1vsSA2.CTRL.AML.RAD21.",p,".SA2dom.txt")), sep = "\t", row.names=T, quote=FALSE)
common<-subset(SA2vsSA1edgeRresult2,!(row.names(SA2vsSA1edgeRresult2) %in% row.names(SA1up) | row.names(SA2vsSA1edgeRresult2) %in% row.names(SA1down))) #the rest is treated as "common"
write.table (common,file = file.path(DIFFDIR,paste0("SA1vsSA2.CTRL.AML.RAD21.",p,".common.txt")), sep = "\t", row.names=T, quote=FALSE)
stats.df<-data.frame(row.names=c("STAG2dominated","STAG1dominated","common"),group=c("STAG2dominated","STAG1dominated","common"),number_of_peaks = c(nrow(SA1down),nrow(SA1up),nrow(common)))
stats.df$prop<-stats.df$number_of_peaks/sum(stats.df$number_of_peaks)
stats.df$ypos<- cumsum(stats.df$prop)- 0.5*stats.df$prop
statslist[[p]]<-stats.df
write.table(statslist[[p]],file=file.path(FIGDIR,paste0("STAGdom.sites.CTRL.AML.",p,".peakset.stats.txt")),quote=FALSE,sep="\t")
}
```


## plot STAG dominance as piechart
```{r}
for (p in PTS){
statslist[[paste0("pie",p)]]<-ggplot(statslist[[p]], aes(x="", y=prop, fill=group)) +
  scale_fill_manual(values = c("STAG2dominated"="seagreen2","STAG1dominated"="darkgoldenrod","common"="grey"))+
  geom_bar(stat="identity", width=1, color="white") +
  coord_polar("y", start=0) + 
  theme_void() +
  #geom_text(aes(y = ypos, label = group), color = "white", size=6)
  geom_text_repel(aes(label = group, y = ypos), 
                   nudge_x = 0.05, nudge_y = 0.05,
                   size = 5, show.legend = F)
}
for (p in PTS){
ggsave(file=file.path(FIGDIR,paste0("STAGdom.sites.CTRL.AML.",p,".peakset.piechart.pdf")), height=6, width=6, plot=statslist[[paste0("pie",p)]])
}
```


# Analysis based on AML differnatial RAD21 peak positions

## read in peak sets
```{r}
CTCFcats<-c("withCTCF","noCTCF")
FILTS<-c("up.FC1","down.FC1")
allAMLRAD21_peaksets<-list()
## read diffpeak positions by CTCF subsets (only SA2mut)
for (CTCF in CTCFcats){
  MUT="SA2mut"
allAMLRAD21_peaksets[[paste0(MUT,"_strRAD21diffpeaks_up.FC1",CTCF)]]<-read.table(file.path(DIFFDIRRAD21,"subsetByCTCF",paste0(MUT,"vsCTRL.RAD21.stringentPeaks_DESEQ.model.2foldup.",CTCF,".txt")),header=FALSE)
allAMLRAD21_peaksets[[paste0(MUT,"_strRAD21diffpeaks_down.FC1",CTCF)]]<-read.table(file.path(DIFFDIRRAD21,"subsetByCTCF",paste0(MUT,"vsCTRL.RAD21.stringentPeaks_DESEQ.model.2folddown.",CTCF,".txt")),header=FALSE)
}
```

## set index
```{r}
for (filt in FILTS){
for (CTCF in CTCFcats){
MUT="SA2mut"
allAMLRAD21_peaksets[[paste0("ind_",MUT,"_strRAD21diffpeaks_",filt,CTCF)]] <- rownames(fit) %in% allAMLRAD21_peaksets[[paste0(MUT,"_strRAD21diffpeaks_",filt,CTCF)]]$V1
}}

#combined index list for complete sets CTCF subsets

allAMLRAD21_peaksets[[paste0("ind_SA2mut_strRAD21diffpeaks_byCTCF")]]<- list(
allAMLRAD21_peaksets[["ind_SA2mut_strRAD21diffpeaks_up.FC1withCTCF"]],allAMLRAD21_peaksets[["ind_SA2mut_strRAD21diffpeaks_up.FC1noCTCF"]],
allAMLRAD21_peaksets[["ind_SA2mut_strRAD21diffpeaks_down.FC1withCTCF"]],allAMLRAD21_peaksets[["ind_SA2mut_strRAD21diffpeaks_down.FC1noCTCF"]]
)
names(allAMLRAD21_peaksets[[paste0("ind_SA2mut_strRAD21diffpeaks_byCTCF")]])<-c("strengthenedRAD21_withCTCF","strengthenedRAD21_noCTCF","weakenedRAD21_withCTCF","weakenedRAD21_noCTCF")
```


## fry contrast for peaksets in CTRL-AML SA1vsSA2
```{r}
###all up/down sets separately
fry_allAMLRAD21_peaksets<-list()
fry_allAMLRAD21_peaksets[[paste0("ind_SA2mut_strRAD21diffpeaks_byCTCF")]]<-fry(ds, index=allAMLRAD21_peaksets[[paste0("ind_SA2mut_strRAD21diffpeaks_byCTCF")]], design=design, contrast=con)


dir.create(file.path(WORKDIRCHIP,"figures","SA1vsSA2_PSEA"))

#barcode plot for the STAG preference at the RAD21 (stringent peak set) diffpeaks

#for the stringent RAD21 diffpeaks
for (KD in MUTs){
cat<-"_strRAD21diffpeaks"
pdf(file=file.path(WORKDIRCHIP,"figures","SA1vsSA2_PSEA",paste0("edgeR_barcodeplot_SA1vsSA2_inCTRL_AMLs_",cat,"SA2mut",".pdf")), height=6, width=6)
barcodeplot(qlf2$table$logFC, 
            index=allAMLRAD21_peaksets[[paste0("ind_","SA2mut",cat,"_up.FC1")]] ,
            index2=allAMLRAD21_peaksets[[paste0("ind_","SA2mut",cat,"_down.FC1")]],
            labels=c("STAG2 dominance","STAG1 dominance"),
            xlab = bquote(log[2]*"FC of SA1 vs SA2 in CTRL-AMLs"),
            cex.axis = 1, 
            main=paste0("SA2mut", ": STAG1 vs STAG2 at ", cat), cex.main=0.5,
            col.bars=c("darkgoldenrod1","seagreen1"))
par(new=TRUE)
plot.new( )
plot.window( xlim=c(-5,5), ylim=c(-5,5) )
text(5,5,bquote('all RAD21 peaks strengthened '~'('*.(fry_allAMLRAD21_peaksets[[paste0("ind_","SA2mut", cat)]]["strengthenedRAD21","NGenes"])*')'), adj = c(1,.3),cex=1)
text(5,4.5,bquote(italic(P)[adj.]*"<"*.(fry_allAMLRAD21_peaksets[[paste0("ind_","SA2mut", cat)]]["strengthenedRAD21","FDR"])*'('*.(fry_allAMLRAD21_peaksets[[paste0("ind_","SA2mut", cat)]]["strengthenedRAD21","Direction"])*')'), adj = c(1,.5),cex=1.5)
text(-5,-5,bquote('all RAD21 peaks weakened  '~'('*.(fry_allAMLRAD21_peaksets[[paste0("ind_","SA2mut", cat)]]["weakenedRAD21","NGenes"])*')'), adj = c(0,.5),cex=1)
text(-5,-4.5,bquote(italic(P)[adj.]*"<"*.(fry_allAMLRAD21_peaksets[[paste0("ind_","SA2mut", cat)]]["weakenedRAD21","FDR"])*'('*.(fry_allAMLRAD21_peaksets[[paste0("ind_","SA2mut", cat)]]["weakenedRAD21","Direction"])*')'), adj = c(0,.5),cex=1.5)
dev.off()
}
```


### Summarize stats for combined stringent RAD21 diffpeak by CTCF association
```{r}
fry_allAMLRAD21_peaksets3<-fry_allAMLRAD21_peaksets[[paste0("ind_SA2mut_strRAD21diffpeaks_byCTCF")]]
#prepare table for plotting
##negative log FDR (positive value)
fry_allAMLRAD21_peaksets3$logFDR<--log(fry_allAMLRAD21_peaksets3$FDR)
fry_allAMLRAD21_peaksets3$logFDR<--log(fry_allAMLRAD21_peaksets3$FDR)
##for down direction use log FDR (negativ value)
fry_allAMLRAD21_peaksets3$logFDR[fry_allAMLRAD21_peaksets3$Direction == "Down"] <- log(subset(fry_allAMLRAD21_peaksets3,Direction=="Down")[,"FDR"])
fry_allAMLRAD21_peaksets3$logFDR[fry_allAMLRAD21_peaksets3$Direction == "Down"] <- log(subset(fry_allAMLRAD21_peaksets3,Direction=="Down")[,"FDR"])
##define categories
fry_allAMLRAD21_peaksets3$RAD21change<-strsplit2(rownames(fry_allAMLRAD21_peaksets3),"_",fixed = TRUE)[,1]
fry_allAMLRAD21_peaksets3$CTCFassociation<-strsplit2(rownames(fry_allAMLRAD21_peaksets3),"_",fixed = TRUE)[,2]
##set levels
fry_allAMLRAD21_peaksets3$RAD21change<-factor(fry_allAMLRAD21_peaksets3$RAD21change,levels=c("weakenedRAD21","strengthenedRAD21"))
fry_allAMLRAD21_peaksets3$CTCFassociation<-factor(fry_allAMLRAD21_peaksets3$CTCFassociation,levels=c("withCTCF","noCTCF"))
write.table(fry_allAMLRAD21_peaksets3,file=file.path(WORKDIRCHIP,"figures","SA1vsSA2_PSEA","PSEA.SA1vsSA2_inCTRL_AMLs_strRAD21.diffpeaks.byCTCF.txt"),sep="\t",quote=FALSE)



#plot
p1<-ggplot(fry_allAMLRAD21_peaksets3, aes(x= logFDR, y= rownames(fry_allAMLRAD21_peaksets3),fill = RAD21change)) +
geom_bar(aes(colour=CTCFassociation),size=1.5, width=0.8, position = position_dodge(width=0.5), stat="identity") +
  scale_fill_manual(values = c("strengthenedRAD21"="seagreen","weakenedRAD21"="firebrick3"))+
  scale_color_manual(values = c("withCTCF"="darkblue","noCTCF"="orange"))+
  scale_x_continuous(limits = c(-10,10),labels=abs,breaks = seq(-10,10,by=2.5)) + 
  xlab("- log2 FDR of SA1 vs SA2 PSEA in CTRL-AMLs") + ylab(" ") + 
  ggtitle(paste0("Differential RAD21 peaks in SA2mut")) +
  theme(
  axis.text.x = element_text(angle=0,vjust=0.2,size=16),
  legend.text = element_text(angle=0,vjust=0.2,size=16),
  axis.text.y = element_text(angle=0,hjust=0.95,vjust=0.2,size=0),
  legend.position="bottom", legend.title=element_blank(),
  axis.title=element_text(size=16,face="bold"), plot.title = element_text(size = 20, face = "bold",hjust = 0.5), 
  panel.border = element_blank(),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.background = element_blank(),
  axis.line.y = element_blank(),
  axis.ticks.y = element_blank(),
  axis.line.x = element_line(colour = "black"),)+
geom_vline(xintercept = -log2(0.05),linetype = 'dashed') + 
geom_vline(xintercept = log2(0.05),linetype = 'dashed',color="black")+
geom_text(aes(x=-8,y=4,label="SA2 dominant"),size=4,vjust=0.2,fontface = "bold")+
geom_text(aes(x=8,y=4,label="SA1 dominant"),size=4,vjust=0.2,fontface = "bold")

pdf(file=file.path(WORKDIRCHIP,"figures","SA1vsSA2_PSEA",paste0("SA2mut.SA1vsSA2_inCTRL_AMLs.PSEA.summary.DIFF.RAD21peaks.stringent.pdf")), height=6, width=6)
plot(p1)
dev.off()
```




# Analysis based on HSPC RAD21 peak-annotated data
## edgeR contrast SA1 vs SA2 - annotated to: all CD34 RAD21 pos. peak set
```{r}
#all RAD21 peak set including X and Y chrom (test for 1:1 comparability?)
dataCD34ann <- read.delim(file.path(WORKDIRCHIP,"annotation_tables","CTRL.AML.matched.SA1.SA2.CD34RAD21.peaks.ann.txt"), row.names=1)
dataCD34ann <-dataCD34ann[,c(19:30)]
colnames(dataCD34ann)<-paste0(group,patient)
#create dgelist object and filter # include the patient ID as covariate for design
dCD34ann <- DGEList(counts=dataCD34ann,group=group)
design <- model.matrix(~0 + group + patient)
rownames(design) <- colnames(dCD34ann)
dCD34ann <- calcNormFactors(dCD34ann)
dCD34ann <- estimateDisp(dCD34ann, design, robust=TRUE)
fitCD34ann <- glmQLFit(dCD34ann, design)
con <- makeContrasts(groupSA1-groupSA2, levels=design)
qlfCD34ann <- glmQLFTest(fitCD34ann, contrast=con)
qstatCD34ann <- topTags(qlfCD34ann, n=Inf)
summary(qs <- decideTestsDGE(qlfCD34ann, lfc=1))
#       1*groupSA1 -1*groupSA2
#Down                     7195
#NotSig                  65884
#Up                       5788
```

## get STAG1/2 dominant sites from CD34+ HSPC analyis
```{r}
###read data
CD34_STAG_DOM_list<-list()
CD34_STAG_DOM_list[["common"]]<-read.table(file.path(STAGDOMCD34DIR,"SA2vsSA1.allCD34.RAD21.peaks_edgeR.common.txt"))
CD34_STAG_DOM_list[["SA1dom"]]<-read.table(file.path(STAGDOMCD34DIR,"SA2vsSA1.allCD34.RAD21.peaks_edgeR.2folddown.txt"))
CD34_STAG_DOM_list[["SA2dom"]]<-read.table(file.path(STAGDOMCD34DIR,"SA2vsSA1.allCD34.RAD21.peaks_edgeR.2foldup.txt"))
###create indices
STAGprefs<-c("common","SA1dom","SA2dom")
for (pref in STAGprefs){
CD34_STAG_DOM_list[[paste0("ind_",pref)]]<- rownames(fitCD34ann) %in% row.names(CD34_STAG_DOM_list[[pref]])
}
###create combined index
CD34_STAG_DOM_list[["ind_comb"]]<-list(CD34_STAG_DOM_list[[paste0("ind_","common")]],CD34_STAG_DOM_list[[paste0("ind_","SA1dom")]],CD34_STAG_DOM_list[[paste0("ind_","SA2dom")]])
names(CD34_STAG_DOM_list[["ind_comb"]])<-STAGprefs
```
## enrichment of HSPC STAG dom/common in AML STAG contrast
```{r}
dir.create(file.path(WORKDIRCHIP,"figures","SA1vsSA2_PSEA"))
#Peak set enrichment using limma fry
CD34_STAG_DOM_GSEA<-fry(dCD34ann, index=CD34_STAG_DOM_list[["ind_comb"]], design=design, contrast=con)
CD34_STAG_DOM_GSEA
#       NGenes Direction       PValue          FDR PValue.Mixed    FDR.Mixed
#SA1dom   2056        Up 0.0001343637 0.0003760395 1.776001e-05 1.776001e-05
#SA2dom   1319      Down 0.0002506930 0.0003760395 1.770371e-05 1.776001e-05
#common  75492        Up 0.0370522388 0.0370522388 6.115183e-06 1.776001e-05
write.table(CD34_STAG_DOM_GSEA,file=file.path(WORKDIRCHIP,"figures","SA1vsSA2_PSEA","PSEA.SA1vsSA2_inCTRL_AMLs_STAGdominant_sites_inCD34.txt"),sep="\t",quote=FALSE)

#generate peak set enrichment barcodeplot showing the fry statistics in plot
pdf(file=file.path(WORKDIRCHIP,"figures","SA1vsSA2_PSEA",paste0("edgeR_barcodeplot_SA1vsSA2_inCTRL_AMLs_STAGdominant_sites_inCD34.pdf")), height=6, width=6)
barcodeplot(qlfCD34ann$table$logFC, 
            index=CD34_STAG_DOM_list[["ind_SA1dom"]] ,
            index2=CD34_STAG_DOM_list[["ind_SA2dom"]],
            labels=c("STAG2 dominance","STAG1 dominance"),
            xlab = bquote(log[2]*"FC of SA1 vs SA2 in CTRL-AMLs"),
            cex.axis = 0.5, 
            main=paste0("CTRL-AML: STAG1 vs STAG2 \n at postions with STAG dominace in HSPCs"), cex.main=0.5,
            col.bars=c("darkgoldenrod3","seagreen3"))
par(new=TRUE)
plot.new( )
plot.window( xlim=c(-5,5), ylim=c(-5,5) )
text(5,5,bquote('STAG1 dominated in HSPCs '~'('*.(CD34_STAG_DOM_GSEA["SA1dom","NGenes"])*')'), adj = c(1,.3),cex=1)
text(5,4.5,bquote(italic(P)[adj.]*"<"*.(CD34_STAG_DOM_GSEA["SA1dom","FDR"])*'('*.(CD34_STAG_DOM_GSEA["SA1dom","Direction"])*')'), adj = c(1,.5),cex=1.5)
text(-5,-5,bquote('STAG2 dominated in HSPCs  '~'('*.(CD34_STAG_DOM_GSEA["SA2dom","NGenes"])*')'), adj = c(0,.5),cex=1)
text(-5,-4.5,bquote(italic(P)[adj.]*"<"*.(CD34_STAG_DOM_GSEA["SA2dom","FDR"])*'('*.(CD34_STAG_DOM_GSEA["SA2dom","Direction"])*')'), adj = c(0,.5),cex=1.5)
dev.off()
```