---
title: "Cohesin KD cell number expansion Analysis in CD34 HSPCs d4"
author: "Alexander Fischer"
date: "Oct 2022"
output: html_document
---

#rbioc_3-12
# set paths and load libraries
```{r setup, include=FALSE}
DIR_DATA="/misc/data"
PROJDIR=file.path(DIR_DATA,"analysis/project_cohesin/CD34")
WORKDIR=file.path(PROJDIR,"KD_expansion_stats")
library(ggplot2)
library(reshape2)
library(tidyverse)
library(rstatix)
library(plyr)
```

# prepare data tables
```{r}
#read in csv file
Expdata<-read.csv(file.path(WORKDIR,"siRNA_expansion.csv"))
row.names(Expdata)<-Expdata$X... 
colnames(Expdata)[1]<-"Sample"

#remove MED12KDdata
Expdata<-subset(Expdata,group != "MED12KD")
#more exact distinction of ctrl types:
Expdata$group2<-Expdata$group
Expdata$group2[Expdata$group_detail=="Mock"]<-"MOCK"
Expdata$group2[Expdata$group_detail=="Untreated"]<-"UNTR"
#label ID for plots
Expdata$labelID<-paste0(Expdata$group2,"_",Expdata$donor)
Expdata$labelIDdetailed<-paste0(Expdata$group_detail,"_",Expdata$donor)

#define group vector
group<-factor(Expdata$group,levels=c("CTRL","STAG2KD","STAG1KD","STAG1_2KD","RAD21KD"))
group2<-factor(Expdata$group2,levels=c("CTRL","MOCK","UNTR","STAG2KD","STAG1KD","STAG1_2KD","RAD21KD"))
#calculate Expansion between split timepoints
Expdata[,"d0_4_expansion"]<-Expdata[,"d4_living"]/Expdata[,"d0"]


#transpose
Expdatatr<-data.frame(t(Expdata))

#means by condition
##stack data and add CFU category as well as group in each row
dat.st<-stack(as.data.frame(Expdatatr[c("d0_4_expansion"),]))   ##stack data to a single column can be used for geom_jitter points
dat.st$perc<- rep(c("d0_4_expansion"), times = ncol(Expdatatr))  ##add timpoints as extra col
dat.st$group=rep(group, each=2)   ##add group identity as extra col
dat.st$group2=rep(group2, each=2)
dat.st$labelID=rep(Expdata$labelID, each=2)  
dat.st$values<-as.numeric(dat.st$values)

##calculate mean values and standard dev. all controls combined
mean.dat<-ddply(dat.st, .(group,perc),
summarize,
mean= mean(as.numeric(values),2),
sd=sd(as.numeric(values))
)
mean.dat

##calculate mean values and standard dev. control subsets
mean.dat2<-ddply(dat.st, .(group2,perc),
summarize,
mean= mean(as.numeric(values),2),
sd=sd(as.numeric(values))
)
mean.dat2

#set levels to desired order in plots
mean.dat$group <- factor(mean.dat$group , levels = c("CTRL","STAG2KD","STAG1KD","STAG1_2KD","RAD21KD"))
mean.dat2$group2 <- factor(mean.dat2$group2 , levels = c("CTRL","MOCK","UNTR","STAG2KD","STAG1KD","STAG1_2KD","RAD21KD"))
dat.st$group <- factor(dat.st$group , levels = c("CTRL","STAG2KD","STAG1KD","STAG1_2KD","RAD21KD"))

```

# test for normal distribution  using histograms and shapiro test
```{r}
sets=c("d0_4_expansion")
for (exp in sets){ 
pdf(file=file.path(WORKDIR,paste0("Hist.",exp,".pdf")))
for (grp in levels(dat.st$group)){ 
shap<-shapiro.test(subset(dat.st,perc==exp & group==grp)[,"values"])
hist(subset(dat.st,perc==exp & group==grp)[,"values"], main=paste0(exp," in ",grp),xlab="")
par(mar=c(2,2,4,2))
mtext(paste0("shapiro.test: p= ", shap[2]," ",shap[1]), side=3)
}
dev.off()
}
###some sets are not normally distributed
```


# non-parametric testing using kruskal - wallis test followed by pairwise wilcoxon tests with adjustment
## combined controls
```{r}
#general testing for significant differences of groups for each CFU cat
kruskal.test(values ~ group, data = subset(dat.st,perc=="d0_4_expansion")) #Kruskal-Wallis chi-squared = 30.299, df = 4, p-value = 4.255e-06

#wilcox tests with Benjamini-Hochberg adjustment of pvalues for the individual comparisons
wil.expansion<-wilcox_test(formula=values ~ group,data=subset(dat.st,perc=="d0_4_expansion"), paired = FALSE, p.adjust.method = "BH")
wil.expansion


#add the sig.stats to mean.dat2 to make plotting SA2AML_GEXup_EPanchup_TSS_intersUp
wilcoxonlist<-list(wil.expansion)
names(wilcoxonlist)<-c("d0_4_expansion")
mean.dat$p.adj.signif.vs.CTRL<-""
mean.dat$p.adj.vs.CTRL<-""
mean.dat$perc<-factor(mean.dat$perc)

KDs=c("STAG1KD","STAG2KD","STAG1_2KD","RAD21KD")

for (cat in levels(mean.dat$perc)){
  for (KD in KDs){ 
mean.dat[mean.dat$perc == cat & mean.dat$group==KD, "p.adj.signif.vs.CTRL"]<-subset(wilcoxonlist[[cat]],group1=="CTRL"&group2==KD)["p.adj.signif"]
mean.dat[mean.dat$perc == cat & mean.dat$group==KD, "p.adj.vs.CTRL"]<-subset(wilcoxonlist[[cat]],group1=="CTRL"&group2==KD)["p.adj"]
}}
```

## also check testing for separate control types
```{r}
#general testing for significant differences of groups for each CFU cat
kruskal.test(values ~ group2, data = subset(dat.st,perc=="d0_4_expansion")) #Kruskal-Wallis chi-squared = 30.731, df = 6, p-value = 2.852e-05

#wilcox tests with Benjamini-Hochberg adjustment of pvalues for the individual comparisons
wil.expansion2<-wilcox_test(formula=values ~ group2,data=subset(dat.st,perc=="d0_4_expansion"), paired = FALSE, p.adjust.method = "BH")
wil.expansion2
```




## show adj. p-value vs CTRL significance level with stars in plot and position legend inside plot
### using combined control testing results
```{r}
# subset mean dat expansion
meandatexp<-subset(mean.dat,perc=="d0_4_expansion")
dat.stexp<-subset(dat.st,perc=="d0_4_expansion")

# plot barplots with replicates as dots: expansion rates -  mark controls by subgroup via shape: control subgroup dots mixed within bar
bplotexp2<-ggplot(meandatexp, aes(group, mean,fill=group)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9),alpha=0.8) +
  geom_errorbar(data=meandatexp,aes(ymin=mean-sd, ymax=mean+sd), position = position_dodge(width = 0.9),width=0.1)+
  geom_jitter(data = dat.stexp, aes(x=group,y=values,fill=group,color=group,shape=group2),alpha=1,size=1,width=0.1)+
  scale_fill_manual(values = c("CTRL" = "firebrick1", "STAG1KD" = "darkgoldenrod2", "STAG2KD" = "lightgreen", "STAG1_2KD" = "lightskyblue2", "RAD21KD" = "mediumorchid2")) +
  scale_color_manual(values = c("CTRL" = "firebrick2", "STAG1KD" = "darkgoldenrod3", "STAG2KD" = "seagreen", "STAG1_2KD" = "lightskyblue3", "RAD21KD" = "mediumorchid3")) +
  scale_shape_manual(values = c("CTRL" = 16,"MOCK" = 17,"UNTR" = 2, "STAG1KD" = 16, "STAG2KD" = 16, "STAG1_2KD" = 16, "RAD21KD" = 16)) +
  geom_text(data=meandatexp,aes(y=mean+sd*3),label=meandatexp$p.adj.signif.vs.CTRL,position = position_dodge(width = 0.9),size=8)+
  xlab("") + ylab("expansion d0-d4") +
  scale_x_discrete(labels=c("CTRL\nHSPCs","STAG2\nKD","STAG1\nKD","STAG1+2\nKD","RAD21\nKD")) +
  theme(
  axis.text.y=element_text(size=12),
  axis.text.x=element_text(size=18),
  legend.text=element_text(size=16),
  legend.title=element_text(size=16,face="bold"),
  legend.background = element_blank(),
  legend.box.background = element_rect(colour = "black"),
  axis.title=element_text(size=16,face="bold"),
  plot.title = element_text(size = 18, face = "bold",hjust=0.5),
  panel.border = element_blank(),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.background = element_blank(),
  legend.position="right",
  axis.line = element_line(colour = "black"),)
# plot barplots with replicates as dots: expansion rates - mark controls by subgroup via shape: control subgroup dots shifted within bar
bplotexp3<-ggplot(meandatexp, aes(group, mean,fill=group)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9),alpha=0.8) +
  geom_errorbar(data=meandatexp,aes(ymin=mean-sd, ymax=mean+sd), position = position_dodge(width = 0.9),width=0.1)+
  geom_jitter(data = dat.stexp, aes(x=group,y=values,fill=group,color=group,shape=group2),alpha=1,size=1,position = position_dodge(width = 0.9))+
  scale_fill_manual(values = c("CTRL" = "firebrick1", "STAG1KD" = "darkgoldenrod2", "STAG2KD" = "lightgreen", "STAG1_2KD" = "lightskyblue2", "RAD21KD" = "mediumorchid2")) +
  scale_color_manual(values = c("CTRL" = "firebrick2", "STAG1KD" = "darkgoldenrod3", "STAG2KD" = "seagreen", "STAG1_2KD" = "lightskyblue3", "RAD21KD" = "mediumorchid3")) +
  scale_shape_manual(values = c("CTRL" = 16,"MOCK" = 17,"UNTR" = 2, "STAG1KD" = 16, "STAG2KD" = 16, "STAG1_2KD" = 16, "RAD21KD" = 16)) +
  geom_text(data=meandatexp,aes(y=mean+sd*3),label=meandatexp$p.adj.signif.vs.CTRL,position = position_dodge(width = 0.9),size=8)+
  xlab("") + ylab("expansion d0-d4") +
  scale_x_discrete(labels=c("CTRL\nHSPCs","STAG2\nKD","STAG1\nKD","STAG1+2\nKD","RAD21\nKD")) +
  theme(
  axis.text.y=element_text(size=12),
  axis.text.x=element_text(size=18),
  legend.text=element_text(size=16),
  legend.title=element_text(size=16,face="bold"),
  legend.background = element_blank(),
  legend.box.background = element_rect(colour = "black"),
  axis.title=element_text(size=16,face="bold"),
  plot.title = element_text(size = 18, face = "bold",hjust=0.5),
  panel.border = element_blank(),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.background = element_blank(),
  legend.position="right",
  axis.line = element_line(colour = "black"),)


#export plots
pdf(file=file.path(WORKDIR,"Expansionstats_CohKDs_sd_transfections.as.replicates.padj.2.pdf"), width = 8, height = 5)
plot(bplotexp2)
dev.off()
pdf(file=file.path(WORKDIR,"Expansionstats_CohKDs_sd_transfections.as.replicates.padj.3.pdf"), width = 8, height = 5)
plot(bplotexp3)
dev.off() 

##add labels for replicates
pdf(file=file.path(WORKDIR,"Expansionstats_CohKDs_sd_transfections.as.replicates.padj.lab.pdf"), width = 12, height = 5)
bplotexp+
geom_text_repel(data = dat.stexp,aes(y=values, label=labelID),
                  size=2,
                  nudge_x=0,
                  segment.size=0.3,
                  min.segment.length=0.3,
                  point.padding=0,
                  segment.alpha=0.5,
                  max.overlaps = 50,
                  force = 2,
                  show.legend = FALSE)
dev.off()
```

# generate a nice metadata table for all siRNA Expansionstats_CohKDs_sd_transfections
```{r}
#select columns and rename
CD34cultures<-Expdata[,c("Sample","group_detail","group","d0","d4_living","donor","d0_4_expansion")]
colnames(CD34cultures)<-c("Sample","transfection","group","Cells_d0","Cells_d4","donor","expansion")
CD34cultures<-CD34cultures[,c("Sample","group","donor","transfection","Cells_d0","Cells_d4","expansion")]
CD34cultures[CD34cultures$transfection=="RAD21-2 (RAD21_2031)","transfection"]<-"RAD21_2031"
CD34cultures[CD34cultures$transfection=="RAD21-1 (RAD21_467)","transfection"]<-"RAD21_467"
CD34cultures[CD34cultures$transfection=="RAD21-467","transfection"]<-"RAD21_467"
CD34cultures[CD34cultures$transfection=="RAD21-2031","transfection"]<-"RAD21_2031"
CD34cultures[CD34cultures$transfection=="RAD21 pool","transfection"]<-"RAD21_pool"
CD34cultures$donortreat<-paste0(CD34cultures$donor,"_",CD34cultures$transfection)
#add information on CFU availability
CD34cultures$CFU<-0
CFUstats<-read.csv(file.path(PROJDIR,"CFU","Summary_stats_CD34_KDs_CFU.2.csv"))
CFUfiltdata<-CFUstats %>% drop_na(CFU.total)
CD34cultures[CD34cultures$Sample %in% CFUfiltdata$X...Sample,"CFU"]<-1

#add information on RNAseq availability
RNAsmeta<-read.table(file.path(PROJDIR,"RNAseq/Metadata/Metadata_RNAseq_HSPCs_cohesin_KD.txt"),sep="\t",header=TRUE)
RNAsmeta[RNAsmeta$siRNA12=="Untr","siRNA12"]<-"Untreated"
RNAsmeta[RNAsmeta$siRNA12=="siLuc","siRNA12"]<-"CTRL"
RNAsmeta[RNAsmeta$siRNA12=="siLuc2x","siRNA12"]<-"CTRLx2"
RNAsmeta[RNAsmeta$siRNA12=="poolSA1","siRNA12"]<-"SA1_pool"
RNAsmeta[RNAsmeta$siRNA12=="poolSA2","siRNA12"]<-"SA2_pool"
RNAsmeta[RNAsmeta$siRNA12=="poolRAD21","siRNA12"]<-"RAD21_pool"
RNAsmeta[RNAsmeta$siRNA12=="RAD21pool","siRNA12"]<-"RAD21_pool"
RNAsmeta[RNAsmeta$siRNA12=="poolSA1_poolSA2","siRNA12"]<-"SA1_SA2_pool"
RNAsmeta$donortreat<-paste0(RNAsmeta$donor,"_",RNAsmeta$siRNA12)
CD34cultures$RNAseq<-0
CD34cultures[CD34cultures$donortreat %in% RNAsmeta$donortreat,"RNAseq"]<-1
duplicatesWO_RNAseq<-c("CD34_6_2","CD34_6_4","CD34_7_1","CD34_7_3")
CD34cultures[CD34cultures$Sample %in% duplicatesWO_RNAseq,"RNAseq"]<-0

#add information on ChIPseq etc availability
METALIST<-list()
METALIST[["H3K27ac"]]<-read.csv(file.path(PROJDIR,"ChIP_KD_analysis/H3K27ac","CD34_cohesinKD_ChIP_H3K27ac_meta.csv"),row.names=1)
METALIST[["HiC"]]<-read.csv(file.path(PROJDIR,"HiC_KDs/DeepSeq_Analysis/CD34_cohesinKD_HiC_meta.csv"),row.names=1)
METALIST[["ATAC"]]<-read.csv(file.path(PROJDIR,"ATAC_RS/CD34_cohesinKD_ATAC_meta.csv"),row.names=1)
CHIPSTF<-c("RAD21","CTCF","SA1","SA2","MED12","PU1")
for (c in CHIPSTF){
METALIST[[c]]<-read.csv(file.path(PROJDIR,"ChIP_KD_analysis/Cohesin_CTCF_MED12/Metadata",paste0("CD34_cohesinKD_ChIP_",c,"_meta.csv")),row.names=1)
}
Assays<-c("H3K27ac","ATAC",CHIPSTF,"HiC")
for (a in Assays){
CD34cultures[,a]<-0
CD34cultures[CD34cultures$Sample %in% METALIST[[a]]$Sample,a]<-1
}

write.table(CD34cultures,file.path(PROJDIR,"CD34_CohesinKD_samples_overview.txt"),sep="\t",col.names=TRUE,row.names=FALSE,quote=FALSE)

```


